### 2020-04-22 inst/paper/paper.Rmd
add-wcentroid

---
title: 'climatrends: Precipitation and temperature indices for climate variability analysis in R'
tags:
- climate data
- climatology
- earth science
- evapotranspiration
- precipitation data
- R
- reproducibility
- weather data
authors:
  - name: Kauê de Sousa
    orcid: 0000-0002-7571-7845
    affiliation: "1, 2"
  - name: Jacob van Etten
    orcid: 0000-0001-7554-2558
    affiliation: 2
  - name: Svein Ø. Solberg
    orcid: 0000-0002-4491-4483
    affiliation: 1
affiliations:
  - name: Department of Agricultural Sciences, Inland Norway University of Applied Sciences, Hamar, Norway
    index: 1
  - name: Bioversity International, Rome, Italy
    index: 2
citation_author: de Sousa et. al.
date: "`r format(Sys.time(), '%d %B %Y')`"
year: 2020
bibliography: paper.bib
output: rticles::joss_article
journal: JOSS
---

# Summary

Understand how climate variability drives ecological processes is a key approach to provide recommendations for climate adaptation and biodiversity conservation. **climatrends** aims to provide the toolkit to compute extreme precipitation and temperature indices that serves as input for climate models, crop modelling, trends in climate change and ecology. The climate variables available in **climatrends** were previously used to identify climate patterns in crop growing cycles [@Kehel2016], regional climate change assessment [@Aguilar2005], the effects of agroclimatic indices on the performance of crop varieties [@vanEtten2019] and compare perceptions to climate change with observed data [@DeSousa2018].

# Implementation

# Application: common beans

# Application: time series

# Further development
Integration with other datasets as they become available in `R` via API client packages.

# Acknowledgements

This work was supported by The Nordic Joint Committee for Agricultural and Food Research (grant num. 202100-2817).

# References

### 2020-04-22 R/get_timeseries.R

@@ -307,10 +307,39 @@ get_timeseries.matrix <- function(object,
 #' .nasapower(dates = c("2010-01-01", "2010-01-30"),
 #'            lonlat = lonlat,
 #'            pars = c("T2M_MAX"))
+#' 
+#' 
+#' lonlat <- data.frame(lon = c(-4.96, 11.09),
+#'                      lat = c(37.82, 60.80))
+#'                      
 #' @noRd
-.nasapower <- function(dates, lonlat, pars){
+.nasapower <- function(dates, lonlat, pars, community = NULL, temporal_average = NULL){
   
   message("Getting climate data from NASA POWER \n")
+  
+  if (is.null(community)) {
+    community <- "AG"
+  }
+  
+  if (is.null(temporal_average)) {
+    temporal_average <- "DAILY"
+  }
+  
+  nr <- dim(lonlat)[[1]]
+  
+  # check if data from multiple regions is required
+  xy <- split(lonlat, seq_len(nr))
+  
+  xy <- lapply(xy, function(x){
+    x <- as.vector(unlist(x))
+    .nearest(x, wcentroid)
+  })
+  
+  regions <- unlist(xy)
+  nregions <- length(unique(regions))
+  if (nregions > 1) {
+    message("Fetching data for ", nregions, " regions with 5 x 5 \n")
+  }
   
   # define geographic boundaries for lonlat
   lims <- with(lonlat, c(floor(min(lonlat[,1])), 
@@ -319,10 +348,10 @@ get_timeseries.matrix <- function(object,
                          ceiling(max(lonlat[,2]))))
   
   # get NASA POWER
-  info <- nasapower::get_power(community = "AG",
+  info <- nasapower::get_power(community = community,
                                lonlat = lims,
                                dates = dates,
-                               temporal_average = "DAILY", 
+                               temporal_average = temporal_average, 
                                pars = pars)
   
   info <- as.data.frame(info)

### 2020-04-22 R/print.R

@@ -45,14 +45,12 @@ print.clima_df <- function(x, ...){
     toprint <- rbind(abbs, 
                      he, 
                      rep("", dim(x)[[2]]), 
-                     ta, 
-                     names(x))
+                     ta)
     
     rownames(toprint) <- c("",
                            row.names(he),
                            "---",
-                           rownames(ta),
-                           " ")
+                           rownames(ta))
     
   }
   

### 2020-04-29 inst/paper/paper.Rmd
29042020

@@
 # Implementation
+Describe the main functions, methods and dependencies.
 
 # Application: common beans
+Replicate part of the analysis in van Etten et al [-@vanEtten2019] with the beans data to show how we can use this package to capture the influence of climate variability on crop performance.
+The idea is to show the same PlackettLuce Tree.
 
@@
 
+Pick some random points in Norway (or Scandinavia??) and check how the trends on temperature indices over the last 20 years.
+
 # Further development
-Integration with other datasets as they become available in `R` via API client packages.
+Integration with other datasets as they become available in `R` via API client packages. New indices related to the physiology of crops to be implemented while I work on the rice data.
 

### 2020-04-29 R/rainfall.R

@@ -1,6 +1,6 @@
 #' Rainfall indices
 #'
-#' Rainfall indices over a timespan
+#' Methods to compute rainfall indices over a time series
 #'
 #' @family precipitation functions
 #' @inheritParams get_timeseries
@@ -62,7 +62,6 @@
 #'          day.one = day,
 #'          span = 11)
 #' 
-#' ###################################
 #' \donttest{
 #' ########################################
 #' 
@@ -82,7 +81,7 @@
 #'                   as.sf = FALSE)
 #' rain1
 #' 
-#' # return a sf object
+#' # same as above but return a sf object
 #' rain2 <- rainfall(lonlatsf,
 #'                   day.one = dates,
 #'                   span = 30)
@@ -93,7 +92,6 @@
 #' rain3 <- rainfall(lonlatsf,
 #'                   day.one = dates,
 #'                   span = 30,
-#'                   as.sf = FALSE,
 #'                   timeseries = TRUE,
 #'                   intervals = 7)
 #' rain3
@@ -197,6 +195,17 @@ rainfall.sf <- function(object, day.one, span,
   if (isTRUE(as.sf) & isFALSE(timeseries)) {
     
     indices <- suppressWarnings(sf::st_bind_cols(object, indices))
+    
+  }
+  
+  if (isFALSE(as.sf) & isTRUE(timeseries)) {
+    
+    xy <- .lonlat_from_sf(object)
+    xy <- as.data.frame(xy)
+    xy$id <- 1:dim(xy)[[1]]
+    indices <- merge(xy, indices, by = "id", all.y = TRUE)
+    
+    class(indices) <- union("clima_df", class(indices))
     
   }
   

### 2020-04-29 R/temperature.R

@@ -1,6 +1,6 @@
 #' Temperature indices
 #'
-#' Methods to compute temperature indices over a time period
+#' Methods to compute temperature indices over a time series
 #'
 #' @param object a data.frame (or object that can be coerced to data.frame) 
 #'  with geographical coordinates (lonlat), or an object of class \code{sf}
@@ -91,7 +91,7 @@
 #'                     as.sf = FALSE)
 #' temp1
 #' 
-#' # return a sf object
+#' # same as above but return a sf object
 #' temp2 <- temperature(lonlatsf,
 #'                     day.one = dates,
 #'                     span = 30)
@@ -221,7 +221,12 @@ temperature.sf <- function(object, day.one,
   
   }
   
-  if (isFALSE(as.sf)) {
+  if (isFALSE(as.sf) & isTRUE(timeseries)) {
+    
+    xy <- .lonlat_from_sf(object)
+    xy <- as.data.frame(xy)
+    xy$id <- 1:dim(xy)[[1]]
+    indices <- merge(xy, indices, by = "id", all.y = TRUE)
     
     class(indices) <- union("clima_df", class(indices))
   

### 2020-04-29 tests/testthat/test-ETo.R

@@ -69,24 +69,6 @@ test_that("accept and return a sf", {
   
 })
 
-test_that("accept a tibble", {
-  
-  coord <- as.data.frame(lonlat)
-  coord <- tibble::as_tibble(coord)
-  
-  ev <- ETo(temp,
-           day.one = d,
-           span = 10,
-           lat = rep(0, 10))
-  
-  ev <- round(ev[[1]], 3)
-  
-  ev <- all(e == ev)
-  
-  expect_true(ev)
-  
-})
-
 
 test_that("error, non Date object in day.one", {
   expect_error(

### 2020-05-06 inst/paper/paper.Rmd
improve_get_timespan

--- 2020-04-22 inst/paper/paper.Rmd
@@
 Understand how climate variability drives ecological processes is a key approach to provide recommendations for climate adaptation and biodiversity conservation.
-**climatrends** aims to provide the toolkit to compute extreme precipitation and temperature indices that serves as input for climate models, crop modelling, trends in climate change and ecology.
-The climate variables available in **climatrends** were previously used to identify climate patterns in crop growing cycles [@Kehel2016], regional climate change assessment [@Aguilar2005], the effects of agroclimatic indices on the performance of crop varieties [@vanEtten2019] and compare perceptions to climate change with observed data [@DeSousa2018].
+**climatrends** aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [**add-reference**].
 
@@
 
+Four main functions are provided, `ETo()`, `GDD()`, `rainfall()` and `temperature()` with methods [@RCoreTeam] for classes 'data.frame', 'matrix' (or array), and 'sf' [@sf].
+
 # Application: common beans
+
+Replicate part of the analysis in van Etten et al [-@vanEtten2019] with the beans data to show how we can use this package to capture the influence of climate variability on crop performance.
+The idea is to show the same PlackettLuce Tree.
 

### 2020-05-06 R/ETo.R

@@ -233,8 +233,15 @@ ETo.sf <- function(object, day.one, span = NULL, Kc = 1, as.sf = TRUE, ...){
 # Compute evapotranspiration
 .eto <- function(day, night, Kc, p) {
   
-  # calculate Tmean
-  Tmean <- (rowMeans(day, na.rm = TRUE) + rowMeans(night, na.rm = TRUE)) / 2
+  temp <- cbind(day, value2 = night$value)
+  
+  temp <- split(temp, temp$id)
+  
+  Tmean <-lapply(temp, function(x){
+    (mean(x$value, na.rm = TRUE) + mean(x$value2, na.rm = TRUE)) / 2
+  })
+  
+  Tmean <- do.call("rbind", Tmean)
   
   # evapotranspiration
   eto <- p * (0.46 * Tmean + 8) * Kc
@@ -260,4 +267,3 @@ ETo.sf <- function(object, day.one, span = NULL, Kc = 1, as.sf = TRUE, ...){
   return(p)
 
 }
- 

### 2020-05-06 R/GDD.R

@@ -8,9 +8,15 @@
 #' 
 #' @inheritParams temperature
 #' @param degree.days an integer for the degree-days required by the 
-#'  organism (look for the physiology of the focal organism)
+#'  organism (look for the physiology of the focal organism). Optional if 
+#'  \var{return} = \code{"daily"}
 #' @param base an integer for the base temperature
-#' @return The number of days which were required to reach the growing degree-days.
+#' @param return character (one of \code{"ndays"} or \code{"daily"}) to select if 
+#'  the function returns the number of days to reach the GDD or the raw daily values 
+#'  of the GDD
+#' @return 
+#'  The number of days which were required to reach the growing degree-days, or 
+#'   the degree days in each day during the time series
 #' @family temperature functions
 #' @details 
 #' The \code{array} method assumes that \var{object} contains climate data provided 
@@ -82,7 +88,7 @@
 #' 
 #'}
 #' @export
-GDD <- function(object, day.one, degree.days,
+GDD <- function(object, day.one, degree.days = NULL,
                 base = 10, ...)
 {
   UseMethod("GDD")
@@ -91,8 +97,8 @@ GDD <- function(object, day.one, degree.days,
 #' @rdname GDD
 #' @method GDD default
 #' @export
-GDD.default <- function(object, day.one, degree.days,
-                        base = 10, span = NULL, ...){
+GDD.default <- function(object, day.one, degree.days = NULL,
+                        base = 10, span = NULL, return = "ndays", ...){
   
   dots <- list(...)
   pars <- dots[["pars"]]
@@ -116,7 +122,7 @@ GDD.default <- function(object, day.one, degree.days,
   
   night <- dat[[pars[[2]]]]
   
-  result <- .gdd(day, night, base, degree.days)
+  result <- .gdd(day, night, base, degree.days, return)
   
   return(result)
 }
@@ -125,8 +131,8 @@ GDD.default <- function(object, day.one, degree.days,
 #' @rdname GDD
 #' @method GDD array
 #' @export
-GDD.array <- function(object, day.one, degree.days,
-                      base = 10, ...){
+GDD.array <- function(object, day.one, degree.days = NULL,
+                      base = 10, return = "ndays", ...){
   
   if(dim(object)[[2]] == 2) {
     UseMethod("GDD", "default")
[…]

### 2020-05-06 R/get_timeseries.R

@@ -1,6 +1,6 @@
 #' Time series climate data
 #' 
-#' General functions and methods to concatenate climate data across a time series.
+#' General functions and methods to concatenate climate data across a time series
 #' 
 #' @param object a \code{data.frame} (or any othwer object that can be coerced to 
 #'  data.frame) with geographical coordinates (lonlat), or an object of class 
@@ -24,7 +24,7 @@
 #'  and will be inputted in the format of a named matrix. 
 #'  See help("modis", "climatrends") for examples.
 #' 
-#' Available remote \var{source}s are: "nasapower"
+#' Available remote \var{source}s: "nasapower"
 #' 
 #' Additional arguments:
 #' 
@@ -40,23 +40,27 @@
 #' # Using local sources
 #' data("modis", package = "climatrends")
 #' 
-#' day <- as.Date("2013-10-28", format = "%Y-%m-%d")
+#' set.seed(9271)
+#' span <- as.integer(runif(10, 6, 15)) 
 #' 
-#' get_timeseries(chirp, day, span = c(10,11))
+#' g <- get_timeseries(modis[,,1], "2013-10-28", span = span)[[1]]
 #' 
+#' # library("ggplot2")
+#' # 
+#' # ggplot(g, aes(x = date, y = value, group = factor(id), color = factor(id))) +
+#' #   geom_point() +
+#' #   geom_line()
 #' 
 #' ########################################################
 #' \donttest{
 #' # Fetch data from NASA POWER using 'sf' method
-#' 
 #' data("lonlatsf", package = "climatrends")
 #' 
-#' do <- as.Date(17667, origin = "1970-01-01")
-#' 
 #' g <- get_timeseries(lonlatsf, 
-#'                     day.one = do, 
-#'                     span = 10,
+#'                     day.one = "2018-05-16", 
+#'                     last.day = "2018-05-30",
 #'                     pars = c("PRECTOT", "T2M", "T10M"))
+#' 
 #' 
 #' g
 #' }
@@ -82,7 +86,7 @@ get_timeseries.default <- function(object, day.one, span = NULL, last.day = NULL
     days.before <- 0
   }
   
-  sts <- .set_span_length(day.one, span, last.day, days.before)
+  sts <- .st_span(day.one, span, last.day, days.before)
 
   object <- as.data.frame(object)
   
@@ -92,14 +96,13 @@ get_timeseries.default <- function(object, day.one, span = NULL, last.day = NULL
                lonlat = object,
                pars = pars)
   
-  
[…]

### 2020-05-06 R/internal_functions.R

@@ -159,83 +159,43 @@
 #' Coerce a vector of characters or integers into class 'Date'
 #' @param x a vector
 #' @examples 
-#' x <- c("1970-01-01", "2020-90-01")
-#' .coerce2Date(x)
+#' # valid entries
+#' .coerce2Date(c(819, 9811))
 #' 
-#' x <- c(819, 9811)
-#' .coerce2Date(x)
+#' .coerce2Date(c("1970-01-01", "2020-09-01"))
 #' 
-#' x <- letters[1:10]
-#' .coerce2Date(x)
+#' .coerce2Date(as.Date(12, origin = "1970-01-01"))
+#' 
+#' # some errors
+#' .coerce2Date(letters[1:10])
+#' 
+#' .coerce2Date(c("1970-01-01", "2020-90-01"))
+#' 
 #' @noRd
 .coerce2Date <- function(x) {
   
-  if (isTRUE(.is_Date(x))) {
-    
-    return(x)
-  
-  }
-  
   if (is.character(x)) {
-    x <- as.Date(x, format = "%Y-%m-%d")
+      
+      x <- as.Date(x, format = "%Y-%m-%d")
+
   }
   
   if (any(is.integer(x), is.numeric(x))) {
+    
     x <- as.Date(x, origin = "1970-01-01")
+  
   }
   
   if (isFALSE(.is_Date(x))) {
-      stop("No visible method to coerce given dates to as.Date \n")
+      stop("No visible method to coerce",
+           " given dates to as.Date \n")
   }
   
   if (any(is.na(x))) {
-    stop("Visible methods to coerce given dates to as.Date returning NAs \n")
+    stop("Visible method to coerce given",
+         " dates to as.Date returning NAs \n")
   }
   
   return(x)
-  
-}
-
-
-
-.set_long <- function(object, id) {
-  
-  object <- split(object, object[, id])
-  
-  object <- lapply(object, function(x){
-    
-    x <- t(x)
-    
-    x <- cbind(row.names(x), x)
-    
-    i <- x[id, 2]
-    
-    x <- cbind(x, id = i)
-    
-    x <- x[ x[,1] != id , ]
-    
-    x <- as.data.frame(x, stringsAsFactors = FALSE)
-    
-    names(x) <- c("variable", "value", "id")
-    
-    x <- x[ ,c("id", "variable", "value")]
-    
-    return(x)
-    
-  })
-  
-  object <- do.call("rbind", object)
-  
[…]

### 2020-05-06 R/rainfall.R

@@ -134,7 +134,7 @@ rainfall.default <- function(object, day.one, span = NULL,
   
   rain <- get_timeseries(object, day.one, span, pars = pars, ...)[[1]]
   
-  indices <- .rainfall_indices(rain, timeseries, intervals, day.one, span, ...)
+  indices <- .rainfall_indices(rain, timeseries, intervals)
   
   return(indices)
   
@@ -156,7 +156,7 @@ rainfall.matrix <- function(object, day.one, span = NULL,
   
   rain <- get_timeseries(object, day.one, span, ...)[[1]]
   
-  indices <- .rainfall_indices(rain, timeseries, intervals, day.one, span, ...)
+  indices <- .rainfall_indices(rain, timeseries, intervals)
   
   return(indices)
   
@@ -181,7 +181,7 @@ rainfall.sf <- function(object, day.one, span = NULL,
   
   rain <- get_timeseries(object, day.one, span, pars = pars, ...)[[1]]
   
-  indices <- .rainfall_indices(rain, timeseries, intervals, day.one, span, ...)
+  indices <- .rainfall_indices(rain, timeseries, intervals)
   
   if (all(as.sf, timeseries)) {
     
@@ -216,41 +216,18 @@ rainfall.sf <- function(object, day.one, span = NULL,
 }
 
 
-.rainfall_indices <- function(rain, timeseries, intervals, day.one, span = NULL, last.day = NULL){
+.rainfall_indices <- function(rain, timeseries, intervals){
   
   index <- c("MLDS","MLWS","R10mm","R20mm","Rx1day",
              "Rx5day","R95p","R99p","Rtotal","SDII")
   
-  nr <- dim(rain)[[1]]
-  
-  nc <- dim(rain)[[2]]
+  nr <- max(unique(rain$id))
   
   if (isTRUE(timeseries)) {
     
-    # check if day.one is a 'Date' else try to coerce to Date
-    if (!.is_Date(day.one)) {
-      
-      day.one <- .coerce2Date(day.one)
-      
-    }
+    ids <- unique(rain$id)
     
-    # the timespan
-    if (!is.null(span)) {
-      span <- as.vector(t(span)) 
-    }
-    
-    # or from last.day
-    if (!is.null(last.day)) {
-      
-      if (!.is_Date(last.day)) {
-        
-        last.day <- .coerce2Date(last.day)
-        
-      }
-      
[…]

### 2020-05-06 R/temperature.R

@@ -125,8 +125,7 @@ temperature <- function(object, day.one,
 #' @export
 temperature.default <- function(object, day.one, 
                                 span = NULL, timeseries = FALSE,
-                                intervals = 5,
-                                ...){
+                                intervals = 5, ...){
   
   dots <- list(...)
   pars <- dots[["pars"]]
@@ -151,7 +150,7 @@ temperature.default <- function(object, day.one,
   
   night <- dat[[pars[[2]]]]
   
-  indices <- .temperature_indices(day, night, timeseries, intervals, day.one, span, last.day)
+  indices <- .temperature_indices(day, night, timeseries, intervals)
   
   class(indices) <- union("clima_df", class(indices))
   
@@ -180,7 +179,7 @@ temperature.array <- function(object, day.one,
   
   night <- get_timeseries(object[, , 2], day.one, span, ...)[[1]]
   
-  indices <- .temperature_indices(day, night, timeseries, intervals, day.one, span, last.day)
+  indices <- .temperature_indices(day, night, timeseries, intervals)
   
   return(indices)
   
@@ -209,7 +208,7 @@ temperature.sf <- function(object, day.one,
   
   night <- dat[[pars[[2]]]]
   
-  indices <- .temperature_indices(day, night, timeseries, intervals, day.one, span, last.day)
+  indices <- .temperature_indices(day, night, timeseries, intervals)
   
   if (all(as.sf, timeseries)) {
   
@@ -248,40 +247,21 @@ temperature.sf <- function(object, day.one,
 #' @param night a data.frame with the night temperature
 #' @inheritParams temperature
 #' @noRd
-.temperature_indices <- function(day, night, timeseries, intervals, day.one, span = NULL, last.day = NULL){
+.temperature_indices <- function(day, night, timeseries, intervals){
   
   index <- c("maxDT", "minDT", "maxNT", "minNT",
              "DTR", "SU", "TR", "CFD",
              "WSDI", "CSDI", "T10p", "T90p")
   
-  n <- nrow(day)
+  nr <- max(unique(day$id))
   
   if (isTRUE(timeseries)) {
     
-    # check if day.one is a 'Date' else try to coerce to Date
-    if (!.is_Date(day.one)) {
[…]

### 2020-05-06 inst/paper/beans.R

@@ -1,0 +1,61 @@
+#set paraters for spatial analysis
+my_proj <- "+proj=longlat +datum=WGS84"
+my_ext <- raster::extent(-87, -83, 12, 15)
+modis_factor <- 0.02
+k_to_c <- 273.15
+
+#add modis data 
+cat("Readind MODIS data. Time:", date(), "\n")
+# #read modis
+# dates <- NULL
+# for(i in c(2003:2016))
+#   dates <- c(dates, c(seq(as.Date(paste0(i, "-01-01"), "%Y-%m-%d"), as.Date(paste0(i, "-12-31"), "%Y-%m-%d"), 8)))
+# dates <- as.Date(c(dates, seq(as.Date("2017-01-01", "%Y-%m-%d"), as.Date("2017-09-06", "%Y-%m-%d"), 8)), origin = "1970-01-01")
+# 
+# full_dates <- seq(dates[1], dates[length(dates)], 1)
+# #read modis day and night
+# modis <- array(NA, dim = c(n, length(dates), 2), dimnames = list(NULL, as.character(dates), c(1:2)))
+# for(i in 1:2){
+#   if(i == 1 ) m="day" else m="night"
+#   file <- raster::stack(list.files("E:/rasters/modis/smooth_Nicaragua/", pattern = m, full.names = T))
+#   raster::crs(file) <- my_proj
+#   x <- raster::extract(file, mydata[,c("lon","lat")])
+#   x <- x * modis_factor - k_to_c #convert to Celsius
+#   x[x<1] <- NA #remove wrong data
+#   modis[ , , i] <- x
+# }
+# sum(is.na(modis))
+# #apply linear interpolation to the 8-day temperature
+# modis_approx <- array(NA, c(n, length(full_dates), 2), dimnames = list(NULL,  as.character(full_dates), c(1:2) ))
+# for(i in 1:dim(modis)[1]){
+#   modis_approx[i,,1] <- approx(modis[i,,1], n = length(full_dates))[[2]]
+#   modis_approx[i,,2] <- approx(modis[i,,2], n = length(full_dates))[[2]]
+# }
+# save(modis, modis_approx, file = "./processing/modis_nicaragua.RData")
+load("./processing/modis_nicaragua.RData")
+
+# # #take CHIRPS info
+# cat("Readind CHIRPS data. Time:", date(), "\n")
+# #source CHIRPS (precipitation data)
+# #http://chg.geog.ucsb.edu/data/chirps/
+# years <- seq(2003 , 2017, 1)
[…]

### 2020-05-06 tests/testthat/test-GDD.R

@@ -10,14 +10,32 @@ test_that("right values array method", {
   dg <- GDD(object = temp,
             day.one = d, 
             degree.days = 45,
-            base = 10,
-            span = 12)
+            base = 10)
   
   dg <- all(realv == dg[, 1])
   
   expect_true(dg)
   
 })
+
+realv <- c(11, 11, 11, 11, 11, 
+           12, 12, 12, 12, 12)
+# return daily gdd
+test_that("return daily gdd", {
+  
+  gdd <- GDD(object = temp[1:2,,],
+             day.one = d[1:2,], 
+             span = 5,
+             base = 10,
+             return = "daily")$gdd
+  
+  gdd <- round(gdd, 0)
+  
+  istrue <- all(gdd == realv)
+  
+  expect_true(istrue)
+})
+
 
 realv <- c(7,6,6)
 test_that("nasapower works", {

### 2020-05-06 tests/testthat/test-rainfall.R

@@ -47,3 +47,13 @@ test_that("larger span", {
 })
 
 
+test_that("non valid date", {
+  
+  expect_error(
+    rainfall(chirp,
+             day.one = "2013-02-31",
+             span = 11)
+  )
+  
+})
+

### 2020-05-07 R/get_timeseries.R
adjust-travis-yml

@@ -34,7 +34,7 @@
 #' \code{days.before}: an integer for the number of days before \var{day.one} to be 
 #'  included in the timespan.
 #' 
-#' @return An object with time series climate data for the chosen period
+#' @return A list of data.frame(s) with the class \code{clima_df} 
 #' @family GET functions
 #' @examples 
 #' # Using local sources

### 2020-05-07 inst/paper/paper.Rmd
adjust-travis-yml
fix-methods-in-ETo-and-GDD

@@
 Understand how climate variability drives ecological processes is a key approach to provide recommendations for climate adaptation and biodiversity conservation.
-**climatrends** aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [**add-reference**].
+**climatrends** aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
 
-# Implementation
+[...continue...]
+
+# Methods and features
+
+## Implementation
 
@@
 
-# Application: common beans
+[...continue...]
+
+## Growing degree-days
+
+Growing degree-days (gdd) is an heuristic tool in phenology that measures heat accumulation and is used to predict plant and animal development rates [@Prentice1992].
+Growing degree-days are calculated by taking the integral of warmth above a base temperature ($T_{0}$).
+The function `GDD()` applies by default the following equation.
+
+Equation [1]
+
+$$GDD = \frac{T_{max} + T_{min}}{2} - T_{0}$$
+
+Where $T_{max}$ is the maximum temperature in the given day, $T_{min}$ is the minimum temperature in the given day and $T_{0}$ is the minimum temperature for growth (as per the physiology of the focal organism).
+
+For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two variants of the previous equation to adjust either $T_{mean}$ (variant A) or $T_{min}$ (variant B).
+The variant A changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.
+
+Equation [2]
+
+$$ GDD = max \left(\frac{T_{max} + T_{min}}{2} - T_{0}, \; 0 \right)$$
+
+The variant B, is calculated using Equation 1, but adjusts $T_{min}$ to $T_{0}$ if $T_{min} < T_{o}$, and is adjusted as follows.
+
+Equation [3]
+
+$$ T_{min} < T_{o} \; \rightarrow \; T_{min} = T_{o} $$
+
+In `GDD()` these equations are defined as 'default', 'variant_a' and 'variant_b', respectively, and can be selected using the argument `equation`.
+
+Additionally, the function may return the number of days that a given organism required to reach a certain number of accumulated degree-days or the raw gdd in each observed day.
+These values are defined by 'ndays' or 'gdd' and can be adjusted using the argument `return.as`.
+The required accumulated gdd is defined with argument `degree.days`.
+For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset of photosynthesis [@JWu2013].
+
+## Reference evapotranspiration
+
+The reference evapotranspiration measures the influence of the climate on a given organism (generally a crop species) water needs [@Brouwer1986].
+The function `ETo()` applies the Blaney-Criddle formula, a general theoretical method used when no measured data on pan evaporation is available locally, using the following equation.
+
+Equation [4]
+
+$$ETo = p \times \left(0.46 \times \frac{T_{max} + T_{min}}{2} + 8 \right) \times K_c$$
+
+Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the relationship between the reference organism and the actually grown of the organism and may change according to the growth stage.
+
+# Examples
+
+## Common beans
 
@@
 
-# Application: time series
+## Time series
 
@@
 # Further development
+
 Integration with other datasets as they become available in `R` via API client packages. New indices related to the physiology of crops to be implemented while I work on the rice data.

### 2020-05-07 R/ETo.R

@@ -1,6 +1,6 @@
-#' Evapotranspiration
+#' Reference evapotranspiration
 #' 
-#' Evapotranspiration using the Blaney-Criddle method. This is general  
+#' Reference evapotranspiration using the Blaney-Criddle method. This is general  
 #' theoretical method used when no measured data on pan evaporation 
 #' is available locally. 
 #' 
@@ -9,6 +9,8 @@
 #'  mean daily percentage of annual daytime hours based on the latitude and month.
 #'  This is extracted automatically in the \code{sf} method. See details
 #' @param Kc a numeric value for the crop factor for water requirement
+#' @param p optional if \var{lat} is given, a numeric for the mean daily percentage 
+#'  of annual daytime hours (p = 0.27 by default)
 #' @return The evapotranspiration in mm/day
 #' @details 
 #' When \var{lat} is provided, it is combined with the month provided in 
@@ -28,9 +30,13 @@
 #' 
 #' Additional arguments:
 #' 
-#' \code{last.day}: optional to \var{span}, an object of class \code{Date} or
+#' \code{last.day}: an object (optional to \var{span}) of class \code{Date} or
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
-#'  YYYY-MM-DD)  for the last day of the time series
+#'  YYYY-MM-DD) for the last day of the time series
+#'  
+#' \code{span}: an integer (optional to \var{last.day}) or a vector with 
+#'  integers (optional if \var{last.day} is given) for the length of 
+#'  the time series to be captured
 #' 
 #' \code{data.from}: character for the source of climate data. Current remote data 
 #'  is: 'nasapower'
@@ -57,41 +63,35 @@
 #'     day.one = "2013-10-28",
 #'     span = 10,
 #'     Kc = 0.92)
-#'     
+#' 
 #' \donttest{
 #' ######################################
-#'  
+#' 
 #' # Using remote data
 #' # random geographic locations around bbox(11, 12, 55, 58)
 #' set.seed(826128)
 #' lonlat <- data.frame(lon = runif(2, 11, 12),
 #'                      lat = runif(2, 55, 58))
[…]

### 2020-05-07 R/GDD.R

@@ -42,7 +42,7 @@
 #' 
 #' \code{last.day}: an object (optional to \var{span}) of class \code{Date} or
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
-#'  YYYY-MM-DD)  for the last day of the time series
+#'  YYYY-MM-DD) for the last day of the time series
 #'  
 #' \code{span}: an integer (optional to \var{last.day}) or a vector with 
 #'  integers (optional if \var{last.day} is given) for the length of 
@@ -156,14 +156,14 @@ GDD.array <- function(object, day.one, base = 10,
   span <- dots[["span"]]
   last.day <- dots[["last.day"]]
   
-  # coerce to vector
   day.one <- as.vector(t(day.one))
   
+  # check if day.one is a 'Date' else try to coerce to Date
+  if (isFALSE(.is_Date(day.one))) {
+    day.one <- .coerce2Date(day.one)
+  }
+  
   if (all(is.null(span), is.null(last.day))) {
-    
-    if (isFALSE(.is_Date(day.one))){
-      day.one <- .coerce2Date(day.one)
-    }
     do <- as.character(max(day.one))
     do <- match(do, dimnames(object[,,1])[[2]])
     span <- dim(object)[[2]] - do

### 2020-05-08 inst/paper/paper.Rmd
add-crop-sensitive-indices

@@
 bibliography: paper.bib
+#output: pdf_document
 output: rticles::joss_article
@@
 For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset of photosynthesis [@JWu2013].
+In that case, the function will calculate the growing degree-days ($gdd$) and sum up the values until it reaches the defined gdd (105 $^\circ C$) and return the number of days needed in the given season ($GDD_{ndays}$), as follows.
+
+Equation [4]
+
+$$\parallel GDD_{ndays} \parallel \: = \; \sum^{ggd_1}_{gdd_n}$$
 
@@
 The reference evapotranspiration measures the influence of the climate on a given organism (generally a crop species) water needs [@Brouwer1986].
-The function `ETo()` applies the Blaney-Criddle formula, a general theoretical method used when no measured data on pan evaporation is available locally, using the following equation.
+The function `ETo()` applies the Blaney-Criddle method, a general theoretical method used when only air-temperature is available locally.
+It should be noted that this method is not very accurate and aims to provide the order of magnitude of evapotranspitation.
+The reference evapotranspiration is calculated using the following equation.
 
-Equation [4]
+Equation [5]
 
@@
 
-Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the relationship between the reference organism and the actually grown of the organism and may change according to the growth stage.
+Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the factor for organism water need may change according to the growth stage.
+
+The percentage of daytime hours ($p$) is calculated internally by most methods in `ETo()` (except the array method) using the given latitude (taken from the inputted `object`) and date (taken from the inputted `day.one`).
+It matches the latitude and date with a table of daylight percentage derived from Brouwer and Heibloem [-@Brouwer1986].
+The table can be verified using `climatrends:::daylight`.
 

### 2020-05-08 R/GDD.R

@@ -1,9 +1,8 @@
 #' Growing degree-days
 #' 
-#' Compute Growing degree-days. This a heuristic tool in phenology 
-#' that measures heat accumulation and is used to predict plant and animal 
-#' development rates. Growing degree-days are calculated by taking the 
-#' integral of warmth above a base temperature.
+#' This a heuristic tool in phenology that measures heat accumulation and 
+#' is used to predict plant and animal development rates. Growing degree-days
+#' are calculated by taking the integral of warmth above a base temperature.
 #' 
 #' @inheritParams temperature
 #' @param base an integer for the minimum temperature for growth (as per the physiology 
@@ -240,41 +239,38 @@ GDD.clima_ls <- function(object,
   
   temp <- cbind(day, tmin = night$value)
   
+  names(temp)[names(temp)=="value"] <- "tmax"
+  
   temp <- split(temp, temp$id)
   
+  suppressWarnings(
   Y <- lapply(temp, function(x){
-    
-    tmax <- x$value
-    tmin <- x$tmin
-    
     
     if (isTRUE(equation == "default")) {
       
-      y <- ((tmax + tmin) / 2) - base
+      y <- ((x$tmax + x$tmin) / 2) - base
     
     }
     
     if (isTRUE(equation == "variant_a")) {
-      # adjust Tmean if Tmean < base
-      tmean <- (tmax + tmin) / 2
+      # take the maximum between y and base
+      y <- ((x$tmax + x$tmin) / 2) - base
       
-      tadj <- ifelse(tmean < base, base, tmean)
-      
-      y <- tadj - base
+      y <- max(y, 0)
       
     }
     
     if (isTRUE(equation == "variant_b")) {
       # set Tmin = base if Tmin < base
-      tadj <- ifelse(tmin < base, base, tmin)
+      tadj <- ifelse(x$tmin < base, base, x$tmin)
       
-      y <- ((tmax + tadj) / 2) - base
+      y <- ((x$tmax + tadj) / 2) - base
       
     }
     
     if (isTRUE(return.as == "ndays")) {
       
-      y <- y[!is.na(y)]
+      y[is.na(y)] <- 0
       
       # sum temperature values until reach the defined degree days
       for (d in seq_along(y)) {
@@ -299,7 +295,7 @@ GDD.clima_ls <- function(object,
     }
   
   })
-  
+  )
   result <- do.call("rbind", Y)
   
   if (isTRUE(return.as == "ndays")) {

### 2020-05-08 R/crop_sensitive.R

@@ -1,0 +1,97 @@
+#' Crop sensitive indices
+#' 
+#' Compute crop sensitive indices
+#' 
+#' @inheritParams temperature
+#' 
+#' @source 
+#' 
+#' Challinor et al. (2016). Nature Climate Change, 6, 954-958(2016)
+#' \cr\url{https://doi.org/10.1038/nclimate3061}
+#' 
+#' @export
+crop_sensitive <- function(object, ...){
+  return(NA)
+}
+
+#' High temperature stress
+#' 
+#' Calculated using daily mean temperature (or the daily max temperature) 
+#'  and given as the percentage number of days a certain threshold is exceeded 
+#' 
+#' @param x maximum day temperature
+#' @param y optional, minimum temperature
+#' @return the HTS index
+#' @examples 
+#' # hts with day max temperature 
+#' x <- modis[3, ,1]
+#' .hts(x, threshold = 36)
+#' 
+#' # hts with mean daily temperature
+#' x <- modis[1,,1]
+#' y <- modis[1,,2]
+#' .hts(x,y, threshold = 32)
+#' 
+#' # lethal temperatures
+#' .hts(x,y, threshold = 43)
+#' 
+#' @noRd
+.hts <- function(x, y = NULL, threshold = NULL){
+  
+  # if y is provided than take the mean of x and y
+  if (isFALSE(is.null(y))) {
+    x <- (x + y) / 2
+  }
+  
+  # length of x
+  l <- length(x)
+  
+  # number of values above the threshold
+  i <- sum(x > threshold, na.rm = TRUE)
+  
+  # fraction of x above the threshold
+  r <- i / l
+  
+  return(r)
+  
+}
+
+#' Crop duration index
+#' 
+#' Calculated using the mean growing season temperature (or the mean daily
+#'  maximum growing season temperature). If Tmean > threshold then 
+#'  CDI = Tmean - threshold, otherwise CDI = 0
+#' 
+#' @param x maximum day temperature
+#' @param y optional, minimum temperature
+#' @return the HTS index
+#' @examples 
+#' # cdi with mean daily maximum temperature
+#' x <- modis[3, ,1]
+#' .cdi(x, threshold = 36)
+#' 
+#' # cdi with mean temperature
+#' x <- modis[1,,1]
+#' y <- modis[1,,2]
+#' .cdi(x,y, threshold = 32)
+#' 
+#' @noRd
[…]

### 2020-05-08 R/print.R

@@ -4,11 +4,14 @@ print.clima_df <- function(x, ...){
   
   x <- as.data.frame(x, stringAsFactor = FALSE)
   
+  rownames(x) <- seq_len(dim(x)[[1]])
+  
   classes <- lapply(x, function(y){
     class(y)
   })
   
   classes <- as.vector(unlist(classes))
+  
   class_abb = c(list = "<list>", integer = "<int>", numeric = "<dbl>", 
                 character = "<chr>", Date = "<date>", complex = "<cpl>", 
                 factor = "<fct>", POSIXct = "<POSc>", logical = "<lgl>", 

### 2020-05-09 R/ETo.R
array-method-get_timeseries

@@ -178,11 +178,9 @@ ETo.array <- function(object, day.one, Kc = 1, lat = NULL, p = 0.27, ...){
     p <- .p_daytime(lat = lat, month = m)
   }
   
-  day <- get_timeseries(object[, , 1], day.one, span, last.day)[[1]]
+  ts <- get_timeseries(object, day.one, span, last.day)
   
-  night <- get_timeseries(object[, , 2], day.one, span, last.day)[[1]]
-  
-  result <- .eto(day, night, Kc, p)
+  result <- .eto(ts[[1]], ts[[2]], Kc, p)
   
   return(result)
   

### 2020-05-09 R/GDD.R

@@ -168,11 +168,9 @@ GDD.array <- function(object, day.one, base = 10,
     span <- dim(object)[[2]] - do
   }
   
-  day <- get_timeseries(object[, , 1], day.one, span, last.day)[[1]]
+  ts <- get_timeseries(object, day.one, span = span, last.day = last.day, ...)
   
-  night <- get_timeseries(object[, , 2], day.one, span, last.day)[[1]]
-  
-  result <- .gdd(day, night, base, degree.days, equation, return.as)
+  result <- .gdd(ts[[1]], ts[[2]], base, degree.days, equation, return.as)
   
   return(result)
 }

### 2020-05-09 R/get_timeseries.R

@@ -5,7 +5,8 @@
 #' @param object a \code{data.frame} (or any other object that can be coerced to 
 #'  data.frame) with geographical coordinates (lonlat), or an object of class 
 #'  \code{sf} with geometry 'POINT' or 'POLYGON', or a named \code{matrix} with 
-#'  climate data. See details.   
+#'  climate data, or an array with two dimensions for max and min temperature.
+#'  See details.   
 #' @param day.one a vector of class \code{Date} or any other object that can be 
 #'  coerced to \code{Date} (e.g. integer, character YYYY-MM-DD) for the starting 
 #'  day to capture the climate data
@@ -39,18 +40,18 @@
 #' @family GET functions
 #' @examples 
 #' # Using local sources
+#' # an array with temperature data
 #' data("modis", package = "climatrends")
 #' 
 #' set.seed(9271)
-#' span <- as.integer(runif(10, 6, 15)) 
+#' span <- as.integer(runif(10, 6, 15))
 #' 
-#' g <- get_timeseries(modis[,,1], "2013-10-28", span = span)[[1]]
+#' get_timeseries(modis, "2013-10-28", span = span)
 #' 
-#' # library("ggplot2")
-#' # 
-#' # ggplot(g, aes(x = date, y = value, group = factor(id), color = factor(id))) +
-#' #   geom_point() +
-#' #   geom_line()
+#' # matrix with precipitation data
+#' data("chirp", package = "climatrends")
+#' 
+#' get_timeseries(chirp, "2013-10-28", span = span)
 #' 
 #' ########################################################
 #' \donttest{
@@ -69,7 +70,7 @@
 #' @importFrom sf st_centroid st_geometry_type st_as_sf
 #' @importFrom stats dist hclust cutree
 #' @export
-get_timeseries <- function(object, day.one, span = NULL, last.day = NULL, ...) {
+get_timeseries <- function(object, day.one, ...) {
   
   UseMethod("get_timeseries")
   
@@ -196,6 +197,24 @@ get_timeseries.matrix <- function(object, day.one, span = NULL, last.day = NULL,
   return(r)
 }
 
+
+#' @rdname get_timeseries
+#' @method get_timeseries array
+#' @export
+get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL, 
+                                  ...){
+  
+  
+  dm1 <- get_timeseries(object[,,1], day.one, span = span, last.day = last.day, ...)
[…]

### 2020-05-09 R/temperature.R

@@ -176,11 +176,9 @@ temperature.array <- function(object, day.one,
   # coerce to data.frame
   day.one <- as.vector(t(day.one))
   
-  day <- get_timeseries(object[, , 1], day.one, span, ...)[[1]]
+  ts <- get_timeseries(object, day.one, span = span, last.day = last.day, ...)
   
-  night <- get_timeseries(object[, , 2], day.one, span, ...)[[1]]
-  
-  indices <- .temperature_indices(day, night, timeseries, intervals)
+  indices <- .temperature_indices(ts[[1]], ts[[2]], timeseries, intervals)
   
   return(indices)
   

### 2020-05-14 inst/paper/paper.Rmd
array-method-get_timeseries
add-anthesis-function

@@
 bibliography: paper.bib
-#output: pdf_document
-output: rticles::joss_article
-journal: JOSS
+output: pdf_document
 ---
 
-# Summary
+>Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
+^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
+^2^ Bioversity International, 00054 Maccarese, Rome, Italy  
+^[*]^Correspondence should be addressed to: <kaue.desousa@inn.no>
 
-Understand how climate variability drives ecological processes is a key approach to provide recommendations for climate adaptation and biodiversity conservation.
+\
+
+# Introduction
+
+Abiotic factors plays an important role in most ecological and crop systems that depends on certain levels of temperature, light and precipitation (and their interplay) to initiate important physiological events [@PlantEcology].
+In the walk of climate change, understand how these factors drives the physiological processes is a key approach to provide recommendations for adaptation and biodiversity conservation.
+If translated to stress
+
+Raw temperature and precipitation values are ... to describe how organisms interacts with the enviroment
+
 **climatrends** aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
@@
 Four main functions are provided, `ETo()`, `GDD()`, `rainfall()` and `temperature()` with methods [@RCoreTeam] for classes 'data.frame', 'matrix' (or array), and 'sf' [@sf].
+
+say that the idea for the functions started with citizen science projects that is why `day.one` and `span` may be variable across locations.
+For time series analysis where fixed periods are defined across many locations the indices can be adjusted with last.day.
 
@@
 
-The variant B, is calculated using Equation 1, but adjusts $T_{min}$ to $T_{0}$ if $T_{min} < T_{o}$, and is adjusted as follows.
+The variant B, is calculated using Equation 1, but adjusts $T_{min}$ to $T_{0}$ if $T_{min} < T_{0}$, and is adjusted as follows.
 
@@
 
-$$ T_{min} < T_{o} \; \rightarrow \; T_{min} = T_{o} $$
+$$ T_{min} < T_{0} \; \rightarrow \; T_{min} = T_{0} $$
 
@@
 
-The reference evapotranspiration measures the influence of the climate on a given organism (generally a crop species) water needs [@Brouwer1986].
+The reference evapotranspiration measures the influence of the climate on a given organism water needs, generally a crop species [@Brouwer1986].
 The function `ETo()` applies the Blaney-Criddle method, a general theoretical method used when only air-temperature is available locally.
@@
 
+## Seed germination or some GDD related analysis
+
+Use the data from seed germination or crop growth to compute GDD. How many GDD a seed need to become a seedling?
+
 # Further development

### 2020-05-14 R/GDD.R

@@ -246,23 +246,19 @@ GDD.clima_ls <- function(object,
     
     if (isTRUE(equation == "default")) {
       
-      y <- ((x$tmax + x$tmin) / 2) - base
+      y <- .gdd_eq_default(x$tmax, x$tmin, base)
     
     }
     
     if (isTRUE(equation == "variant_a")) {
-      # take the maximum between y and base
-      y <- ((x$tmax + x$tmin) / 2) - base
       
-      y <- max(y, 0)
+      y <- .gdd_eq_variant_a(x$tmax, x$tmin, base)
       
     }
     
     if (isTRUE(equation == "variant_b")) {
-      # set Tmin = base if Tmin < base
-      tadj <- ifelse(x$tmin < base, base, x$tmin)
       
-      y <- ((x$tmax + tadj) / 2) - base
+      y <- .gdd_eq_variant_b(x$tmax, x$tmin, base)
       
     }
     
@@ -286,7 +282,8 @@ GDD.clima_ls <- function(object,
       
       y <- data.frame(id = x$id,
                       date = x$date,
-                      gdd = y)
+                      gdd = y,
+                      stringsAsFactors = FALSE)
       
       return(y)
       
@@ -294,6 +291,7 @@ GDD.clima_ls <- function(object,
   
   })
   )
+  
   result <- do.call("rbind", Y)
   
   if (isTRUE(return.as == "ndays")) {
@@ -309,3 +307,48 @@ GDD.clima_ls <- function(object,
   return(result)
   
 }
+
+.gdd_eq_default <- function(x, y, base) {
+  
+  g <- ((x + y) / 2) - base
+  
+  return(g)
+}
+
+
+.gdd_eq_variant_a <- function(x, y, base) {
+  
+  # take the maximum between y and base
+  g <- ((x + y) / 2) - base
+  
+  g[g < 0] <- 0
+  
+  return(g)
+  
+}
+
+
+.gdd_eq_variant_b <- function(x, y, base){
+  # set Tmin = base if Tmin < base
+  tadj <- y
+  
+  tadj[tadj < base] <- base
+  
+  g <- ((x + tadj) / 2) - base
+  
+  return(g)
+}
+
+
+# // [[Rcpp::export]]
+# NumericVector gdd_variant_aC(NumericVector x, NumericVector y, double base) {
+#   
+#   NumericVector result = clone(x);
+#   
+#   result = ((x + y) / 2) - base;
+#   
[…]

### 2020-05-14 R/anthesis.R

@@ -1,0 +1,393 @@
+#' Crop sensitive indices during anthesis
+#' 
+#' Compute crop sensitive indices. These indices are intended to capture 
+#'  the changes in temperature extremes during anthesis, but can also be 
+#'  applied to other phenological stages.
+#' 
+#' @family temperature functions
+#' @inheritParams temperature
+#' @details 
+#' 
+#' The function uses pre-defined threshold to compute the indices. For hts_mean (32,
+#'  35, 38 Celsius), for hts_max (36, 39, 42 Celsius) and for hse (31 Celsius). 
+#'  See Additional arguments 
+#' 
+#' Additional arguments:
+#' 
+#' The thresholds can be adjusted using the arguments \code{hts_mean.threshold}, 
+#'  \code{hts_max.threshold} and \code{hse.threshold}, which are a numeric (or
+#'  vector of numeric)
+#' 
+#' \code{last.day}: an object (optional to \var{span}) of class \code{Date} or
+#'  any other object that can be coerced to \code{Date} (e.g. integer, character 
+#'  YYYY-MM-DD) for the last day of the time series
+#'  
+#' \code{span}: an integer (optional to \var{last.day}) or a vector with 
+#'  integers (optional if \var{last.day} is given) for the length of 
+#'  the time series to be captured
+#' 
+#' @return A dataframe with crop sensitive indices with n colunms depending on the 
+#'  number of thresholds passed to each index:
+#' \item{hts_mean}{high temperature stress using daily MEAN temperature, 
+#'  and given as percentage number of days a certain threshold is exceeded}
+#' \item{hts_max}{high temperature stress using daily MAX temperature,
+#'  and given as percentage number of days a certain threshold is exceeded}
+#' \item{hse}{heat stress event, and given as percentage number of days a 
+#'  a certain threshold is exceeded for at least two consecutive days}
+#' \item{hse_ms}{heat stress event, and given the maximum number of days 
+#'  a certain threshold is exceeded for at least two consecutive days}
+#' @source 
+#' Challinor et al. (2016). Nature Climate Change 6(10):6954-958
[…]

### 2020-05-14 R/climatrends.R

@@ -3,7 +3,6 @@
 #' @name climatrends
 #' @author Kauê de Sousa and Jacob van Etten and Svein Ø. Solberg
 #' @docType package
-#'
 #' @import methods
 #' @seealso
 #'\strong{Useful links:}

### 2020-05-14 R/crop_sensitive.R

@@ -1,97 +1,0 @@
-#' Crop sensitive indices
-#' 
-#' Compute crop sensitive indices
-#' 
-#' @inheritParams temperature
-#' 
-#' @source 
-#' 
-#' Challinor et al. (2016). Nature Climate Change, 6, 954-958(2016)
-#' \cr\url{https://doi.org/10.1038/nclimate3061}
-#' 
-#' @export
-crop_sensitive <- function(object, ...){
-  return(NA)
-}
-
-#' High temperature stress
-#' 
-#' Calculated using daily mean temperature (or the daily max temperature) 
-#'  and given as the percentage number of days a certain threshold is exceeded 
-#' 
-#' @param x maximum day temperature
-#' @param y optional, minimum temperature
-#' @return the HTS index
-#' @examples 
-#' # hts with day max temperature 
-#' x <- modis[3, ,1]
-#' .hts(x, threshold = 36)
-#' 
-#' # hts with mean daily temperature
-#' x <- modis[1,,1]
-#' y <- modis[1,,2]
-#' .hts(x,y, threshold = 32)
-#' 
-#' # lethal temperatures
-#' .hts(x,y, threshold = 43)
-#' 
-#' @noRd
-.hts <- function(x, y = NULL, threshold = NULL){
-  
-  # if y is provided than take the mean of x and y
-  if (isFALSE(is.null(y))) {
-    x <- (x + y) / 2
-  }
-  
-  # length of x
-  l <- length(x)
-  
-  # number of values above the threshold
-  i <- sum(x > threshold, na.rm = TRUE)
-  
-  # fraction of x above the threshold
-  r <- i / l
-  
-  return(r)
-  
-}
-
-#' Crop duration index
-#' 
-#' Calculated using the mean growing season temperature (or the mean daily
-#'  maximum growing season temperature). If Tmean > threshold then 
-#'  CDI = Tmean - threshold, otherwise CDI = 0
-#' 
-#' @param x maximum day temperature
-#' @param y optional, minimum temperature
-#' @return the HTS index
-#' @examples 
-#' # cdi with mean daily maximum temperature
-#' x <- modis[3, ,1]
-#' .cdi(x, threshold = 36)
-#' 
-#' # cdi with mean temperature
-#' x <- modis[1,,1]
-#' y <- modis[1,,2]
-#' .cdi(x,y, threshold = 32)
-#' 
-#' @noRd
[…]

### 2020-05-14 R/season_sensitive.R

@@ -1,0 +1,55 @@
+#' Seasonal crop sensitive indices
+#' 
+#' Compute seasonal crop sensitive indices. These indices are intended to 
+#'  capture the changes in temperature extremes during a whole season, 
+#'  but can also be applied to specific phenological stages.
+#' 
+#' @inheritParams temperature
+
+
+
+
+#' Crop duration index
+#' 
+#' Calculated using the mean growing season temperature (or the mean daily
+#'  maximum growing season temperature). If Tmean > threshold then 
+#'  CDI = Tmean - threshold, otherwise CDI = 0
+#' 
+#' @param x maximum day temperature
+#' @param y optional, minimum temperature
+#' @return the HTS index
+#' @examples 
+#' # cdi with mean daily maximum temperature
+#' x <- modis[3, ,1]
+#' .cdi(x, threshold = 36)
+#' 
+#' # cdi with mean temperature
+#' x <- modis[1,,1]
+#' y <- modis[1,,2]
+#' .cdi(x,y, threshold = 32)
+#' 
+#' @noRd
+.cdi <- function(x, y = NULL, threshold = NULL){
+  
+  # if y is provided than take the mean of x and y
+  if (isFALSE(is.null(y))) {
+    x <- (x + y) / 2
+  }
+  
+  # take the mean
+  tmean <- mean(x, na.rm = TRUE)
+  
+  # the mean minus the threshold
+  tmean <- tmean - threshold
+  
+  # take the max between tmean and 0
+  r <- max(tmean, 0)
+  
+  return(r)
+  
+}
+
+
+.lsf <- function(x, y = NULL, threshold = -2){
+  
+}

### 2020-05-14 R/temperature.R

@@ -2,10 +2,11 @@
 #'
 #' Methods to compute temperature indices over a time series
 #'
+#' @family temperature functions
 #' @param object a data.frame (or any object that can be coerced to data.frame) 
 #'  with geographical coordinates (lonlat), or an object of class \code{sf}
 #'  with geometry 'POINT' or 'POLYGON', or an \code{array} with two dimensions 
-#'  containing the temperature data or a \code{clima_ls} with day and night 
+#'  containing the temperature data or a \code{clima_ls} with maximum and minimum 
 #'  temperature, in that order. See details 
 #' @inheritParams get_timeseries
 #' @param timeseries logical, \code{FALSE} for a single point time series
@@ -62,8 +63,6 @@
 #' temperature < 10th percentile}
 #' \item{T10p}{the 10th percentile of night tempeture (degree Celsius)}
 #' \item{T90p}{the 90th percentile of day tempeture (degree Celsius)}
-#' 
-#' @family temperature functions
 #' @references 
 #' Aguilar E., et al. (2005). Journal of Geophysical Research, 
 #' 110(D23), D23107. \cr\url{https://doi.org/10.1029/2005JD006119}
@@ -430,7 +429,7 @@ temperature.clima_ls <- function(object,
 #' 
 #' @param x a numeric vector
 #' @return the maximum warm spell duration index
-#' 
+#' @examples
 #' set.seed(871)
 #' x <- rnorm(30, 34)
 #' 
@@ -475,7 +474,7 @@ temperature.clima_ls <- function(object,
 #' 
 #' @param x a numeric vector
 #' @return the maximum cool spell duration index
-#' 
+#' @examples
 #' set.seed(871)
 #' x <- runif(14, 6, 34)
 #' 

### 2020-05-18 inst/paper/paper.Rmd
update-gdd-function

@@
 
-Where $T_{max}$ is the maximum temperature in the given day, $T_{min}$ is the minimum temperature in the given day and $T_{0}$ is the minimum temperature for growth (as per the physiology of the focal organism).
+Where $T_{max}$ is the maximum temperature in the given day, $T_{min}$ is the minimum temperature in the given day and $T_{0}$ is the minimum temperature for growth (as per the physiology of the focal organism or ecosystem averages).
 
@@
 
-Additionally, the function may return the number of days that a given organism required to reach a certain number of accumulated degree-days or the raw gdd in each observed day.
-These values are defined by 'ndays' or 'gdd' and can be adjusted using the argument `return.as`.
+By default, the function returns the degree-days that is accumulated over the time series.
+Additionally, the function may return the daily values of degree-days or the number of days that a given organism required to reach a certain number of accumulated degree-days.
+These values are defined by 'acc', 'daily' or 'ndays' and can be adjusted using the argument `return.as`.
 The required accumulated gdd is defined with argument `degree.days`.
@@
 
-$$\parallel GDD_{ndays} \parallel \: = \; \sum^{ggd_1}_{gdd_n}$$
+$$\parallel GDD_{ndays} \parallel \: = \; ggd_1 \;+ \; ...  \; +  \; gdd_n$$
 

### 2020-05-18 R/GDD.R

@@ -14,13 +14,14 @@
 #' @details 
 #' Additional arguments:
 #' 
-#' \code{return.as} character (one of, the default, \code{"gdd"} or \code{"ndays"})
-#'  to select if the function returns the number of days to reach the accumulated 
-#'  \var{degree.days} or the daily values of the GDD
+#' \code{return.as} character (one of, the default, \code{"acc"} or \code{"daily"},
+#'  \code{"ndays"}) to select if the function returns the accumulated gdd, or the 
+#'  daily values of gdd across the series of the number of days required to reach a
+#'  certain number of \code{degree.days}
 #'  
 #' \code{degree.days} an integer for the accumulated degree-days required by the 
 #'  organism (as per the physiology of the focal organism). Optional if 
-#'  \var{return.as} = \code{"gdd"}
+#'  \var{return.as} = \code{"gdd"} or \var{return.as} = \code{"acc"} 
 #' 
 #' \code{equation} character to specify the equation to be used, one of \code{"default"},
 #' \code{"variant_a"} or \code{"variant_b"}. See Equations below 
@@ -102,7 +103,7 @@ GDD.default <- function(tmax, tmin, dates = NULL, ..., base = 10) {
   degree.days <- dots[["degree.days"]]
   
   if (is.null(return.as)) {
-    return.as <- "gdd" 
+    return.as <- "acc" 
   }
   
   if (is.null(equation)) {
@@ -148,7 +149,7 @@ GDD.data.frame <- function(object, day.one, ..., base = 10){
   degree.days <- dots[["degree.days"]]
   
   if (is.null(return.as)) {
-    return.as <- "gdd" 
+    return.as <- "acc" 
   }
   
   if (is.null(equation)) {
@@ -195,7 +196,7 @@ GDD.array <- function(object, day.one, ..., base = 10){
   degree.days <- dots[["degree.days"]]
   
   if (is.null(return.as)) {
-    return.as <- "gdd" 
+    return.as <- "acc" 
   }
   
   if (is.null(equation)) {
@@ -239,7 +240,7 @@ GDD.sf <- function(object, day.one, ..., base = 10, as.sf = TRUE){
   degree.days <- dots[["degree.days"]]
   
   if (is.null(return.as)) {
-    return.as <- "gdd" 
+    return.as <- "acc" 
   }
   
   if (is.null(equation)) {
[…]

### 2020-05-18 R/late_frost.R

@@ -247,7 +247,7 @@ late_frost.clima_ls <- function(object, ..., base = 4, tfrost = -2){
     g <- .gdd(temp,
               base = base,
               equation = equation,
-              return.as = "gdd")$gdd
+              return.as = "daily")$gdd
 
     # get a vector where 1 are the frost events and 0 are the non-frost
     frost <- as.integer(tmin <= tfrost & g == 0)

### 2020-05-19 R/ETo.R
add-new-eq-for-GDD

@@ -87,7 +87,7 @@
 #' }
 #' @importFrom sf st_bind_cols
 #' @export
-ETo <- function(..., Kc = 1){
+ETo <- function(object, ..., Kc = 1){
   
   UseMethod("ETo")
 
@@ -96,7 +96,7 @@ ETo <- function(..., Kc = 1){
 #' @rdname ETo
 #' @method ETo default
 #' @export
-ETo.default <- function(tmax, tmin, ..., Kc = 1, lat = NULL, month = NULL){
+ETo.default <- function(object, tmin, ..., Kc = 1, lat = NULL, month = NULL){
   
   dots <- list(...)
   p <- dots[["p"]]
@@ -111,11 +111,12 @@ ETo.default <- function(tmax, tmin, ..., Kc = 1, lat = NULL, month = NULL){
     p <- 0.27
   }
   
-  tmax <- data.frame(id = 1, value = tmax, stringsAsFactors = FALSE)
+  temp <- data.frame(id = 1, 
+                     tmax = object, 
+                     tmin = tmin,
+                     stringsAsFactors = FALSE)
   
-  tmin <- data.frame(id = 1, value = tmin, stringsAsFactors = FALSE)
-  
-  result <- .eto(tmax, tmin, Kc, p)
+  result <- .eto(temp, Kc, p)
   
   return(result)
   
@@ -159,11 +160,11 @@ ETo.data.frame <- function(object, day.one, ..., Kc = 1){
   
   dat <- get_timeseries(object, day.one, pars = pars, ...)
   
-  day <- dat[[pars[[1]]]]
+  temp <- cbind(dat[[pars[[1]]]], tmin = dat[[pars[[2]]]]$value)
   
-  night <- dat[[pars[[2]]]]
+  names(temp)[names(temp) == "value"] <- "tmax"
   
-  result <- .eto(day, night, Kc, p)
+  result <- .eto(temp, Kc, p)
   
   return(result)
 }
@@ -198,9 +199,13 @@ ETo.array <- function(object, day.one, ..., Kc = 1, lat = NULL, p = 0.27){
     p <- .p_daytime(lat = lat, month = m)
   }
   
-  ts <- get_timeseries(object, day.one, span, last.day)
+  dat <- get_timeseries(object, day.one, span, last.day)
   
-  result <- .eto(ts[[1]], ts[[2]], Kc, p)
+  temp <- cbind(dat[[1]], tmin = dat[[2]]$value)
+  
[…]

### 2020-05-19 R/GDD.R

@@ -7,12 +7,15 @@
 #' @family temperature functions
 #' @family GDD functions
 #' @inheritParams temperature
-#' @param base an integer for the minimum temperature for growth
+#' @param tbase an integer for the minimum temperature for growth
 #' @return 
 #'  The number of days to reach the accumulated \var{degree.days} or the daily degree-days 
 #'   as defined with the argument \var{return.as}
 #' @details 
 #' Additional arguments:
+#' 
+#' \code{tbase_max} optional, the maximum tbase temperature, 
+#'  required if \code{equation = "c"}   
 #' 
 #' \code{return.as} character (one of, the default, \code{"acc"} or \code{"daily"},
 #'  \code{"ndays"}) to select if the function returns the accumulated gdd, or the 
@@ -20,23 +23,23 @@
 #'  certain number of \code{degree.days}
 #'  
 #' \code{degree.days} an integer for the accumulated degree-days required by the 
-#'  organism (as per the physiology of the focal organism). Optional if 
-#'  \var{return.as} = \code{"gdd"} or \var{return.as} = \code{"acc"} 
+#'  organism. Optional if \var{return.as = "daily"} or \var{return.as = "acc"} 
 #' 
 #' \code{equation} character to specify the equation to be used, one of \code{"default"},
-#' \code{"variant_a"} or \code{"variant_b"}. See Equations below 
+#' \code{"a"}, \code{"b"} or \code{"c"}. See Equations below 
+#' 
+#' \code{dates}: a character (or Date or numeric) vector for the dates of tmax and tmin
+#'  in the \code{default} method
 #' 
 #' \code{last.day}: an object (optional to \var{span}) of class \code{Date} or
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
-#'  YYYY-MM-DD) for the last day of the time series
+#'  YYYY-MM-DD) for the last day of the time series. For \code{data.frame}, 
+#'  \code{array} and \code{sf} methods
 #'  
 #' \code{span}: an integer (optional to \var{last.day}) or a vector with 
 #'  integers (optional if \var{last.day} is given) for the length of 
-#'  the time series to be captured
-#'  
[…]

### 2020-05-19 R/clima_data.R

@@ -19,8 +19,8 @@
 #' \code{innlandet}: matrix with maximum and minimum temperature for a random point in 
 #'  the Innlandet county in Norway, spanning from "2019-01-01" to "2019-07-01"
 #'  
-#' \code{lonlatsf}: a 'sf' object with 'POINT' geometry with five random points across 
-#'  the Innlandet county in Norway 
+#' \code{lonlatsf}: a 'sf' object with 'POINT' geometry with five random points around 
+#'  Sinop, Brazil 
 #'  
 #' In \code{temp_dat} and \code{rain_dat}, rows represents the coordinates for the 
 #'  given lonlat provided in \code{raster::extract()} and colunms represents the dates 

### 2020-05-19 R/crop_sensitive.R

@@ -22,11 +22,13 @@
 #' 
 #' \code{last.day}: an object (optional to \var{span}) of class \code{Date} or
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
-#'  YYYY-MM-DD) for the last day of the time series
+#'  YYYY-MM-DD) for the last day of the time series. For \code{data.frame}, \code{array} 
+#'  and \code{sf} methods
 #'  
 #' \code{span}: an integer (optional to \var{last.day}) or a vector with 
 #'  integers (optional if \var{last.day} is given) for the length of 
-#'  the time series to be captured
+#'  the time series to be captured. For \code{data.frame}, \code{array} 
+#'  and \code{sf} methods
 #' 
 #' @return A dataframe with crop sensitive indices with n colunms depending on the 
 #'  number of thresholds passed to each index:
@@ -77,14 +79,14 @@
 #'                hts_mean.threshold = c(24),
 #'                hts_max.threshold = c(31, 33))
 #' @export
-crop_sensitive <- function(...){
+crop_sensitive <- function(object, ...){
   UseMethod("crop_sensitive")
 }
 
 #' @rdname crop_sensitive
 #' @method crop_sensitive default
 #' @export
-crop_sensitive.default <- function(tmax, tmin, ...) {
+crop_sensitive.default <- function(object, tmin, ...) {
   
   dots <- list(...)
   hts_mean <- dots[["hts_mean.threshold"]]
@@ -94,7 +96,10 @@ crop_sensitive.default <- function(tmax, tmin, ...) {
   cdi_max <- dots[["cdi_max.threshold"]]
   lethal <- dots[["lethal.threshold"]]
   
-  temp <- data.frame(id = 1, tmax = tmax, tmin = tmin, stringsAsFactors = FALSE)
+  temp <- data.frame(id = 1, 
+                     tmax = object, 
+                     tmin = tmin,
+                     stringsAsFactors = FALSE)
   
   result <- .crop_sensitive(temp = temp, 
                             hts_mean.threshold = hts_mean,
@@ -151,7 +156,6 @@ crop_sensitive.data.frame <- function(object, day.one, ...) {
   
 }
 
-
 #' @rdname crop_sensitive
 #' @method crop_sensitive array
 #' @export
@@ -173,35 +177,6 @@ crop_sensitive.array <- function(object, day.one, ...){
[…]

### 2020-05-19 R/late_frost.R

@@ -10,8 +10,11 @@
 #' @details 
 #' Additional arguments:
 #' 
-#' \code{equation}: character to specify the equation to be used, \code{"variant_a"} 
+#' \code{equation}: character to specify the equation to be used, \code{"b"} 
 #'  is set by default. See GDD() 
+#' 
+#' \code{dates}: a character (or Date or numeric) vector for the dates of tmax and tmin
+#'  in the \code{default} method
 #' 
 #' \code{last.day}: an object (optional to \var{span}) of class \code{Date} or
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
@@ -37,15 +40,24 @@
 #' \cr\url{https://doi.org/10.1073/pnas.1920816117}
 #' 
 #' @examples 
-#' 
 #' # default method
 #' data("innlandet", package = "climatrends")
 #' 
-#' tmax <- innlandet[, "tmax"]
-#' tmin <- innlandet[, "tmin"]
-#' date <- innlandet[, "dates"]
+#' # equation b is set by default
+#' # where tmin and tmax are adjusted if below tbase
+#' late_frost(innlandet$tmax, 
+#'            innlandet$tmin, 
+#'            dates = innlandet$date, 
+#'            base = 2)
 #' 
-#' late_frost(tmax, tmin, dates = date)
+#' # slightly different series if equation a is used
+#' late_frost(innlandet$tmax, 
+#'            innlandet$tmin, 
+#'            dates = innlandet$date, 
+#'            base = 2,
+#'            equation = "a")
+#' 
+#' #####################################################
 #' 
 #' # demo of the array method but no frost event is returned 
 #' # because the data comes from the tropics
@@ -53,6 +65,7 @@
 #' 
 #' late_frost(temp_dat, day.one = "2013-10-27")
 #' 
+#' #####################################################
 #' 
 #' \donttest{
 #' # Some random points in Norway
@@ -65,7 +78,7 @@
 #' }
 #' 
 #' @export
-late_frost <- function(..., base = 4, tfrost = -2) {
+late_frost <- function(object, ..., tbase = 4, tfrost = -2) {
   
   UseMethod("late_frost")
 
@@ -75,13 +88,14 @@ late_frost <- function(..., base = 4, tfrost = -2) {
 #' @rdname late_frost
 #' @method late_frost default
 #' @export
[…]

### 2020-05-19 R/rainfall.R

@@ -10,28 +10,19 @@
 #'  precipitation data. See details.   
 #' @param timeseries logical, \code{FALSE} for a single point time series
 #'  observation or \code{TRUE} for a time series based on \var{intervals}
-#' @param intervals integer (no lower than 5), for the days intervals when
-#'  \var{timeseries} = \code{TRUE}
-#' @param dates optional, a character (or Date or numeric) vector for the dates in 
-#'  \var{object} in the dafault method
 #' @param as.sf logical, to return an object of class 'sf'
 #' @details 
-#' The \code{matrix} method assumes that \var{object} contains climate data available in your 
-#'  R section; see help("rain_dat", package = "climatrends") for an example on input 
-#'  structure.
+#' #' Additional arguments:
 #' 
-#' The \code{default} method and the \code{sf} method assumes that the climate data
-#'  will e fetched from a remote (cloud) source that be adjusted using the argument 
-#'  \var{data.from}.
-#'
-#' When \var{timeseries} = \code{TRUE}, an id is created, 
-#'  which is the index for the rownames of the inputted \var{object}.
-#' 
-#' Additional arguments:
-#' 
+#' \code{intervals}: an integer (no lower than 5), for the days intervals when
+#'  \var{timeseries = TRUE}
+#'  
 #' \code{last.day}: optional to \var{span}, an object of class \code{Date} or
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
 #'  YYYY-MM-DD)  for the last day of the time series
+#'  
+#' \code{dates}: a character (or Date or numeric) vector for the dates of tmax and tmin
+#'  in the \code{default} method
 #' 
 #' \code{data.from}: character for the source of remote data. Current remote source 
 #'  is: 'nasapower'
@@ -42,6 +33,16 @@
 #' \code{days.before}: optional, an integer for the number of days before 
 #'  \var{day.one} to be included in the timespan.
 #'  
+#' # S3 Methods 
+#' 
+#' The \code{matrix} method assumes that \var{object} contains climate data available in
+#'  your R section; see help("rain_dat", package = "climatrends") for an example on input 
+#'  structure.
+#' 
+#' The \code{data.frame} and the \code{sf} methods assumes that the climate data
[…]

### 2020-05-19 R/temperature.R

@@ -3,40 +3,31 @@
 #' Methods to compute temperature indices over a time series
 #'
 #' @family temperature functions
-#' @param object a data.frame with geographical coordinates (lonlat), 
+#' @param object a numeric vector with the maximum temperature, 
+#'  or a data.frame with geographical coordinates (lonlat), 
 #'  or an object of class \code{sf} with geometry 'POINT' or 'POLYGON', 
-#'  or an \code{array} with two dimensions containing the temperature data 
-#'  or a \code{clima_ls} with maximum and minimum temperature, 
-#'  in that order. See details 
+#'  or an \code{array} with two dimensions containing the
+#'  maximum and minimum temperature, in that order. See details 
 #' @inheritParams get_timeseries
 #' @param timeseries logical, \code{FALSE} for a single point time series
 #'  observation or \code{TRUE} for a time series based on \var{intervals}
-#' @param intervals integer (no lower than 5), for the days intervals when
-#'  \var{timeseries} = \code{TRUE}
-#' @param dates optional, a character (or Date or numeric) vector for the dates in 
-#'  \var{tmax} and \var{tmin}
-#' @param tmax a numeric vector with the maximum temperature
 #' @param tmin a numeric vector with the minimum temperature
 #' @param as.sf logical, to return an object of class 'sf'
 #' @param ... additional arguments passed to methods. See details
 #' @details 
-#' The \code{array} method assumes that \var{object} contains climate data available 
-#'  in your R section; this requires an array with two dimensions, 1st dimension 
-#'  contains the day temperature and 2nd dimension the night temperature, 
-#'  see help("temp_dat", package = "climatrends") for an example on input structure.
-#' 
-#' The \code{data.frame} method and the \code{sf} method assumes that the climate data
-#'  will e fetched from a remote (cloud) source that be adjusted using the argument 
-#'  \var{data.from}.
-#'
-#' When \var{timeseries} = \code{TRUE}, an id is created, 
-#'  which is the index for the rownames of the inputted \var{object}.
 #' 
 #' Additional arguments:
 #' 
+#' \code{intervals}: an integer (no lower than 5), for the days intervals when
+#'  \var{timeseries = TRUE}
+#' 
[…]

### 2020-05-19 tests/testthat/test-GDD.R

@@ -4,13 +4,13 @@ library("sf")
 # load("tests/test_data.rda")
 load("../test_data.rda")
 
-realv <- c(5,4,5,11,5,7,5,6,6,8)
+realv <- c(4,3,4,10,4,6,4,5,5,7)
 test_that("right values array method", {
   
   dg <- GDD(object = temp,
             day.one = d, 
             degree.days = 45,
-            base = 10,
+            tbase = 10,
             return.as = "ndays")
   
   dg <- all(realv == dg[, 1])
@@ -24,11 +24,11 @@ realv <- c(11, 11, 11, 11, 11,
 # return daily gdd
 test_that("return gdd", {
   
-  gdd <- GDD(object = temp[1:2,,],
+  gdd <- GDD(object = temp[1:2, , ],
              day.one = d[1:2,], 
              span = 5,
-             base = 10,
-             return.as = "gdd")$gdd
+             tbase = 10,
+             return.as = "daily")$gdd
   
   gdd <- round(gdd, 0)
   
@@ -49,7 +49,7 @@ test_that("nasapower works", {
             day.one = "2013-10-28",
             last.day = "2013-11-27",
             degree.days = 30, 
-            base = 5,
+            tbase = 5,
             return.as = "ndays")
   
   dg <- all(dg$GDD == realv)
@@ -64,7 +64,7 @@ test_that("missing degree.days return.as ndays", {
   expect_error(
     GDD(object = temp,
         day.one = d, 
-        base = 10,
+        tbase = 10,
         return.as = "ndays")
   )
   

### 2020-05-19 vignettes/precompile.R

@@ -1,22 +1,0 @@
-# vignettes that depend on internet access need to be precompiled and take a
-# while to run
-library("knitr")
-knit("vignettes/Overview.Rmd.orig", "vignettes/Overview.Rmd")
-
-# remove file path such that vignettes will build with figures
-replace <- readLines("vignettes/Overview.Rmd")
-replace <- gsub("<img src=\"vignettes/", "<img src=\"", replace)
-fileConn <- file("vignettes/Overview.Rmd")
-writeLines(replace, fileConn)
-close(fileConn)
-
-
-library("devtools")
-build_vignettes()
- 
-# move resource files to /doc
-resources <-
-  list.files("vignettes/", pattern = ".png$", full.names = TRUE)
-file.copy(from = resources,
-          to = "docs",
-          overwrite =  TRUE)

### 2020-05-19 R/late_frost.R
late_frost_examples

@@ -71,10 +71,11 @@
 #' # Some random points in Norway
 #' # get data from NASA Power
 #' library("nasapower")
-#' data("lonlatsf", package = "climatrends")
-#'   
-#' late_frost(lonlatsf, day.one = "2019-01-01", last.day = "2019-07-01")
-#'   
+#' lonlat <- data.frame(lon = c(10.93, 10.57),
+#'                      lat = c(60.77, 61.10))
+#' 
+#' late_frost(lonlat, day.one = "2019-01-01", last.day = "2019-07-01")
+#'     
 #' }
 #' 
 #' @export

### 2020-05-19 vignettes/Overview.R

@@ -1,0 +1,73 @@
+## ----temperature, message=TRUE, eval=TRUE, echo=TRUE--------------------------
+library("climatrends")
+
+data("innlandet", package = "climatrends")
+
+temp1 <- temperature(innlandet$tmax, innlandet$tmin)
+
+temp1
+
+
+## ----temperature2, message=TRUE, eval=TRUE, echo=TRUE-------------------------
+temp2 <- temperature(innlandet$tmax, innlandet$tmin,
+                     dates = innlandet$dates,
+                     timeseries = TRUE, 
+                     intervals = 30)
+
+temp2
+
+## ----gdd, message=TRUE, eval=TRUE, echo=TRUE----------------------------------
+gdd <- GDD(innlandet$tmax, innlandet$tmin, tbase = 2, equation = "b")
+
+gdd
+
+## ----gdd2, message=FALSE, eval=FALSE, echo=TRUE-------------------------------
+#  lonlat <- data.frame(lon = 129.19,
+#                       lat = 36.39)
+#  
+#  GDD(lonlat,
+#      day.one = "2019-04-01",
+#      last.day = "2019-10-01",
+#      degree.days = 150,
+#      return.as = "ndays")
+
+## ----latefrost, message=TRUE, eval=TRUE, echo=TRUE----------------------------
+lf <- late_frost(innlandet$tmax, 
+                 innlandet$tmin, 
+                 dates = innlandet$date, 
+                 base = 2)
+
+lf
+
+## ----rain, message=FALSE, eval=FALSE, echo=TRUE-------------------------------
+#  library("nasapower")
+#  
+#  lonlat <- data.frame(lon = c(-73.3, -74.5),
+#                       lat = c(-6.1, - 6.2))
+#  
+#  rain <- rainfall(lonlat,
+#                   day.one = "2018-11-01",
+#                   last.day = "2018-12-31")
+#  
+#  rain
+
+## ----csenstive, message=FALSE, eval=FALSE, echo=TRUE--------------------------
+#  data("lonlatsf", package = "climatrends")
+#  
+#  crop_sensitive(lonlatsf,
+#                 day.one = "2018-12-01",
+#                 last.day = "2019-01-31",
+#                 as.sf = FALSE)
+
+## ----eto, message=TRUE, eval=TRUE, echo=TRUE----------------------------------
+data("temp_dat", package = "climatrends")
+
+eto <- ETo(temp_dat, 
[…]

### 2020-05-20 inst/paper/paper.Rmd
fix-gdd-bug

@@
 \
+```{r get data, include=FALSE}
+library("knitr")
+library("kableExtra")
+library("magrittr")
+funtbl <- read.csv("available_functions.csv")
+```
 
@@
 
-Four main functions are provided, `ETo()`, `GDD()`, `rainfall()` and `temperature()` with methods [@RCoreTeam] for classes 'data.frame', 'matrix' (or array), and 'sf' [@sf].
+Six main functions are provided, `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector's and additional methods [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
+The later two are designed to fetch data from cloud sources.
 
@@
 [...continue...]
+
+```{r characteristics, echo=FALSE, message=TRUE}
+kable(funtbl, 
+      row.names = FALSE,
+      align = "l", 
+      caption = "Main functions available in climatrends.") %>% 
+  kable_styling(latex_options = "scale_down") %>% 
+  column_spec(2, width = "1in") %>% 
+  column_spec(3, width = "1.7in") %>% 
+  column_spec(4, width = "1.7in")
+```
 
@@
 
-For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two variants of the previous equation to adjust either $T_{mean}$ (variant A) or $T_{min}$ (variant B).
-The variant A changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.
+Additionally, the function `GDD()` offers three modified equations (two) designed for cold environments and (one) for tropical environments.
+For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two variants of the previous equation to adjust either $T_{mean}$ (variant a) or $T_{min}$ (variant b).
+The variant a changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.
 
@@
 
-The variant B, is calculated using Equation 1, but adjusts $T_{min}$ to $T_{0}$ if $T_{min} < T_{0}$, and is adjusted as follows.
+The variant b, is calculated using Equation 1, but adjusts $T_{min}$ or $T_{max}$ to $T_{0}$ if $T < T_{0}$, the equation is adjusted as follows.
 
@@
 
-$$ T_{min} < T_{0} \; \rightarrow \; T_{min} = T_{0} $$
+$$ T < T_{0} \; \rightarrow \; T = T_{0} $$
 
-In `GDD()` these equations are defined as 'default', 'variant_a' and 'variant_b', respectively, and can be selected using the argument `equation`.
+Where $T$ may refer to $T_{min}$ and/or $T_{max}$ when the condition of being below $T_{0}$ applies.
 
-By default, the function returns the degree-days that is accumulated over the time series.
+For tropical areas, where the temperature may surpass a maximum threshold ($T_{0^{max}}$), resulting in limited development, the minimum temperature is adjusted using Equation 3 and the maximum temperature is adjusted to a maximum base temperature as follow.
+
+Equation [4]
+
+$$ T_{max} > T_{0^{max}} \; \rightarrow \; T_{max} = T_{0^{max}} $$
+
+Where $T_{0^{max}}$ is the maximum base temperature for growth, defined in `GDD()` using the argument `tbase_max`.
+
+These modified equations are defined as 'a', 'b' and 'c', respectively, and can be selected using the argument `equation`.
+
+By default, the function returns the degree-days that is accumulated over the time series using Equation 1.
 Additionally, the function may return the daily values of degree-days or the number of days that a given organism required to reach a certain number of accumulated degree-days.
@@
 The required accumulated gdd is defined with argument `degree.days`.
-For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset of photosynthesis [@JWu2013].
-In that case, the function will calculate the growing degree-days ($gdd$) and sum up the values until it reaches the defined gdd (105 $^\circ C$) and return the number of days needed in the given season ($GDD_{ndays}$), as follows.
+For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset the photosynthesis [@JWu2013].
+In that case, `GDD()` will calculate the growing degree-days ($gdd$) and sum up the values until it reaches the defined gdd (105 $^\circ C$) and return the number of days needed in the given season ($GDD_{ndays}$), as follows.
 
-Equation [4]
+Equation [5]
 
@@
 
-Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the factor for organism water need may change according to the growth stage.
+Where $p$ is the mean daily percentage of annual daytime hours, $T_{max}$ is the maximum temperature, $T_{min}$ is the minimum temperature, and $K_c$ is the factor for organism water need.
 

### 2020-05-20 R/GDD.R

@@ -472,7 +472,7 @@ GDD.sf <- function(object, day.one, ..., tbase = 10, as.sf = TRUE){
 
   tmax_adj[tmax_adj > tbase_max] <- tbase_max
   
-  tmin_adj[tmin_adj > tbase] <- tbase
+  tmin_adj[tmin_adj < tbase] <- tbase
   
   g <- ((tmax_adj + tmin_adj) / 2) - tbase
 

### 2020-05-20 R/crop_sensitive.R

@@ -1,6 +1,6 @@
 #' Crop sensitive indices
 #' 
-#' Compute crop sensitive indices. These indices are intended to capture 
+#' Compute crop sensitive indices. These indices are designed to capture 
 #'  the changes in temperature extremes during key phenological stages 
 #'  (e.g. anthesis), but can also be applied to other phenological stages.
 #' 

### 2020-05-20 R/late_frost.R

@@ -24,7 +24,7 @@
 #'  integers (optional if \var{last.day} is given) for the length of 
 #'  the time series to be captured
 #'  
-#' @return a data.frame with the late frost events
+#' @return A data.frame with the late frost events
 #' \item{id}{the id generated using the indices for the rows in \var{object}}
 #' \item{date}{the first day of the event}
 #' \item{gdd}{the growing degree-days accumulated during the event}

### 2020-05-22 inst/paper/paper.Rmd
speed-gdd-late_frost

@@
 ---
-title: 'climatrends: Precipitation and temperature indices for climate variability analysis in R'
+title: 'Climate variability indices for ecological and crop models in R: the `climatrends` package'
 tags:
@@
 
+```{r get data, include=FALSE}
+library("knitr")
+library("kableExtra")
+library("magrittr")
+funtbl <- read.csv("available_functions.csv")
+```
+
 >Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
@@
 \
-```{r get data, include=FALSE}
-library("knitr")
-library("kableExtra")
-library("magrittr")
-funtbl <- read.csv("available_functions.csv")
-```
 
@@
 
-Six main functions are provided, `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector's and additional methods [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
+Six main functions are provided, `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector's and additional methods implemented via the package `methods` [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
 The later two are designed to fetch data from cloud sources.
@@
 
+## Temperature and precipitation indices
+
 ## Growing degree-days
@@
 
-Additionally, the function `GDD()` offers three modified equations (two) designed for cold environments and (one) for tropical environments.
-For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two variants of the previous equation to adjust either $T_{mean}$ (variant a) or $T_{min}$ (variant b).
+Additionally, the function `GDD()` offers three modified equations designed for cold environments and for tropical environments.
+For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two modified equations that adjusts either $T_{mean}$ (variant a) or $T_{min}$ (variant b).
 The variant a changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.
@@
 For example, the Korean pine (*Pinus koraiensis*) requires 105 $^\circ C$ accumulated gdd to onset the photosynthesis [@JWu2013].
-In that case, `GDD()` will calculate the growing degree-days ($gdd$) and sum up the values until it reaches the defined gdd (105 $^\circ C$) and return the number of days needed in the given season ($GDD_{ndays}$), as follows.
+In that case, `GDD()` will calculate the growing degree-days ($gdd$) and sum up the values until it reaches 105 $^\circ C$ and return the number of days required in the given season ($GDD_{r}$), as follows.
 
@@
 
-$$\parallel GDD_{ndays} \parallel \: = \; ggd_1 \;+ \; ...  \; +  \; gdd_n$$
+$$\parallel GDD_{r} \parallel \: = \; ggd_1 \;+ \; ...  \; +  \; gdd_n$$
 
-## Reference evapotranspiration
+## Late-spring frost
+
+## Crop-related indices
+
+Two functions in **climatrends** are mainly designed to capture the effects of climate on the development and stress of crop species, `crop_sensitive` computes indices that aims to capture the changes in temperature extremes during key phenological stages (e.g.
+anthesis), and `ETo()` computes the reference evapotranspiration.
 
@@
 
-Equation [5]
+Equation [6]
 
@@
 
-The percentage of daytime hours ($p$) is calculated internally by most methods in `ETo()` (except the array method) using the given latitude (taken from the inputted `object`) and date (taken from the inputted `day.one`).
+The percentage of daytime hours ($p$) is calculated internally by the 'data.frame' and 'sf' methods in `ETo()` using the given latitude (taken from the inputted `object`) and date (taken from the inputted `day.one`).
 It matches the latitude and date with a table of daylight percentage derived from Brouwer and Heibloem [-@Brouwer1986].
@@
 
+# Authors' contributions
+
+KdS and JvE conceived the ideas of implementing an `R` package. JvE and SØS assisted in the implementation of methods. KdS wrote and documented the software. [add other contributions...]
+
+# Data availability statement
+
+The source code has been archived at [**add link**] as `climatrends` version **X.Y.Z**.
+To explore the latest functionalities of `climatrends`, please check the package's updates at CRAN (https://cran.r-project.org/package=climatrends).
+
 # References

### 2020-05-22 R/GDD.R

@@ -14,22 +14,19 @@
 #' @details 
 #' Additional arguments:
 #' 
+#' \code{equation} character to specify the equation to be used, one of \code{"default"},
+#' \code{"a"}, \code{"b"} or \code{"c"}. See Equations below 
+#' 
 #' \code{tbase_max} optional, the maximum tbase temperature, 
 #'  required if \code{equation = "c"}   
 #' 
 #' \code{return.as} character (one of, the default, \code{"acc"} or \code{"daily"},
 #'  \code{"ndays"}) to select if the function returns the accumulated gdd, or the 
-#'  daily values of gdd across the series of the number of days required to reach a
+#'  daily values of gdd across the series, or the number of days required to reach a
 #'  certain number of \code{degree.days}
 #'  
 #' \code{degree.days} an integer for the accumulated degree-days required by the 
 #'  organism. Optional if \var{return.as = "daily"} or \var{return.as = "acc"} 
-#' 
-#' \code{equation} character to specify the equation to be used, one of \code{"default"},
-#' \code{"a"}, \code{"b"} or \code{"c"}. See Equations below 
-#' 
-#' \code{dates}: a character (or Date or numeric) vector for the dates of tmax and tmin
-#'  in the \code{default} method
 #' 
 #' \code{last.day}: an object (optional to \var{span}) of class \code{Date} or
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
@@ -57,7 +54,7 @@
 #'  
 #'  \code{"default"}: GDD = ((tmax + tmin) / 2) - tbase
 #'  
-#'  \code{"a"}: GDD = max(GDD, 0)
+#'  \code{"a"}: adjust tmean = tbase if tmeam < tbase
 #'  
 #'  \code{"b"}: adjust tmin = tbase if tmin < tbase,
 #'   adjust tmax = tbase if tmax < tbase
@@ -75,7 +72,7 @@
 #' @examples
 #' data("innlandet", package = "climatrends")
 #' 
-#' # use the default equation, but in this case it returns negative values
+#' # use the default equation
 #' GDD(innlandet$tmax, innlandet$tmin, tbase = 2)
 #' 
 #' # set the equation "b", which is a better option for this case
@@ -127,7 +124,6 @@ GDD.default <- function(object, tmin, ..., tbase = 10) {
   
   dots <- list(...)
   equation <- dots[["equation"]]
[…]

### 2020-05-22 R/crop_sensitive.R

@@ -227,7 +227,7 @@ crop_sensitive.sf <- function(object, day.one, ..., as.sf = TRUE){
   
   if (isTRUE(as.sf)) {
     
-    result <- suppressWarnings(sf::st_bind_cols(object, result))
+    result <- cbind(object, result)
     
   }
   

### 2020-05-22 R/late_frost.R

@@ -48,13 +48,15 @@
 #' late_frost(innlandet$tmax, 
 #'            innlandet$tmin, 
 #'            dates = innlandet$date, 
-#'            base = 2)
+#'            tbase = 2, 
+#'            tfrost = -2)
 #' 
 #' # slightly different series if equation a is used
 #' late_frost(innlandet$tmax, 
 #'            innlandet$tmin, 
 #'            dates = innlandet$date, 
-#'            base = 2,
+#'            tbase = 2,
+#'            tfrost = -2,
 #'            equation = "a")
 #' 
 #' #####################################################
@@ -71,11 +73,11 @@
 #' # Some random points in Norway
 #' # get data from NASA Power
 #' library("nasapower")
-#' lonlat <- data.frame(lon = c(10.93, 10.57),
-#'                      lat = c(60.77, 61.10))
+#' lonlat <- data.frame(lon = c(10.93, 10.57, 11.21),
+#'                      lat = c(60.77, 61.10, 60.33))
 #' 
 #' late_frost(lonlat, day.one = "2019-01-01", last.day = "2019-07-01")
-#'     
+#'      
 #' }
 #' 
 #' @export
@@ -107,13 +109,7 @@ late_frost.default <- function(object, tmin, ..., tbase = 4, tfrost = -2) {
     dates <- rep(NA, times = length(object))
   }
   
-  temp <- data.frame(id = 1, 
-                     tmax = object, 
-                     tmin = tmin, 
-                     date = dates, 
-                     stringsAsFactors = FALSE)
-  
-  result <- .late_frost(temp, tbase, tfrost, equation)
+  result <- latefrost(object, tmin, dates, tbase, tfrost, equation)
   
   return(result)
   
@@ -152,7 +148,19 @@ late_frost.data.frame <- function(object, day.one, ..., tbase = 4, tfrost = -2){
   
   names(temp)[names(temp)=="value"] <- "tmax"
   
-  result <- .late_frost(temp, tbase, tfrost, equation)
+  temp <- split(temp, temp$id)
+  
+  result <- lapply(temp, function(x){
+    lf <- latefrost(x$tmax, x$tmin, dates = x$date, tbase, tfrost, equation)
+    i <- rep(x$id[1], times = dim(lf)[[1]])
+    lf <- cbind(id = i, lf)
+  })
+  
+  result <- do.call("rbind", result)
+  
+  row.names(result) <- seq_len(dim(result)[[1]])
+  
[…]

### 2020-05-22 R/ETo.R
rm-sf-from-dependencies

@@ -78,6 +78,7 @@
 #' \donttest{
 #' #######################################
 #' library("nasapower")
+#' library("sf")
 #' data("lonlatsf", package = "climatrends")
 #' 
 #' ETo(lonlatsf,
@@ -85,7 +86,6 @@
 #'     last.day = "2015-06-22",
 #'     pars = c("T10M_MAX", "T10M_MIN"))
 #' }
-#' @importFrom sf st_bind_cols
 #' @export
 ETo <- function(object, ..., Kc = 1){
   
@@ -250,7 +250,7 @@ ETo.sf <- function(object, day.one, ..., Kc = 1, as.sf = TRUE){
   result <- .eto(temp, Kc, p)
   
   if (isTRUE(as.sf)) {
-    result <- suppressWarnings(sf::st_bind_cols(object, result))
+    result <- cbind(object, result)
   }
   
   return(result)

### 2020-05-22 R/get_timeseries.R

@@ -56,7 +56,7 @@
 #' ########################################################
 #' \donttest{
 #' library("nasapower")
-#' 
+#' library("sf")
 #' # Fetch data from NASA POWER using 'sf' method
 #' data("lonlatsf", package = "climatrends")
 #' 
@@ -69,7 +69,6 @@
 #' g
 #' }
 #' 
-#' @importFrom sf st_centroid st_geometry_type st_as_sf
 #' @importFrom stats dist hclust cutree
 #' @export
 get_timeseries <- function(object, day.one, ...) {

### 2020-05-22 R/internal_functions.R

@@ -75,17 +75,20 @@
 #' 
 #' .lonlat_from_sf(lonlat)
 #' 
-#' @importFrom sf st_geometry_type st_centroid
 #' @noRd
 .lonlat_from_sf <- function(object){
   # check geometry type
   type <- c("POINT", "POLYGON")
   
-  # check for supported types 
-  supp_type <- c(all(grepl(type[[1]], sf::st_geometry_type(object))),
-                 all(grepl(type[[2]], sf::st_geometry_type(object))))
+  args <- list(x = object)
   
-  if (!any(supp_type)) {
+  geo <- as.character(do.call("st_geometry_type", args))
+  
+  # check for supported types 
+  supp_type <- c(all(grepl(type[[1]], geo)),
+                 all(grepl(type[[2]], geo)))
+  
+  if (isFALSE(any(supp_type))) {
     stop("The sf geometry type is not supported. ",
          "Please provide a sf object of geometry type ",
          "'POINT' or 'POLYGON'\n")
@@ -93,15 +96,13 @@
   
   type <- type[which(supp_type)]
   
-  nr <- dim(object)[[1]]
-  
-  # find the sf_column
-  index <- attr(object, "sf_column")
-  
-  # get the sf column
-  lonlat <- object[[index]]
-  
   if (type == "POINT") {
+    
+    # find the sf_column
+    index <- attr(object, "sf_column")
+    
+    # get the sf column
+    lonlat <- object[[index]]
     
     # unlist the sf_column
     lonlat <- unlist(object[[index]])
@@ -111,12 +112,15 @@
   if (type == "POLYGON") {
     
     # set centroid to validade lonlat
-    lonlat <- sf::st_centroid(lonlat)
+    args <- list(x = object)
+    lonlat <- do.call("st_centroid", args)
     
     # unlist the sf_column
     lonlat <- unlist(lonlat)
     
   }
+  
+  nr <- length(lonlat) / 2
   
   lonlat <- matrix(lonlat,
                    nrow = nr,

### 2020-05-22 R/rainfall.R

@@ -87,6 +87,7 @@
 #' \donttest{
 #' # Using remote sources of climate data
 #' library("nasapower")
+#' library("sf")
 #' data("lonlatsf", package = "climatrends")
 #' 
 #' # some random dates provided as integers and coerced to Dates internally
@@ -222,20 +223,9 @@ rainfall.sf <- function(object, day.one, span = NULL, ...,
   
   indices <- .rainfall_indices(rain, timeseries, intervals)
   
-  if (all(as.sf, timeseries)) {
-    
-    xy <- .lonlat_from_sf(object)
-    xy <- as.data.frame(xy)
-    xy$id <- 1:dim(xy)[[1]]
-    xy <- merge(xy, indices, by = "id")
-    
-    indices <- sf::st_as_sf(xy, coords = c("lon", "lat"), crs = 4326)
-    
-  }
-  
   if (isTRUE(as.sf) & isFALSE(timeseries)) {
     
-    indices <- suppressWarnings(sf::st_bind_cols(object, indices))
+    indices <- cbind(object, indices)
     
   }
   

### 2020-05-22 R/temperature.R

@@ -105,6 +105,7 @@
 #' 
 #' # Using remote sources of climate data
 #' library("nasapower")
+#' library("sf")
 #' data("lonlatsf", package = "climatrends")
 #' 
 #' # some random dates provided as integers and coerced to Dates internally
@@ -256,20 +257,9 @@ temperature.sf <- function(object, day.one, span = NULL, ...,
   
   indices <- .temperature_indices(temp, timeseries, intervals)
   
-  if (all(as.sf, timeseries)) {
-  
-    xy <- .lonlat_from_sf(object)
-    xy <- as.data.frame(xy)
-    xy$id <- 1:dim(xy)[[1]]
-    xy <- merge(xy, indices, by = "id")
-    
-    indices <- sf::st_as_sf(xy, coords = c("lon", "lat"), crs = 4326)
-      
-  }
-  
   if (isTRUE(as.sf) & isFALSE(timeseries)) {
     
-    indices <- suppressWarnings(sf::st_bind_cols(object, indices))
+    indices <- cbind(object, indices)
   
   }
   

### 2020-05-24 inst/paper/paper.Rmd
beans_data

@@
 library("kableExtra")
-library("magrittr")
+library("tidyverse")
+library("climatrends")
+library("PlackettLuce")
 funtbl <- read.csv("available_functions.csv")
+load("inst/paper/cbean.rda")
 ```
@@
 
+```{r }
+```
+
 ## Time series

### 2020-05-24 inst/paper/cbean.R

@@ -1,0 +1,88 @@
+library("climatrends")
+library("tidyverse")
+library("PlackettLuce")
+
+load("inst/paper/cbean.rda")
+
+
+# compute the number of days required to accumulate gdd from 
+# planting date to maturity
+gdd <- GDD(modis, 
+           day.one = cbean$planting_date, 
+           degree.days = 900, 
+           return.as = "ndays")
+
+# add gdd to the cbean data
+# and take the average of gdd per season
+cbean %<>%  
+  mutate(gdd = gdd$gdd) %>% 
+  group_by(season) %>% 
+  mutate(gdds = as.integer(mean(gdd)))
+  
+# compute the temperature indices from planting date to the number of days 
+# required to accumulate the gdd in each season
+temp <- temperature(modis, 
+                    day.one = cbean$planting_date, 
+                    span = cbean$gdds)
+
+cbean <- cbind(cbean, temp)
+
+plot(density(cbean$maxNT))
+plot(density(cbean$gdd))
+
+plt <- pltree(G ~ maxNT, data = cbean, minsize = 200)
+
+plot(plt, abbrev = 6)
+
+library("patchwork")
+
+A <- 
+ggplot(cbean, aes(gdd)) +
+  geom_density() +
+  labs(x = "GDD", y = "Density", title = "A") +
+  theme_bw() + 
+  theme(panel.background = element_blank(), 
+        panel.grid = element_blank(),
+        plot.title = element_text(size=16, 
+                                  colour = "grey20", 
+                                  face = "bold"),
+        axis.text.x = element_text(size = 12, angle = 0,
+                                   face="plain", colour = "grey20"),
+        axis.title.x = element_text(size=12, colour = "grey20"),
+        axis.text.y = element_text(size=12, angle = 0,
+                                   hjust=1, vjust=0.5,
+                                   face="plain", colour = "grey20"),
+        axis.title.y = element_text(size=12, colour = "grey20"))
+
+B <- 
+ggplot(cbean, aes(season, maxNT)) +
+  geom_violin(aes(fill = season), show.legend = FALSE) +
+  geom_jitter(height = 0, width = 0.1) +
+  scale_fill_manual(values = c('#d7191c','#fdae61','#ffffbf','#abd9e9','#2c7bb6')) +
+  labs(x = "Season", y = "maxNT", title = "B") +
+  theme_bw() + 
+  theme(panel.background = element_blank(), 
+        panel.grid = element_blank(),
+        plot.title = element_text(size=16, 
[…]

### 2020-05-24 inst/paper/redbeans.R

@@ -1,32 +1,0 @@
-library("climatrends")
-library("tidyverse")
-library("PlackettLuce")
-
-load("inst/paper/redbeans.rda")
-
-
-gdd <- GDD(modis, 
-          day.one = redbeans$planting_date, 
-          degree.days = 900, 
-          return.as = "ndays")
-
-
-gdd %<>%
-  mutate(season = redbeans$season) %>% 
-  group_by(season) %>% 
-  summarise(gdd = round(mean(gdd), 0))
-  
-
-redbeans <- merge(redbeans, gdd, by = "season")
-
-temp <- temperature(modis, day.one = redbeans$planting_date, span = redbeans$gdd)
-
-dat <- cbind(G = redbeans$G, temp)
-
-plot(density(dat$maxNT))
-
-pl <- pltree(G ~ maxNT, data = dat, minsize = 200, bonferroni = TRUE)
-
-print(pl)
-
-plot(pl)

### 2020-05-25 inst/paper/paper.Rmd
paper_cbeans

--- 2020-05-22 inst/paper/paper.Rmd
@@
 bibliography: paper.bib
-output: pdf_document
+output:
+  pdf_document:
+    keep_tex: true
+header-includes: \usepackage{caption} \captionsetup[figure]{labelformat=empty}
 ---
@@
 library("kableExtra")
-library("magrittr")
+library("tidyverse")
+library("climatrends")
+library("PlackettLuce")
 funtbl <- read.csv("available_functions.csv")
+load("cbean.rda")
 ```
@@
 
-**climatrends** aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
+`climatrends` aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
 
@@
 [...continue...]
-
-```{r characteristics, echo=FALSE, message=TRUE}
-kable(funtbl, 
-      row.names = FALSE,
-      align = "l", 
-      caption = "Main functions available in climatrends.") %>% 
-  kable_styling(latex_options = "scale_down") %>% 
-  column_spec(2, width = "1in") %>% 
-  column_spec(3, width = "1.7in") %>% 
-  column_spec(4, width = "1.7in")
-```
 
@@
 
+## Seed germination or some GDD related analysis
+
+Use the data from seed germination or crop growth to compute GDD. How many GDD a seed need to become a seedling?
+
 ## Common beans
 
-Replicate part of the analysis in van Etten et al [-@vanEtten2019] with the beans data to show how we can use this package to capture the influence of climate variability on crop performance.
-The idea is to show the same PlackettLuce Tree.
+During five growing seasons (from 2015 to 2016 in Nicaragua), van Etten et al [-@vanEtten2019] conducted a crowdsourcing citizen-science trial testing 11 common beans varieties (*Phaseolus vulgaris* L.) in 842 farmer-managed experimental plots.
+Sets of three varieties were allocated randomly to farms as incomplete blocks, maintaining spatial balance by assigning roughly equal frequencies of the varieties to each area.
+For the analysis the authors applied the Plackett–Luce model, which estimates for each variety the probability that it wins, beating all other varieties in the set [@Turner2020].
+The research showed that this approach can register known specific effects of climate variation on crop varietal performance.
+An earlier version of `climatrends` was used in this research to capture the seasonal climate variation, here we replicate part of this analysis in the matter that concerns the calculation and application of the climate indices.
+
+Is important to mention that this analysis shaped the implementation of the methods `temperature.array()` and `rainfall.matrix()`.
+
+```{r ggd_bean, message=FALSE, echo=TRUE, eval=FALSE}
+library("climatrends")
+library("PlackettLuce")
+library("tidyverse")
+
+# compute the number of days required to accumulate
+# gdd from planting date to maturity
+gdd <- GDD(modis, 
+           day.one = cbean$planting_date, 
+           degree.days = 900, 
+           return.as = "ndays")
+
+# add gdd to the cbean data and take the average 
+# of gdd per season
+cbean %<>%  
+  mutate(gdd = gdd$gdd) %>% 
+  group_by(season) %>% 
+  mutate(gdds = as.integer(mean(gdd)))
+```
+
+```{r pltree_bean, message=FALSE, echo=TRUE, eval=FALSE}
+# compute the temperature indices from planting date to the 
+# number of days required to accumulate the gdd in each season
+temp <- temperature(modis, 
+                    day.one = cbean$planting_date, 
+                    span = cbean$gdds)
+
+# combine the indices with the main data
+cbean <- cbind(cbean, temp)
+
+# fit a Plackett-Luce tree
+plt <- pltree(G ~ maxNT, data = cbean, minsize = 200)
+```
 
@@
 Pick some random points in Norway (or Scandinavia??) and check how the trends on temperature indices over the last 20 years.
-
-## Seed germination or some GDD related analysis
-
-Use the data from seed germination or crop growth to compute GDD. How many GDD a seed need to become a seedling?
 
@@
 # References
+
+<div id="refs"></div>
+
+\pagebreak
+
+# Figures and tables
+
+```{r table1, echo=FALSE, message=TRUE}
+kable(funtbl, 
+      row.names = FALSE,
+      align = "l", 
+      caption = "Main functions available in climatrends.") %>% 
+  kable_styling(latex_options = "scale_down") %>% 
+  column_spec(2, width = "1in") %>% 
+  column_spec(4, width = "1.7in")
+```
+
+\pagebreak
+
+```{r fig_cbean, fig.cap="Fig.
+1.
+Application of climatrends functions to support the analysis of a citizen-science data testing 11 common bean varieties in Nicaragua.
+(A) Days required to reach 900 growing-degree days from planting date calculated using the function GDD().
+(B) Maximum night temperature (°C) distributed across seasons computed using the function temperature().
+(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in two different nodes splitted with the maximum night temperature (°C).", out.width = '80%', echo=FALSE}
+knitr::include_graphics("cbean.png")
+```

### 2020-05-25 inst/paper/paper.tex
paper_cbeans

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Climate variability indices for ecological and crop models in R: the climatrends package},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption} \captionsetup[figure]{labelformat=empty}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{Climate variability indices for ecological and crop models in R:
the \texttt{climatrends} package}
\author{}
\date{\vspace{-2.5em}25 May 2020}

\begin{document}
\maketitle

\begin{quote}
Kauê de Sousa\textsuperscript{1,2{[}*{]}}, Jacob van
Etten\textsuperscript{2}, Svein Øivind Solberg\textsuperscript{1}\\
\textsuperscript{1} Department of Agricultural Sciences, Inland Norway
University of Applied Sciences, 2318 Hamar, Norway\\
\textsuperscript{2} Bioversity International, 00054 Maccarese, Rome,
Italy\\
\textsuperscript{{[}*{]}}Correspondence should be addressed to:
\href{mailto:kaue.desousa@inn.no}{\nolinkurl{kaue.desousa@inn.no}}
\end{quote}

\hfill\break

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Abiotic factors plays an important role in most ecological and crop
systems that depends on certain levels of temperature, light and
precipitation (and their interplay) to initiate important physiological
events (Schulze et al. 2019). In the walk of climate change, understand
how these factors drives the physiological processes is a key approach
to provide recommendations for adaptation and biodiversity conservation.
If translated to stress

Raw temperature and precipitation values are \ldots{} to describe how
organisms interacts with the enviroment

\texttt{climatrends} aims to provide the R (R Core Team 2020) toolkit to
compute extreme precipitation and temperature indices that serves as
input for climate and crop models (van Etten et al. 2019; Kehel, Crossa,
and Reynolds 2016), trends in climate change (Aguilar et al. 2005; de
Sousa et al. 2018) and applied ecology (Prentice et al. 1992; Liu and
El-Kassaby 2018).

{[}\ldots continue\ldots{]}

\hypertarget{methods-and-features}{%
\section{Methods and features}\label{methods-and-features}}

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

Six main functions are provided, \texttt{crop\_sensitive()},
\texttt{ETo()}, \texttt{GDD()}, \texttt{late\_frost()},
\texttt{rainfall()} and \texttt{temperature()} with a default method for
numeric `vector's and additional methods implemented via the package
\texttt{methods} (R Core Team 2020) for classes 'matrix' (or array),
`data.frame', and `sf' (of geometry POINT or POLYGON) (Pebesma 2018).
The later two are designed to fetch data from cloud sources.

say that the idea for the functions started with citizen science
projects that is why \texttt{day.one} and \texttt{span} may be variable
across locations. For time series analysis where fixed periods are
defined across many locations the indices can be adjusted with last.day.

{[}\ldots continue\ldots{]}

\hypertarget{temperature-and-precipitation-indices}{%
\subsection{Temperature and precipitation
indices}\label{temperature-and-precipitation-indices}}

\hypertarget{growing-degree-days}{%
\subsection{Growing degree-days}\label{growing-degree-days}}

Growing degree-days (gdd) is an heuristic tool in phenology that
measures heat accumulation and is used to predict plant and animal
development rates (Prentice et al. 1992). Growing degree-days are
calculated by taking the integral of warmth above a base temperature
(\(T_{0}\)). The function \texttt{GDD()} applies by default the
following equation.

Equation {[}1{]}

\[GDD = \frac{T_{max} + T_{min}}{2} - T_{0}\]

Where \(T_{max}\) is the maximum temperature in the given day,
\(T_{min}\) is the minimum temperature in the given day and \(T_{0}\) is
the minimum temperature for growth (as per the physiology of the focal
organism or ecosystem averages).

Additionally, the function \texttt{GDD()} offers three modified
equations designed for cold environments and for tropical environments.
For cold environments, where \(T_{min}\) may be lower than \(T_{0}\),
there are two modified equations that adjusts either \(T_{mean}\)
(variant a) or \(T_{min}\) (variant b). The variant a changes
\(T_{mean}\) to \(T_{0}\) if \(T_{mean} < T_{0}\) and is expressed as
follows.

Equation {[}2{]}

\[ GDD = max \left(\frac{T_{max} + T_{min}}{2} - T_{0}, \; 0 \right)\]

The variant b, is calculated using Equation 1, but adjusts \(T_{min}\)
or \(T_{max}\) to \(T_{0}\) if \(T < T_{0}\), the equation is adjusted
as follows.

Equation {[}3{]}

\[ T < T_{0} \; \rightarrow \; T = T_{0} \]

Where \(T\) may refer to \(T_{min}\) and/or \(T_{max}\) when the
condition of being below \(T_{0}\) applies.

For tropical areas, where the temperature may surpass a maximum
threshold (\(T_{0^{max}}\)), resulting in limited development, the
minimum temperature is adjusted using Equation 3 and the maximum
temperature is adjusted to a maximum base temperature as follow.

Equation {[}4{]}

\[ T_{max} > T_{0^{max}} \; \rightarrow \; T_{max} = T_{0^{max}} \]

Where \(T_{0^{max}}\) is the maximum base temperature for growth,
defined in \texttt{GDD()} using the argument \texttt{tbase\_max}.

These modified equations are defined as `a', `b' and `c', respectively,
and can be selected using the argument \texttt{equation}.

By default, the function returns the degree-days that is accumulated
over the time series using Equation 1. Additionally, the function may
return the daily values of degree-days or the number of days that a
given organism required to reach a certain number of accumulated
degree-days. These values are defined by `acc', `daily' or `ndays' and
can be adjusted using the argument \texttt{return.as}. The required
accumulated gdd is defined with argument \texttt{degree.days}. For
example, the Korean pine (\emph{Pinus koraiensis}) requires 105
\(^\circ C\) accumulated gdd to onset the photosynthesis (Wu et al.
2013). In that case, \texttt{GDD()} will calculate the growing
degree-days (\(gdd\)) and sum up the values until it reaches 105
\(^\circ C\) and return the number of days required in the given season
(\(GDD_{r}\)), as follows.

Equation {[}5{]}

\[\parallel GDD_{r} \parallel \: = \; ggd_1 \;+ \; ...  \; +  \; gdd_n\]

\hypertarget{late-spring-frost}{%
\subsection{Late-spring frost}\label{late-spring-frost}}

\hypertarget{crop-related-indices}{%
\subsection{Crop-related indices}\label{crop-related-indices}}

Two functions in \textbf{climatrends} are mainly designed to capture the
effects of climate on the development and stress of crop species,
\texttt{crop\_sensitive} computes indices that aims to capture the
changes in temperature extremes during key phenological stages
(e.g.~anthesis), and \texttt{ETo()} computes the reference
evapotranspiration.

The reference evapotranspiration measures the influence of the climate
on a given organism water needs, generally a crop species (Brouwer and
Heibloem 1986). The function \texttt{ETo()} applies the Blaney-Criddle
method, a general theoretical method used when only air-temperature is
available locally. It should be noted that this method is not very
accurate and aims to provide the order of magnitude of
evapotranspitation. The reference evapotranspiration is calculated using
the following equation.

Equation {[}6{]}

\[ETo = p \times \left(0.46 \times \frac{T_{max} + T_{min}}{2} + 8 \right) \times K_c\]

Where \(p\) is the mean daily percentage of annual daytime hours,
\(T_{max}\) is the maximum temperature, \(T_{min}\) is the minimum
temperature, and \(K_c\) is the factor for organism water need.

The percentage of daytime hours (\(p\)) is calculated internally by the
`data.frame' and `sf' methods in \texttt{ETo()} using the given latitude
(taken from the inputted \texttt{object}) and date (taken from the
inputted \texttt{day.one}). It matches the latitude and date with a
table of daylight percentage derived from Brouwer and Heibloem (1986).
The table can be verified using \texttt{climatrends:::daylight}.

\hypertarget{examples}{%
\section{Examples}\label{examples}}

\hypertarget{seed-germination-or-some-gdd-related-analysis}{%
\subsection{Seed germination or some GDD related
analysis}\label{seed-germination-or-some-gdd-related-analysis}}

Use the data from seed germination or crop growth to compute GDD. How
many GDD a seed need to become a seedling?

\hypertarget{common-beans}{%
\subsection{Common beans}\label{common-beans}}

During five growing seasons (from 2015 to 2016 in Nicaragua), van Etten
et al (2019) conducted a crowdsourcing citizen-science trial testing 11
common beans varieties (\emph{Phaseolus vulgaris} L.) in 842
farmer-managed experimental plots. Sets of three varieties were
allocated randomly to farms as incomplete blocks, maintaining spatial
balance by assigning roughly equal frequencies of the varieties to each
area. For the analysis the authors applied the Plackett--Luce model,
which estimates for each variety the probability that it wins, beating
all other varieties in the set (Turner et al. 2020). The research showed
that this approach can register known specific effects of climate
variation on crop varietal performance. An earlier version of
\texttt{climatrends} was used in this research to capture the seasonal
climate variation, here we replicate part of this analysis in the matter
that concerns the calculation and application of the climate indices.

Is important to mention that this analysis shaped the implementation of
the methods \texttt{temperature.array()} and \texttt{rainfall.matrix()}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"climatrends"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"PlackettLuce"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"tidyverse"}\NormalTok{)}

\CommentTok{\# compute the number of days required to accumulate}
\CommentTok{\# gdd from planting date to maturity}
\NormalTok{gdd <{-}}\StringTok{ }\KeywordTok{GDD}\NormalTok{(modis, }
           \DataTypeTok{day.one =}\NormalTok{ cbean}\OperatorTok{$}\NormalTok{planting\_date, }
           \DataTypeTok{degree.days =} \DecValTok{900}\NormalTok{, }
           \DataTypeTok{return.as =} \StringTok{"ndays"}\NormalTok{)}

\CommentTok{\# add gdd to the cbean data and take the average }
\CommentTok{\# of gdd per season}
\NormalTok{cbean }\OperatorTok{\%<>\%}\StringTok{  }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{gdd =}\NormalTok{ gdd}\OperatorTok{$}\NormalTok{gdd) }\OperatorTok{\%>\%}\StringTok{ }
\StringTok{  }\KeywordTok{group\_by}\NormalTok{(season) }\OperatorTok{\%>\%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{gdds =} \KeywordTok{as.integer}\NormalTok{(}\KeywordTok{mean}\NormalTok{(gdd)))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# compute the temperature indices from planting date to the }
\CommentTok{\# number of days required to accumulate the gdd in each season}
\NormalTok{temp <{-}}\StringTok{ }\KeywordTok{temperature}\NormalTok{(modis, }
                    \DataTypeTok{day.one =}\NormalTok{ cbean}\OperatorTok{$}\NormalTok{planting\_date, }
                    \DataTypeTok{span =}\NormalTok{ cbean}\OperatorTok{$}\NormalTok{gdds)}

\CommentTok{\# combine the indices with the main data}
\NormalTok{cbean <{-}}\StringTok{ }\KeywordTok{cbind}\NormalTok{(cbean, temp)}

\CommentTok{\# fit a Plackett{-}Luce tree}
\NormalTok{plt <{-}}\StringTok{ }\KeywordTok{pltree}\NormalTok{(G }\OperatorTok{\textasciitilde{}}\StringTok{ }\NormalTok{maxNT, }\DataTypeTok{data =}\NormalTok{ cbean, }\DataTypeTok{minsize =} \DecValTok{200}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{time-series}{%
\subsection{Time series}\label{time-series}}

Pick some random points in Norway (or Scandinavia??) and check how the
trends on temperature indices over the last 20 years.

\hypertarget{further-development}{%
\section{Further development}\label{further-development}}

Integration with other datasets as they become available in \texttt{R}
via API client packages. New indices related to the physiology of crops
to be implemented while I work on the rice data.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This work was supported by The Nordic Joint Committee for Agricultural
and Food Research (grant num. 202100-2817).

\hypertarget{authors-contributions}{%
\section{Authors' contributions}\label{authors-contributions}}

KdS and JvE conceived the ideas of implementing an \texttt{R} package.
JvE and SØS assisted in the implementation of methods. KdS wrote and
documented the software. {[}add other contributions\ldots{]}

\hypertarget{data-availability-statement}{%
\section{Data availability
statement}\label{data-availability-statement}}

The source code has been archived at {[}\textbf{add link}{]} as
\texttt{climatrends} version \textbf{X.Y.Z}. To explore the latest
functionalities of \texttt{climatrends}, please check the package's
updates at CRAN (\url{https://cran.r-project.org/package=climatrends}).

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Aguilar2005}{}%
Aguilar, E., T. C. Peterson, P. Ramírez Obando, R. Frutos, J. A. Retana,
M. Solera, J. Soley, et al. 2005. ``Changes in precipitation and
temperature extremes in Central America and northern South America,
1961--2003.'' \emph{Journal of Geophysical Research} 110 (D23): D23107.
\url{https://doi.org/10.1029/2005JD006119}.

\leavevmode\hypertarget{ref-Brouwer1986}{}%
Brouwer, C., and M. Heibloem. 1986. \emph{Irrigation water management:
Irrigation water needs}. Training m. Rome, Italy: Food; Agriculture
Organization of The United Nations.
\url{http://www.fao.org/3/S2022E/s2022e00.htm}.

\leavevmode\hypertarget{ref-deSousa2018}{}%
de Sousa, Kauê, Fernando Casanoves, Jorge Sellare, Alejandra Ospina,
Jose Gabriel Suchini, Amilcar Aguilar, and Leida Mercado. 2018. ``How
climate awareness influences farmers' adaptation decisions in Central
America?'' \emph{Journal of Rural Studies} 64: 11--19.
\url{https://doi.org/10.1016/j.jrurstud.2018.09.018}.

\leavevmode\hypertarget{ref-Kehel2016}{}%
Kehel, Z., J. Crossa, and M. Reynolds. 2016. ``Identifying Climate
Patterns during the Crop-Growing Cycle from 30 Years of CIMMYT Elite
Spring Wheat International Yield Trials.'' In \emph{Applied Mathematics
and Omics to Assess Crop Genetic Resources for Climate Change Adaptive
Traits}, edited by Abdallah Bari, Ardeshir B. Damania, Michael Mackay,
and Selvadurai Dayanandan, 151--74. CRC Press.

\leavevmode\hypertarget{ref-YLiu2018}{}%
Liu, Yang, and Yousry A. El-Kassaby. 2018. ``Evapotranspiration and
favorable growing degree-days are key to tree height growth and
ecosystem functioning: Meta-analyses of Pacific Northwest historical
data.'' \emph{Scientific Reports} 8228 (October).
\url{https://doi.org/10.1038/s41598-018-26681-1}.

\leavevmode\hypertarget{ref-sf}{}%
Pebesma, Edzer. 2018. ``Simple Features for R: Standardized Support for
Spatial Vector Data.'' \emph{The R Journal} 10 (1): 439--46.
\url{https://doi.org/10.32614/RJ-2018-009}.

\leavevmode\hypertarget{ref-Prentice1992}{}%
Prentice, I. Colin, Wolfgang Cramer, Sandy P. Harrison, Rik Leemans,
Robert A. Monserud, and Allen M. Solomon. 1992. ``Special Paper: A
Global Biome Model Based on Plant Physiology and Dominance, Soil
Properties and Climate.'' \emph{Journal of Biogeography} 19 (2): 117.
\url{https://doi.org/10.2307/2845499}.

\leavevmode\hypertarget{ref-RCoreTeam}{}%
R Core Team. 2020. ``R: A language and environment for statistical
computing. version 4.0.0.'' Vienna, Austria: CRAN R Project.
\url{https://www.r-project.org/}.

\leavevmode\hypertarget{ref-PlantEcology}{}%
Schulze, Ernst-Detlef, Erwin Beck, Nina Buchmann, Stephan Clemens, Klaus
Müller-Hohenstein, and Michael Scherer-Lorenzen. 2019. \emph{Plant
Ecology}. Second Edi. Berlin, Germany: Springer.
\url{https://doi.org/10.1007/978-3-662-56233-8}.

\leavevmode\hypertarget{ref-Turner2020}{}%
Turner, Heather L, Jacob van Etten, David Firth, and Ioannis Kosmidis.
2020. ``Modelling rankings in R: the PlackettLuce package.''
\emph{Computational Statistics}.
\url{https://doi.org/10.1007/s00180-020-00959-3}.

\leavevmode\hypertarget{ref-vanEtten2019}{}%
van Etten, Jacob, Kauê de Sousa, Amílcar Aguilar, Mirna Barrios, Allan
Coto, Matteo Dell'Acqua, Carlo Fadda, et al. 2019. ``Crop variety
management for climate adaptation supported by citizen science.''
\emph{Proceedings of the National Academy of Sciences} 116 (10):
4194--9. \url{https://doi.org/10.1073/pnas.1813720116}.

\leavevmode\hypertarget{ref-JWu2013}{}%
Wu, Jiabing, Dexin Guan, Fenhui Yuan, Anzhi Wang, and Changjie Jin.
2013. ``Soil Temperature Triggers the Onset of Photosynthesis in Korean
Pine.'' Edited by Carl J. Bernacchi. \emph{PLoS ONE} 8 (6): e65401.
\url{https://doi.org/10.1371/journal.pone.0065401}.
\end{cslreferences}

\pagebreak

\hypertarget{figures-and-tables}{%
\section{Figures and tables}\label{figures-and-tables}}

\begin{table}

\caption{\label{tab:table1}Main functions available in climatrends.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{1in}|l|>{}p{1.7in}}
\hline
Name & Function & Output\\
\hline
crop\_sensitive() & Compute crop sensitive indices & A data frame with crop sensitive indices with n columns depending on the number of thresholds passed to each index\\
\hline
Eto() & Reference evapotranspiration & The reference evapotranspiration\\
\hline
GDD() & Compute growing degree-days & Either the cumulative sum of gdd across time (the default), or the gdd obtained in each day, or the number of days required to reach a certain amount of gdd\\
\hline
late\_frost() & Compute the occurrence of late-spring frost & A data.frame with the duration and gdd accumulated during the events of frost, latency (where there is no frost event, but also there is no gdd), and warming (where gdd is accumulated)\\
\hline
rainfall() & Precipitation indices & Either the indices considering the series as a whole, or time series indices splitted into equal intervals of days\\
\hline
temperature() & Temperature indices & Either the indices considering the series as a whole, or time series indices splitted into equal intervals of days\\
\hline
\end{tabular}}
\end{table}

\pagebreak

\begin{figure}
\includegraphics[width=0.8\linewidth]{cbean} \caption{Fig. 1. Application of climatrends functions to support the analysis of a citizen-science data testing 11 common bean varieties in Nicaragua. (A) Days required to reach 900 growing-degree days from planting date calculated using the function GDD(). (B) Maximum night temperature (°C) distributed across seasons computed using the function temperature(). (C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in two different nodes splitted with the maximum night temperature (°C).}\label{fig:fig_cbean}
\end{figure}

\end{document}

### 2020-05-25 inst/paper/cbean.R

@@ -1,6 +1,9 @@
 library("climatrends")
 library("tidyverse")
 library("PlackettLuce")
+library("patchwork")
+library("ggparty")
+
 
 load("inst/paper/cbean.rda")
 
@@ -25,17 +28,19 @@ temp <- temperature(modis,
                     day.one = cbean$planting_date, 
                     span = cbean$gdds)
 
+# round values to 3 decimals, this will not change the main result, but will 
+# help in visualizing the tree 
+temp[1:ncol(temp)] <- lapply(temp[1:ncol(temp)], function(x) round(x, 3))
+
+# combine the indices with the main data
 cbean <- cbind(cbean, temp)
 
-plot(density(cbean$maxNT))
-plot(density(cbean$gdd))
-
+# fit a PL tree
 plt <- pltree(G ~ maxNT, data = cbean, minsize = 200)
 
-plot(plt, abbrev = 6)
-
-library("patchwork")
-
+# ..................................
+# ..................................
+# Prepare charts ####
 A <- 
 ggplot(cbean, aes(gdd)) +
   geom_density() +
@@ -57,9 +62,9 @@ ggplot(cbean, aes(gdd)) +
 B <- 
 ggplot(cbean, aes(season, maxNT)) +
   geom_violin(aes(fill = season), show.legend = FALSE) +
-  geom_jitter(height = 0, width = 0.1) +
+  #geom_jitter(height = 0, width = 0.1) +
   scale_fill_manual(values = c('#d7191c','#fdae61','#ffffbf','#abd9e9','#2c7bb6')) +
-  labs(x = "Season", y = "maxNT", title = "B") +
+  labs(x = "", y = "maxNT", title = "B") +
   theme_bw() + 
   theme(panel.background = element_blank(), 
         panel.grid = element_blank(),
@@ -74,15 +79,58 @@ ggplot(cbean, aes(season, maxNT)) +
                                    face="plain", colour = "grey20"),
         axis.title.y = element_text(size=12, colour = "grey20"))
 
+
+
+coeff <- as.vector(coef(plt, log = FALSE))
+items <- rep(dimnames(coef(plt))[[2]], each = 2)
+nit <- length(unique(items))
+
+coeff <- data.frame(coeff = coeff, items = items, 
+                    node = rep(c("Node 2","Node 3"),
+                               times = length(items)))
+
+D <-
+ggplot(coeff, aes(x = coeff, y = items, group = node)) +
[…]

### 2020-05-26 inst/paper/paper.Rmd
bean_plot

@@
 bibliography: paper.bib
+csl: citationstyle.csl
 output:
@@
 In the walk of climate change, understand how these factors drives the physiological processes is a key approach to provide recommendations for adaptation and biodiversity conservation.
-If translated to stress
 
-Raw temperature and precipitation values are ... to describe how organisms interacts with the enviroment
+[...continue...]
 
@@
 
+[...continue...]
+
 say that the idea for the functions started with citizen science projects that is why `day.one` and `span` may be variable across locations.
@@
 ## Temperature and precipitation indices
+
+Some explanation on the `temperature()` and `rainfall()` functions.
 
@@
 
-For tropical areas, where the temperature may surpass a maximum threshold ($T_{0^{max}}$), resulting in limited development, the minimum temperature is adjusted using Equation 3 and the maximum temperature is adjusted to a maximum base temperature as follow.
+For tropical areas, where the temperature may surpass a maximum threshold ($T_{0_{max}}$), resulting in limited development, the minimum temperature is adjusted using Equation 3 and the maximum temperature is adjusted to a maximum base temperature as follow.
 
@@
 
-$$ T_{max} > T_{0^{max}} \; \rightarrow \; T_{max} = T_{0^{max}} $$
+$$ T_{max} > T_{0_{max}} \; \rightarrow \; T_{max} = T_{0_{max}} $$
 
-Where $T_{0^{max}}$ is the maximum base temperature for growth, defined in `GDD()` using the argument `tbase_max`.
+Where $T_{0_{max}}$ is the maximum base temperature for growth, defined in `GDD()` using the argument `tbase_max`.
 
@@
 
+Explain how the `late_frost()` is implemented.
+
 ## Crop-related indices
 
-Two functions in **climatrends** are mainly designed to capture the effects of climate on the development and stress of crop species, `crop_sensitive` computes indices that aims to capture the changes in temperature extremes during key phenological stages (e.g.
+Two functions in **climatrends** are mainly designed to capture the effects of climate on the development and stress of crop species, `crop_sensitive()` computes indices that aims to capture the changes in temperature extremes during key phenological stages (e.g.
 anthesis), and `ETo()` computes the reference evapotranspiration.
 
-The reference evapotranspiration measures the influence of the climate on a given organism water needs, generally a crop species [@Brouwer1986].
+The reference evapotranspiration measures the influence of the climate on a given plant water needs [@Brouwer1986].
 The function `ETo()` applies the Blaney-Criddle method, a general theoretical method used when only air-temperature is available locally.
@@
 
-## Seed germination or some GDD related analysis
+## Some GDD related analysis
 
-Use the data from seed germination or crop growth to compute GDD. How many GDD a seed need to become a seedling?
+Use the data from seed germination or crop growth to compute GDD. Maybe some GDD analysis in Scandinavia with data from the reindeer project.
 
@@
 
-During five growing seasons (from 2015 to 2016 in Nicaragua), van Etten et al [-@vanEtten2019] conducted a crowdsourcing citizen-science trial testing 11 common beans varieties (*Phaseolus vulgaris* L.) in 842 farmer-managed experimental plots.
-Sets of three varieties were allocated randomly to farms as incomplete blocks, maintaining spatial balance by assigning roughly equal frequencies of the varieties to each area.
-For the analysis the authors applied the Plackett–Luce model, which estimates for each variety the probability that it wins, beating all other varieties in the set [@Turner2020].
-The research showed that this approach can register known specific effects of climate variation on crop varietal performance.
-An earlier version of `climatrends` was used in this research to capture the seasonal climate variation, here we replicate part of this analysis in the matter that concerns the calculation and application of the climate indices.
+During five growing seasons (from 2015 to 2017 in Nicaragua), van Etten et al.
+[-@vanEtten2019] conducted a crowdsourcing citizen-science trial testing 11 common beans varieties (*Phaseolus vulgaris* L.) in 842 farmer-managed experimental plots.
+Sets of three varieties were allocated randomly to farms as incomplete blocks.
+A Plackett–Luce model was used to analyse the data, this model estimates for each variety the probability that it wins, beating all other varieties in the set [@Turner2020].
+An earlier version of `climatrends` was used in this research to capture the seasonal climate variation, here we reproduce part of this analysis regarding calculation and application of the climate indices.
+The approach here is slightly different because if considers the growing-degree days from planting date to maturity (the earlier study used planting date to the end of reproductive stage) and add new indices to illustrate the package implementation.
 
-Is important to mention that this analysis shaped the implementation of the methods `temperature.array()` and `rainfall.matrix()`.
+The data used here is available as supplementary materials (**zenodo ref**) as `cbean`.
+This contains a data.frame with a Plackett-Luce grouped rankings, the geographical coordinates of each sampled plot and the planting dates from where each farmer decided to start the experiment.
+The dates can differ from each other in the same season.
+The temperature data used was the land surface temperature MODIS (MYD11A2) [@Wan2015] and is storaged as an array with two layers (1st for the day and 2nd for the night temperatures).
+Each column corresponds to the dates (from 2015-09-10 to 2017-06-09) and the rows corresponds to the rows in the `cbean` data.frame.
+
+Since the phenological stages were not available, we estimate these stages based on the amount of growing degree-days required to reach a given stage using the function `GDD()`.
+For common beans, we define 900 degree-days, from planting date to maturity [@deMedeiros2016].
+The input data is the array with the temperature data, the vector with planting dates (`cbean$planting_date`), the required amount of degree-days passed to the argument `degree.days` and the character string 'ndays' specifying that the function must return the values as number of days.
+`GDD()` calls internally the function `get_timeseries()` which will match the given dates in `day.one` with the column names in the array and concatenate the values for each row.
+Then `GDD()` computes the degree-days for the time series and return the length of the vector where the accumulated gdd reached the pre-defined threshold (900).
+
+The degree-days spanned from 54 to 100 days as shown in Fig. 2a. For simplicity we take the average per season and use this vector to compute the temperature indices.
 
@@
 
+To compute the temperature indices we use the array with tempeture data, the vector with planting dates, and the seasonal averaged degree-days passed as a vector using the argument `span`.
+`temperature()` concatenates the data from the given `day.one` to the given `span` and compute the indices for each row.
+
+In van Etten [-@vanEtten2019], a forward variable selection was applied to retain the most representative covariates based on the deviance reduction.
+This analysis retained the maximum night temperature (maxNT) as the most representative covariate.
+To illustrate how the Plackett-Luce trees can grow in complexity as we add more indices, we included the summer days (SU, number of days with maximum day temperature > 30 $^\circ C$) together with maxNT.
+
 ```{r pltree_bean, message=FALSE, echo=TRUE, eval=FALSE}
@@
 # fit a Plackett-Luce tree
-plt <- pltree(G ~ maxNT, data = cbean, minsize = 200)
+plt <- pltree(G ~ maxNT + SU, data = cbean, minsize = 50)
 ```
+
+Across-season distribution of maxNT captured for each sample plot in this experiment is shown in Fig.
+2b.
+The data has a bimodal distribution which is reflected in the splitting value (18.7 $^\circ C$) for the Plackett-Luce trees in Fig.
+2c.
+The upper node splits with 49 summer days (SU).
+We can interpret these results as that differences in growing performance of common beans is led by a considerable amount of diurnal temperature above a warmer threshold of 30 $^\circ C$ (in this case >70% of the growing days) and warmer nights (> 18.7 $^\circ C$).
 
@@
 (B) Maximum night temperature (°C) distributed across seasons computed using the function temperature().
-(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in two different nodes splitted with the maximum night temperature (°C).", out.width = '80%', echo=FALSE}
+(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in three different nodes splitted with the summer days (day temperature > 30 °C) and maximum night temperature (°C).
+Note: the first season (primera, Pr) spans from May to August, the second (postrera, Po) from September to October, and the third (apante, Ap) from November to January.", out.width = '80%', echo=FALSE}
 knitr::include_graphics("cbean.png")

### 2020-05-26 inst/paper/paper.tex
bean_plot

@@
 precipitation (and their interplay) to initiate important physiological
-events (Schulze et al. 2019). In the walk of climate change, understand
+events (Schulze et al., 2019). In the walk of climate change, understand
 how these factors drives the physiological processes is a key approach
 to provide recommendations for adaptation and biodiversity conservation.
-If translated to stress
 
-Raw temperature and precipitation values are \ldots{} to describe how
-organisms interacts with the enviroment
+{[}\ldots continue\ldots{]}
 
-\texttt{climatrends} aims to provide the R (R Core Team 2020) toolkit to
-compute extreme precipitation and temperature indices that serves as
-input for climate and crop models (van Etten et al. 2019; Kehel, Crossa,
-and Reynolds 2016), trends in climate change (Aguilar et al. 2005; de
-Sousa et al. 2018) and applied ecology (Prentice et al. 1992; Liu and
-El-Kassaby 2018).
+\texttt{climatrends} aims to provide the R (R Core Team, 2020) toolkit
+to compute extreme precipitation and temperature indices that serves as
+input for climate and crop models (Kehel, Crossa, \& Reynolds, 2016; van
+Etten et al., 2019), trends in climate change (Aguilar et al., 2005; de
+Sousa et al., 2018) and applied ecology (Prentice et al., 1992; Liu \&
+El-Kassaby, 2018).
 
@@
 numeric `vector's and additional methods implemented via the package
-\texttt{methods} (R Core Team 2020) for classes 'matrix' (or array),
-`data.frame', and `sf' (of geometry POINT or POLYGON) (Pebesma 2018).
+\texttt{methods} (R Core Team, 2020) for classes 'matrix' (or array),
+`data.frame', and `sf' (of geometry POINT or POLYGON) (Pebesma, 2018).
 The later two are designed to fetch data from cloud sources.
+
+{[}\ldots continue\ldots{]}
 
@@
 
+Some explanation on the \texttt{temperature()} and \texttt{rainfall()}
+functions.
+
 \hypertarget{growing-degree-days}{%
@@
 measures heat accumulation and is used to predict plant and animal
-development rates (Prentice et al. 1992). Growing degree-days are
+development rates (Prentice et al., 1992). Growing degree-days are
 calculated by taking the integral of warmth above a base temperature
@@
 example, the Korean pine (\emph{Pinus koraiensis}) requires 105
-\(^\circ C\) accumulated gdd to onset the photosynthesis (Wu et al.
-2013). In that case, \texttt{GDD()} will calculate the growing
-degree-days (\(gdd\)) and sum up the values until it reaches 105
-\(^\circ C\) and return the number of days required in the given season
-(\(GDD_{r}\)), as follows.
+\(^\circ C\) accumulated gdd to onset the photosynthesis (Wu, Guan,
+Yuan, Wang, \& Jin, 2013). In that case, \texttt{GDD()} will calculate
+the growing degree-days (\(gdd\)) and sum up the values until it reaches
+105 \(^\circ C\) and return the number of days required in the given
+season (\(GDD_{r}\)), as follows.
 
@@
 \subsection{Late-spring frost}\label{late-spring-frost}}
+
+Explain how the \texttt{late\_frost()} is implemented.
 
@@
 The reference evapotranspiration measures the influence of the climate
-on a given organism water needs, generally a crop species (Brouwer and
-Heibloem 1986). The function \texttt{ETo()} applies the Blaney-Criddle
+on a given organism water needs, generally a crop species (Brouwer \&
+Heibloem, 1986). The function \texttt{ETo()} applies the Blaney-Criddle
 method, a general theoretical method used when only air-temperature is
@@
 
-\hypertarget{seed-germination-or-some-gdd-related-analysis}{%
-\subsection{Seed germination or some GDD related
-analysis}\label{seed-germination-or-some-gdd-related-analysis}}
+\hypertarget{some-gdd-related-analysis}{%
+\subsection{Some GDD related analysis}\label{some-gdd-related-analysis}}
 
-Use the data from seed germination or crop growth to compute GDD. How
-many GDD a seed need to become a seedling?
+Use the data from seed germination or crop growth to compute GDD. Maybe
+some GDD analysis in Scandinavia with data from the reindeer project.
 
@@
 
-During five growing seasons (from 2015 to 2016 in Nicaragua), van Etten
-et al (2019) conducted a crowdsourcing citizen-science trial testing 11
+During five growing seasons (from 2015 to 2017 in Nicaragua), van Etten
+et al.~(2019) conducted a crowdsourcing citizen-science trial testing 11
 common beans varieties (\emph{Phaseolus vulgaris} L.) in 842
 farmer-managed experimental plots. Sets of three varieties were
-allocated randomly to farms as incomplete blocks, maintaining spatial
-balance by assigning roughly equal frequencies of the varieties to each
-area. For the analysis the authors applied the Plackett--Luce model,
-which estimates for each variety the probability that it wins, beating
-all other varieties in the set (Turner et al. 2020). The research showed
-that this approach can register known specific effects of climate
-variation on crop varietal performance. An earlier version of
+allocated randomly to farms as incomplete blocks. A Plackett--Luce model
+was used to analyse the data, this model estimates for each variety the
+probability that it wins, beating all other varieties in the set
+(Turner, van Etten, Firth, \& Kosmidis, 2020). An earlier version of
 \texttt{climatrends} was used in this research to capture the seasonal
-climate variation, here we replicate part of this analysis in the matter
-that concerns the calculation and application of the climate indices.
+climate variation, here we reproduce part of this analysis regarding
+calculation and application of the climate indices. The approach here is
+slightly different because if considers the growing-degree days from
+planting date to maturity (the earlier study used planting date to the
+end of reproductive stage) and add new indices to illustrate the package
+implementation.
 
-Is important to mention that this analysis shaped the implementation of
-the methods \texttt{temperature.array()} and \texttt{rainfall.matrix()}.
+The data used here is available as supplementary materials
+(\textbf{zenodo ref}) as \texttt{cbean}. This contains a data.frame with
+a Plackett-Luce grouped rankings, the geographical coordinates of each
+sampled plot and the planting dates from where each farmer decided to
+start the experiment. The dates can differ from each other in the same
+season. The temperature data used was the land surface temperature MODIS
+(MYD11A2) (Wan, Hook, \& Hulley, 2015) and is storaged as an array with
+two layers (1st for the day and 2nd for the night temperatures). Each
+column corresponds to the dates (from 2015-09-10 to 2017-06-09) and the
+rows corresponds to the rows in the \texttt{cbean} data.frame.
+
+Since the phenological stages were not available, we estimate these
+stages based on the amount of growing degree-days required to reach a
+given stage using the function \texttt{GDD()}. For common beans, we
+define 900 degree-days, from planting date to maturity (de Medeiros,
+Daniel, \& Fengler, 2016). The input data is the array with the
+temperature data, the vector with planting dates
+(\texttt{cbean\$planting\_date}), the required amount of degree-days
+passed to the argument \texttt{degree.days} and the character string
+`ndays' specifying that the function must return the values as number of
+days. \texttt{GDD()} calls internally the function
+\texttt{get\_timeseries()} which will match the given dates in
+\texttt{day.one} with the column names in the array and concatenate the
+values for each row. Then \texttt{GDD()} computes the degree-days for
+the time series and return the length of the vector where the
+accumulated gdd reached the pre-defined threshold (900).
+
+The degree-days spanned from 54 to 100 days as shown in Fig. 2a. For
+simplicity we take the average per season and use this vector to compute
+the temperature indices.
 
@@
 
+To compute the temperature indices we use the array with tempeture data,
+the vector with planting dates, and the seasonal averaged degree-days
+passed as a vector using the argument \texttt{span}.
+\texttt{temperature()} concatenates the data from the given
+\texttt{day.one} to the given \texttt{span} and compute the indices for
+each row.
+
+In van Etten (2019), a forward variable selection was applied to retain
+the most representative covariates based on the deviance reduction. This
+analysis retained the maximum night temperature (maxNT) as the most
+representative covariate. To illustrate how the Plackett-Luce trees can
+grow in complexity as we add more indices, we included the summer days
+(SU, number of days with maximum day temperature \textgreater{} 30
+\(^\circ C\)) together with maxNT.
+
 \begin{Shaded}
@@
 \CommentTok{\# fit a Plackett{-}Luce tree}
-\NormalTok{plt <{-}}\StringTok{ }\KeywordTok{pltree}\NormalTok{(G }\OperatorTok{\textasciitilde{}}\StringTok{ }\NormalTok{maxNT, }\DataTypeTok{data =}\NormalTok{ cbean, }\DataTypeTok{minsize =} \DecValTok{200}\NormalTok{)}
+\NormalTok{plt <{-}}\StringTok{ }\KeywordTok{pltree}\NormalTok{(G }\OperatorTok{\textasciitilde{}}\StringTok{ }\NormalTok{maxNT }\OperatorTok{+}\StringTok{ }\NormalTok{SU, }\DataTypeTok{data =}\NormalTok{ cbean, }\DataTypeTok{minsize =} \DecValTok{50}\NormalTok{)}
 \end{Highlighting}
 \end{Shaded}
+
+Across-season distribution of maxNT captured for each sample plot in
+this experiment is shown in Fig. 2b. The data has a bimodal distribution
+which is reflected in the splitting value (18.7 \(^\circ C\)) for the
+Plackett-Luce trees in Fig. 2c. The upper node splits with 49 summer
+days (SU). We can interpret these results as that differences in growing
+performance of common beans is led by a considerable amount of diurnal
+temperature above a warmer threshold of 30 \(^\circ C\) (in this case
+\textgreater70\% of the growing days) and warmer nights (\textgreater{}
+18.7 \(^\circ C\)).
 
@@
 \leavevmode\hypertarget{ref-Aguilar2005}{}%
-Aguilar, E., T. C. Peterson, P. Ramírez Obando, R. Frutos, J. A. Retana,
-M. Solera, J. Soley, et al. 2005. ``Changes in precipitation and
+Aguilar, E., Peterson, T. C., Obando, P. R., Frutos, R., Retana, J. A.,
+Solera, M., \ldots{} Mayorga, R. (2005). Changes in precipitation and
 temperature extremes in Central America and northern South America,
-1961--2003.'' \emph{Journal of Geophysical Research} 110 (D23): D23107.
-\url{https://doi.org/10.1029/2005JD006119}.
+1961--2003. \emph{Journal of Geophysical Research}, \emph{110}(D23),
+D23107.
+doi:\href{https://doi.org/10.1029/2005JD006119}{10.1029/2005JD006119}
 
 \leavevmode\hypertarget{ref-Brouwer1986}{}%
-Brouwer, C., and M. Heibloem. 1986. \emph{Irrigation water management:
-Irrigation water needs}. Training m. Rome, Italy: Food; Agriculture
-Organization of The United Nations.
-\url{http://www.fao.org/3/S2022E/s2022e00.htm}.
+Brouwer, C., \& Heibloem, M. (1986). \emph{Irrigation water management:
+Irrigation water needs} (Training m). Rome, Italy: Food; Agriculture
+Organization of The United Nations. Retrieved from
+\url{http://www.fao.org/3/S2022E/s2022e00.htm}
+
+\leavevmode\hypertarget{ref-deMedeiros2016}{}%
+de Medeiros, G. A., Daniel, L. A., \& Fengler, F. H. (2016). Growth,
+Development, and Water Consumption of Irrigated Bean Crop Related to
+Growing Degree-Days on Different Soil Tillage Systems in Southeast
+Brazil. \emph{International Journal of Agronomy}, \emph{2016}, 8065985.
+doi:\href{https://doi.org/10.1155/2016/8065985}{10.1155/2016/8065985}
 
 \leavevmode\hypertarget{ref-deSousa2018}{}%
-de Sousa, Kauê, Fernando Casanoves, Jorge Sellare, Alejandra Ospina,
-Jose Gabriel Suchini, Amilcar Aguilar, and Leida Mercado. 2018. ``How
-climate awareness influences farmers' adaptation decisions in Central
-America?'' \emph{Journal of Rural Studies} 64: 11--19.
-\url{https://doi.org/10.1016/j.jrurstud.2018.09.018}.
+de Sousa, K., Casanoves, F., Sellare, J., Ospina, A., Suchini, J. G.,
+Aguilar, A., \& Mercado, L. (2018). How climate awareness influences
+farmers' adaptation decisions in Central America? \emph{Journal of Rural
+Studies}, \emph{64}, 11--19.
+doi:\href{https://doi.org/10.1016/j.jrurstud.2018.09.018}{10.1016/j.jrurstud.2018.09.018}
 
 \leavevmode\hypertarget{ref-Kehel2016}{}%
-Kehel, Z., J. Crossa, and M. Reynolds. 2016. ``Identifying Climate
+Kehel, Z., Crossa, J., \& Reynolds, M. (2016). Identifying Climate
 Patterns during the Crop-Growing Cycle from 30 Years of CIMMYT Elite
-Spring Wheat International Yield Trials.'' In \emph{Applied Mathematics
-and Omics to Assess Crop Genetic Resources for Climate Change Adaptive
-Traits}, edited by Abdallah Bari, Ardeshir B. Damania, Michael Mackay,
-and Selvadurai Dayanandan, 151--74. CRC Press.
+Spring Wheat International Yield Trials. In A. Bari, A. B. Damania, M.
+Mackay, \& S. Dayanandan (Eds.), \emph{Applied mathematics and omics to
+assess crop genetic resources for climate change adaptive traits} (pp.
+151--174). CRC Press.
 
 \leavevmode\hypertarget{ref-YLiu2018}{}%
-Liu, Yang, and Yousry A. El-Kassaby. 2018. ``Evapotranspiration and
-favorable growing degree-days are key to tree height growth and
-ecosystem functioning: Meta-analyses of Pacific Northwest historical
-data.'' \emph{Scientific Reports} 8228 (October).
-\url{https://doi.org/10.1038/s41598-018-26681-1}.
+Liu, Y., \& El-Kassaby, Y. A. (2018). Evapotranspiration and favorable
+growing degree-days are key to tree height growth and ecosystem
+functioning: Meta-analyses of Pacific Northwest historical data.
+\emph{Scientific Reports}, \emph{8228}.
+doi:\href{https://doi.org/10.1038/s41598-018-26681-1}{10.1038/s41598-018-26681-1}
 
 \leavevmode\hypertarget{ref-sf}{}%
-Pebesma, Edzer. 2018. ``Simple Features for R: Standardized Support for
-Spatial Vector Data.'' \emph{The R Journal} 10 (1): 439--46.
-\url{https://doi.org/10.32614/RJ-2018-009}.
+Pebesma, E. (2018). Simple Features for R: Standardized Support for
+Spatial Vector Data. \emph{The R Journal}, \emph{10}(1), 439--446.
+doi:\href{https://doi.org/10.32614/RJ-2018-009}{10.32614/RJ-2018-009}
 
 \leavevmode\hypertarget{ref-Prentice1992}{}%
-Prentice, I. Colin, Wolfgang Cramer, Sandy P. Harrison, Rik Leemans,
-Robert A. Monserud, and Allen M. Solomon. 1992. ``Special Paper: A
-Global Biome Model Based on Plant Physiology and Dominance, Soil
-Properties and Climate.'' \emph{Journal of Biogeography} 19 (2): 117.
-\url{https://doi.org/10.2307/2845499}.
+Prentice, I. C., Cramer, W., Harrison, S. P., Leemans, R., Monserud, R.
+A., \& Solomon, A. M. (1992). Special Paper: A Global Biome Model Based
+on Plant Physiology and Dominance, Soil Properties and Climate.
+\emph{Journal of Biogeography}, \emph{19}(2), 117.
+doi:\href{https://doi.org/10.2307/2845499}{10.2307/2845499}
 
 \leavevmode\hypertarget{ref-RCoreTeam}{}%
-R Core Team. 2020. ``R: A language and environment for statistical
-computing. version 4.0.0.'' Vienna, Austria: CRAN R Project.
-\url{https://www.r-project.org/}.
+R Core Team. (2020). R: A language and environment for statistical
+computing. version 4.0.0. Vienna, Austria: CRAN R Project. Retrieved
+from \url{https://www.r-project.org/}
 
 \leavevmode\hypertarget{ref-PlantEcology}{}%
-Schulze, Ernst-Detlef, Erwin Beck, Nina Buchmann, Stephan Clemens, Klaus
-Müller-Hohenstein, and Michael Scherer-Lorenzen. 2019. \emph{Plant
-Ecology}. Second Edi. Berlin, Germany: Springer.
-\url{https://doi.org/10.1007/978-3-662-56233-8}.
+Schulze, E.-D., Beck, E., Buchmann, N., Clemens, S., Müller-Hohenstein,
+K., \& Scherer-Lorenzen, M. (2019). \emph{Plant Ecology} (Second Edi, p.
+925). Berlin, Germany: Springer.
+doi:\href{https://doi.org/10.1007/978-3-662-56233-8}{10.1007/978-3-662-56233-8}
 
 \leavevmode\hypertarget{ref-Turner2020}{}%
-Turner, Heather L, Jacob van Etten, David Firth, and Ioannis Kosmidis.
-2020. ``Modelling rankings in R: the PlackettLuce package.''
-\emph{Computational Statistics}.
-\url{https://doi.org/10.1007/s00180-020-00959-3}.
+Turner, H. L., van Etten, J., Firth, D., \& Kosmidis, I. (2020).
+Modelling rankings in R: the PlackettLuce package. \emph{Computational
+Statistics}.
+doi:\href{https://doi.org/10.1007/s00180-020-00959-3}{10.1007/s00180-020-00959-3}
 
 \leavevmode\hypertarget{ref-vanEtten2019}{}%
-van Etten, Jacob, Kauê de Sousa, Amílcar Aguilar, Mirna Barrios, Allan
-Coto, Matteo Dell'Acqua, Carlo Fadda, et al. 2019. ``Crop variety
-management for climate adaptation supported by citizen science.''
-\emph{Proceedings of the National Academy of Sciences} 116 (10):
-4194--9. \url{https://doi.org/10.1073/pnas.1813720116}.
+van Etten, J., de Sousa, K., Aguilar, A., Barrios, M., Coto, A.,
+Dell'Acqua, M., \ldots{} Steinke, J. (2019). Crop variety management for
+climate adaptation supported by citizen science. \emph{Proceedings of
+the National Academy of Sciences}, \emph{116}(10), 4194--4199.
+doi:\href{https://doi.org/10.1073/pnas.1813720116}{10.1073/pnas.1813720116}
+
+\leavevmode\hypertarget{ref-Wan2015}{}%
+Wan, Z., Hook, S., \& Hulley, G. (2015). MYD11A1 MODIS/Aqua Land Surface
+Temperature/Emissivity Daily L3 Global 1km SIN Grid V006 {[}Data set{]}.
+\emph{NASA EOSDIS LP DAAC}.
+doi:\href{https://doi.org/10.5067/MODIS/MYD11A1.006}{10.5067/MODIS/MYD11A1.006}
 
 \leavevmode\hypertarget{ref-JWu2013}{}%
-Wu, Jiabing, Dexin Guan, Fenhui Yuan, Anzhi Wang, and Changjie Jin.
-2013. ``Soil Temperature Triggers the Onset of Photosynthesis in Korean
-Pine.'' Edited by Carl J. Bernacchi. \emph{PLoS ONE} 8 (6): e65401.
-\url{https://doi.org/10.1371/journal.pone.0065401}.
+Wu, J., Guan, D., Yuan, F., Wang, A., \& Jin, C. (2013). Soil
+Temperature Triggers the Onset of Photosynthesis in Korean Pine.
+\emph{PLoS ONE}, \emph{8}(6), e65401.
+doi:\href{https://doi.org/10.1371/journal.pone.0065401}{10.1371/journal.pone.0065401}
 \end{cslreferences}
@@
 (B) Maximum night temperature (°C) distributed across seasons computed using the function temperature().
-(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in two different nodes splitted with the maximum night temperature (°C).}\label{fig:fig_cbean}
+(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in three different nodes splitted with the summer days (day temperature > 30 °C) and maximum night temperature (°C).
+Note: the first season (primera, Pr) spans from May to August, the second (postrera, Po) from September to October, and the third (apante, Ap) from November to January.}\label{fig:fig_cbean}
 \end{figure}

### 2020-05-26 inst/paper/cbean.R

@@ -2,6 +2,7 @@ library("climatrends")
 library("tidyverse")
 library("PlackettLuce")
 library("patchwork")
+library("qvcalc")
 library("ggparty")
 
 
@@ -30,13 +31,15 @@ temp <- temperature(modis,
 
 # round values to 3 decimals, this will not change the main result, but will 
 # help in visualizing the tree 
-temp[1:ncol(temp)] <- lapply(temp[1:ncol(temp)], function(x) round(x, 3))
+temp[1:ncol(temp)] <- lapply(temp[1:ncol(temp)], function(x) round(x, 2))
 
 # combine the indices with the main data
 cbean <- cbind(cbean, temp)
 
 # fit a PL tree
-plt <- pltree(G ~ maxNT, data = cbean, minsize = 200)
+plt <- pltree(G ~ maxNT + SU, data = cbean, minsize = 50)
+
+plot(plt)
 
 # ..................................
 # ..................................
@@ -44,82 +47,78 @@ plt <- pltree(G ~ maxNT, data = cbean, minsize = 200)
 A <- 
 ggplot(cbean, aes(gdd)) +
   geom_density() +
-  labs(x = "GDD", y = "Density", title = "A") +
+  labs(x = "GDD", y = "Density", title = "(a)") +
   theme_bw() + 
   theme(panel.background = element_blank(), 
         panel.grid = element_blank(),
         plot.title = element_text(size=16, 
                                   colour = "grey20", 
                                   face = "bold"),
-        axis.text.x = element_text(size = 12, angle = 0,
+        axis.text.x = element_text(size = 10, angle = 0,
                                    face="plain", colour = "grey20"),
-        axis.title.x = element_text(size=12, colour = "grey20"),
-        axis.text.y = element_text(size=12, angle = 0,
+        axis.title.x = element_text(size=10, colour = "grey20"),
+        axis.text.y = element_text(size=10, angle = 0,
                                    hjust=1, vjust=0.5,
                                    face="plain", colour = "grey20"),
-        axis.title.y = element_text(size=12, colour = "grey20"))
+        axis.title.y = element_text(size=10, colour = "grey20"))
 
 B <- 
 ggplot(cbean, aes(season, maxNT)) +
-  geom_violin(aes(fill = season), show.legend = FALSE) +
+  geom_violin(aes(fill = season), alpha = 0.5, show.legend = FALSE) +
[…]

### 2020-06-30 inst/paper/paper.Rmd
paper

@@
   pdf_document:
-    keep_tex: true
+    keep_tex: false
 header-includes: \usepackage{caption} \captionsetup[figure]{labelformat=empty}
@@
 library("PlackettLuce")
-funtbl <- read.csv("available_functions.csv")
 load("cbean.rda")
@@
 
-Abiotic factors plays an important role in most ecological and crop systems that depends on certain levels of temperature, light and precipitation (and their interplay) to initiate important physiological events [@PlantEcology].
-In the walk of climate change, understand how these factors drives the physiological processes is a key approach to provide recommendations for adaptation and biodiversity conservation.
+Abiotic factors play an important role in most ecological and crop systems that depend on certain levels of temperature, light and precipitation (and their interplay) to initiate important physiological events [@PlantEcology].
+In the walk of climate change, understanding how these factors drive the physiological processes is a key approach to provide recommendations for adaptation and biodiversity conservation.
 
@@
 
-`climatrends` aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serves as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
+`climatrends` aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serve as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
 
@@
 
-Six main functions are provided, `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector's and additional methods implemented via the package `methods` [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
-The later two are designed to fetch data from cloud sources.
+Six main functions are provided, `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector' and additional methods implemented via the package `methods` [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
+The last two methods are designed to fetch data from cloud sources.
 
@@
 
-Some explanation on the `temperature()` and `rainfall()` functions.
+The indices computed by the functions `temperature()` and `rainfall()` are described in Table 2.
 
@@
 
-Where $T_{max}$ is the maximum temperature in the given day, $T_{min}$ is the minimum temperature in the given day and $T_{0}$ is the minimum temperature for growth (as per the physiology of the focal organism or ecosystem averages).
+where $T_{max}$ is the maximum temperature in the given day, $T_{min}$ is the minimum temperature in the given day and $T_{0}$ is the minimum temperature for growth (as per the physiology of the focal organism or ecosystem averages).
 
 Additionally, the function `GDD()` offers three modified equations designed for cold environments and for tropical environments.
-For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two modified equations that adjusts either $T_{mean}$ (variant a) or $T_{min}$ (variant b).
-The variant a changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follows.
+For cold environments, where $T_{min}$ may be lower than $T_{0}$, there are two modified equations that adjust either $T_{mean}$ (variant a) or $T_{min}$ (variant b).
+The variant a changes $T_{mean}$ to $T_{0}$ if $T_{mean} < T_{0}$ and is expressed as follow.
 
@@
 
-Where $T$ may refer to $T_{min}$ and/or $T_{max}$ when the condition of being below $T_{0}$ applies.
+where $T$ may refer to $T_{min}$ and/or $T_{max}$ when the condition of being below $T_{0}$ applies.
 
@@
 
-Where $T_{0_{max}}$ is the maximum base temperature for growth, defined in `GDD()` using the argument `tbase_max`.
+where $T_{0_{max}}$ is the maximum base temperature for growth, defined in `GDD()` using the argument `tbase_max`.
 
@@
 
+where $GDD_{r}$ is the length of the vector with accumulated degree-days from day 1 to $n$.
+
 ## Late-spring frost
 
-Explain how the `late_frost()` is implemented.
+Late-spring frost is a freezing event occurring after a substantial accumulation of warmth.
+Frost damage is a known issue in temperate and boreal regions, it is associated with the formation of extracellular ice crystals that cause damage in the membranes [@Lambers2008].
+Freezing occurring after an advanced phenological stage during spring may harm some plant species, resulting in lost of productivity in crop systems [@Trnka2014] and important ecological impacts [@Zohner2020].
+
+The function `late_frost()` supports the computation of late-spring frost events.
+The function counts for the number of freezing days with minimum temperature below a certain threshold (argument `tfrost`).
+And returns the number of days spanned by frost events (temperature below `tfrost`), latency (event with no freezing temperature but also no accumulation of growing degree-days) and warming (when growing degree-days are accumulated enabling the development of the target organism).
+Additionally the function returns the first day of the events.
+The function calculates the growing degree-days applying the variant b (Eq.
+3), which can be adjusted using the argument `equation` passed to `GDD()` as explained in the later section.
+The main inputs are a vector with maximum and minimum temperatures to compute the degree-days, a vector of dates (argument `date`), and, if needed, the `tbase` and `tfrost`, set by default to 4 and -2 $^\circ C$.
 
@@
 
-Two functions in **climatrends** are mainly designed to capture the effects of climate on the development and stress of crop species, `crop_sensitive()` computes indices that aims to capture the changes in temperature extremes during key phenological stages (e.g.
+Two functions in **climatrends** are mainly designed to capture the effects of climate on the development and stress of crop species, `crop_sensitive()` computes indices that aim to capture the changes in temperature extremes during key phenological stages (e.g.
 anthesis), and `ETo()` computes the reference evapotranspiration.
 
-The reference evapotranspiration measures the influence of the climate on a given plant water needs [@Brouwer1986].
+The crop sensitive indices available in **climatrends** are described in Table 3.
+These indices were previously used in crop models to project the impacts of climate change on crop yield [@Challinor2016; @Trnka2014].
+Each index has a default temperature threshold(s) which can be adjusted by using the arguments `*.threshold`.
+Where the `*` means the index.
+For example, to change the defaults for hts_max (high temperature stress), a vector with the temperature thresholds is passed through the argument `hts_max.thresholds`.
+
+The reference evapotranspiration measures the influence of the climate on a given plant's water need [@Brouwer1986].
 The function `ETo()` applies the Blaney-Criddle method, a general theoretical method used when only air-temperature is available locally.
@@
 
-Use the data from seed germination or crop growth to compute GDD. Maybe some GDD analysis in Scandinavia with data from the reindeer project.
+Use the data from seed germination or crop growth to compute GDD. Maybe a GDD analysis in Scandinavia with data from the reindeer project.
 
-## Common beans
+## Common bean
 
-During five growing seasons (from 2015 to 2017 in Nicaragua), van Etten et al.
-[-@vanEtten2019] conducted a crowdsourcing citizen-science trial testing 11 common beans varieties (*Phaseolus vulgaris* L.) in 842 farmer-managed experimental plots.
+During five growing seasons (from 2015 to 2017) in Nicaragua, van Etten et al.
+[-@vanEtten2019] conducted a crowdsourcing citizen-science experiment testing 11 common bean varieties (*Phaseolus vulgaris* L.) in 842 farmer-managed plots.
 Sets of three varieties were allocated randomly to farms as incomplete blocks.
@@
 An earlier version of `climatrends` was used in this research to capture the seasonal climate variation, here we reproduce part of this analysis regarding calculation and application of the climate indices.
-The approach here is slightly different because if considers the growing-degree days from planting date to maturity (the earlier study used planting date to the end of reproductive stage) and add new indices to illustrate the package implementation.
+The approach here is slightly different because it considers the growing-degree days from planting date to maturity (the earlier study used planting date to the end of reproductive stage) and add new indices to illustrate the package implementation.
 
-The data used here is available as supplementary materials (**zenodo ref**) as `cbean`.
+The data used here is available as supplementary material (**zenodo ref**) as `cbean`.
 This contains a data.frame with a Plackett-Luce grouped rankings, the geographical coordinates of each sampled plot and the planting dates from where each farmer decided to start the experiment.
-The dates can differ from each other in the same season.
+The planting dates differ from each other in the same season.
 The temperature data used was the land surface temperature MODIS (MYD11A2) [@Wan2015] and is storaged as an array with two layers (1st for the day and 2nd for the night temperatures).
@@
 
-```{r table1, echo=FALSE, message=TRUE}
-kable(funtbl, 
-      row.names = FALSE,
-      align = "l", 
-      caption = "Main functions available in climatrends.") %>% 
-  kable_styling(latex_options = "scale_down") %>% 
-  column_spec(2, width = "1in") %>% 
-  column_spec(4, width = "1.7in")
-```
-
-\pagebreak
-
 ```{r fig_cbean, fig.cap="Fig.
@@
 ```
+
+\pagebreak
+
+```{r table1, echo=FALSE, results="asis"}
+tabl <- as.data.frame(matrix(
+  c("crop_sensitive()", "Compute crop sensitive indices",
+    "ETo()", "Reference evapotranspiration using the Blaney-Criddle method",
+    "GDD()", "Compute growing degree-days",
+    "late_frost()","Compute the occurrence of late-spring frost",
+    "rainfall()","Precipitation indices",
+    "temperature()","Temperature indices"),
+  nrow = 6, ncol = 2, byrow = TRUE
+))
+
+names(tabl) <- c("Function", "Definition")
+
+kable(tabl, 
+      align = "l", 
+      caption = "Main functions available in climatrends.") %>% 
+  kable_styling(latex_options = "hold_position")
+```
+
+\pagebreak
+
+```{r table2, echo=FALSE, results="asis"}
+index <- as.data.frame(matrix(
+  c("maxDT", "Maximun day temperature", "°C",
+    "minDT", "Minimum day temperature", "°C",
+    "maxNT", "Maximun night temperature", "°C",
+    "minNT", "Minimum night temperature", "°C",
+    "DTR", "Diurnal temperature range (mean difference between DT and NT)", "°C",
+    "SU", "Summer days, number of days with maximum temperature > 30 °C", "days", 
+    "TR", "Tropical nights, number of nights with maximum temperature > 25 °C", "days",
+    "CFD", "Consecutive frosty days, number of days with temperature < 0 °C", "days",
+    "WSDI","Maximum warm spell duration, consecutive days with temperature > 90th percentile", "days",
+    "CSDI","Maximum cold spell duration, consecutive nights with temperature < 10th percentile","days",
+    "T10p","The 10th percentile of night tempeture","°C",
+    "T90p","The 90th percentile of day tempeture","°C",
+    "MLDS", "Maximum length of consecutive dry day, rain < 1 mm", "days",
+    "MLWS", "Maximum length of consecutive wet day, rain >= 1 mm", "days",
+    "R10mm", "Heavy precipitation days 10 >= rain < 20 mm", "days",
+    "R20mm", "Very heavy precipitation days rain >= 20", "days",
+    "Rx1day", "Maximum 1-day precipitation", "mm",
+    "Rx5day", "Maximum 5-day precipitation", "mm",
+    "R95p", "Total precipitation when rain > 95th percentile", "mm",
+    "R99p", "Total precipitation when rain > 99th percentile", "mm",
+    "Rtotal", "Total precipitation in wet days, rain >= 1 mm", "mm",
+    "SDII", "Simple daily intensity index, total precipitation divided by the number of wet days", "mm/days"),
+  nrow = 22, ncol = 3, byrow = TRUE
+))
+
+names(index) <- c("Index", "Definition", "Unit")
+
+kable(index,
+      caption = "Temperature and precipitation indices available in climatrends.") %>% 
+  kable_styling(latex_options = "hold_position") %>% 
+  column_spec(2, width = "20em")
+```
+
+```{r cropsentbl, echo=FALSE, results="asis"}
+index <- as.data.frame(matrix(
+  c("hts_mean", "High temperature stress using daily mean temperature, and given as percentage number of days a certain threshold is exceeded", "32, 35, 38 °C",
+    "hts_max", "High temperature stress using daily max temperature, and given as percentage number of days a certain threshold is exceeded", "36, 39, 42 °C",
+    "hse", "Heat stress event, and given as percentage number of days a certain threshold is exceeded for at least two consecutive days", "31 °C", 
+    "hse_ms", "Heat stress event, and given the maximum number of days a certain threshold is exceeded for at least two consecutive days", "31 °C",
+    "cdi_mean", "Crop duration index using daily mean temperature, and given as max(Tmean - threshold, 0)", "22, 23, 24 °C",
+    "cdi_max", "Crop duration index using daily max temperature, and given as max(Tmax - threshold, 0)", "27, 28, 29 °C",
+    "lethal", "Lethal temperatures, defined as percentage of days during the timeseries where daily mean temperature exceeds a given threshold", "43, 46, 49 °C"),
+  nrow = 7, ncol = 3, byrow = TRUE
+))
+
+names(index) <- c("Index", "Definition", "Default thresholds")
+
+kable(index,
+      caption = "Crop sensitive indices computed by climatrends.") %>% 
+  kable_styling(latex_options = "hold_position") %>% 
+  column_spec(2, width = "20em")
+
+```

### 2020-06-30 R/rainfall.R

@@ -315,7 +315,7 @@ rainfall.sf <- function(object, day.one, span = NULL, ...,
              .r_one_day(y),
              .r_five_day(y),
              .very_wet_days(y),
-             .extrem_wet_days(y),
+             .extreme_wet_days (y),
              .r_total(y),
              .sdii(y))
       
@@ -362,7 +362,7 @@ rainfall.sf <- function(object, day.one, span = NULL, ...,
                       .r_one_day(x),
                       .r_five_day(x),
                       .very_wet_days(x),
-                      .extrem_wet_days(x),
+                      .extreme_wet_days (x),
                       .r_total(x),
                       .sdii(x),
                       stringsAsFactors = FALSE)
@@ -398,7 +398,7 @@ rainfall.sf <- function(object, day.one, span = NULL, ...,
 #' set.seed(12)
 #' r <- runif(20, 0, 9)
 #' r[c(1,4,9:12,17)] <- 0
-#' chirps:::.dryspell(r)
+#' climatrends:::.dryspell(r)
 #' @noRd
 .dryspell <- function(x){
   if (all(is.na(x))) {
@@ -444,7 +444,7 @@ rainfall.sf <- function(object, day.one, span = NULL, ...,
 #' set.seed(12)
 #' r <- runif(20, 0, 9)
 #' r[c(1,4,9:11)] <- 0.1
-#' chirps:::.wetspell(r)
+#' climatrends:::.wetspell(r)
 #' @noRd
 .wetspell <- function(x){
   
@@ -488,7 +488,7 @@ rainfall.sf <- function(object, day.one, span = NULL, ...,
 #' set.seed(12)
 #' r <- runif(20, 0, 12)
 #' r[c(1,4,9:11)] <- 0.1
-#' chirps:::.r_ten_mm(r)
+#' climatrends:::.r_ten_mm(r)
 #' @noRd
 .r_ten_mm <- function(x) {
   
@@ -510,7 +510,7 @@ rainfall.sf <- function(object, day.one, span = NULL, ...,
 #' set.seed(12)
 #' r <- runif(20, 10, 23)
 #' r[c(1,4,9:11)] <- 0.1
-#' chirps:::.r_twenty_mm(r)
+#' climatrends:::.r_twenty_mm(r)
 #' @noRd
[…]

### 2020-06-30 inst/paper/cbean.R

@@ -29,12 +29,17 @@ temp <- temperature(modis,
                     day.one = cbean$planting_date, 
                     span = cbean$gdds)
 
+rain <- rainfall(chirps, 
+                    day.one = cbean$planting_date, 
+                    span = cbean$gdds)
+
+
 # round values to 3 decimals, this will not change the main result, but will 
 # help in visualizing the tree 
 temp[1:ncol(temp)] <- lapply(temp[1:ncol(temp)], function(x) round(x, 2))
 
 # combine the indices with the main data
-cbean <- cbind(cbean, temp)
+cbean <- cbind(cbean, temp, rain)
 
 # fit a PL tree
 plt <- pltree(G ~ maxNT + SU, data = cbean, minsize = 50)
@@ -130,6 +135,5 @@ ggsave(paste0("inst/paper/cbean.png"),
        height = 14,
        dpi = 1000,
        units = "cm")
-
 
 

### 2020-06-30 inst/paper/scandinavia.R

@@ -1,0 +1,98 @@
+library("climatrends")
+library("patchwork")
+library("sf")
+library("nasapower")
+library("tidyverse")
+library("mapview")
+
+e <- matrix(c(7, 59,
+              17, 59,
+              17, 63,
+              7, 63,
+              7, 59),
+            nrow = 5, ncol = 2, byrow = TRUE)
+
+# e <- matrix(c(-7.5, 37.2,
+#                  -2, 37.2,
+#                  -2, 40.5,
+#               -7.5, 40.5,
+#               -7.5, 37.2),
+#             nrow = 5, ncol = 2, byrow = TRUE)
+e
+
+e <- st_polygon(list(e))
+plot(e)
+set.seed(435)
+p <- st_sample(e, 100, type = "hexagonal")
+p <- st_as_sf(p, crs = 4326)
+
+mapview(p)
+
+temp <- temperature(p, 
+                    day.one = "2020-03-01", 
+                    last.day = "2020-04-01", 
+                    timeseries = TRUE, 
+                    intervals = 365)
+
+i <- c("CSDI","WSDI")
+
+
+temp %>% 
+  #filter(index %in% i) %>% 
+  group_by(index) %>% 
+  mutate(ab = mean(value)) %>% 
+  ggplot() +
+  geom_line(aes(x = date, y = value, group = id)) + 
+  geom_hline(aes(yintercept = ab), colour = "red", lwd = 0.7) +
+  geom_smooth(aes(x = date, y = value), method = "loess", se = FALSE) +
+  facet_wrap(. ~ index, scales = "free") +
+  labs(x = "", y = "Index (days)") +
+  theme(axis.text.y = element_text(size = 12, angle = 0, hjust = 1, 
+                                   vjust = 0.5, face = "plain", colour = "black"),
+        axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5, 
+                                   vjust = 1, face = "plain", colour = "black"),
+        axis.title = element_text(size = 12, face = "plain"),
+        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
+        plot.background = element_blank(),
+        panel.background = element_blank(),
+        strip.text.x = element_text(size = 12, colour = "black"),
+        strip.text.y = element_text(size = 12, colour = "black"),
+        strip.background = element_rect(colour = "black", fill="#FFFFFF")) ->
+  gg1
+
+gg1
+
+
+i <- c("T10p","T90p")
+
+
+temp %>% 
[…]

### 2020-08-14 inst/paper/paper.Rmd
paper

@@
 
->Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
+>Kauê de Sousa^1,2[*]^, Magne Neby^3^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
 ^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
 ^2^ Bioversity International, 00054 Maccarese, Rome, Italy  
-^[*]^Correspondence should be addressed to: <kaue.desousa@inn.no>
+^3^ Department of Forest and Field Sciences, Inland Norway University of Applied Sciences, 2480 Koppang, Norway    
+^[*]^ Correspondence should be addressed to: <kaue.desousa@inn.no>
 
@@
 
-## Some GDD related analysis
-
-Use the data from seed germination or crop growth to compute GDD. Maybe a GDD analysis in Scandinavia with data from the reindeer project.
-
 ## Common bean
@@
 
-## Time series
+## Trends in climate variability in Norway and Sweden
 
-Pick some random points in Norway (or Scandinavia??) and check how the trends on temperature indices over the last 20 years.
+Pick some random points in Norway and Sweden and check how the trends on temperature indices over the last 20 years.
+
+## Impact of early thaw on vole populations
+
+From [Johnsen et al (2016)](https://doi.org/10.1002/ece3.2635) and [Boonstra et al (2016)](http://bioscience.oxfordjournals.org/content/early/2016/07/19/biosci.biw080.full.pdf).
+The snowpack of European boreal forest contains relatively dense snow because of snow falling at temperatures just below freezing and because of the formation of ice crusts owing to thaw–freeze cycles during the relatively mild winters.
+The formation of ice due to freeze–thaw cycles during winter may be limiting vole availability to bilberries and other edible food items by encrusting them in ice and therefore limiting the peak phase of the cycles.
+If there falls snow in autumn but then it becomes above zero and the snow starts melting but then it starts freezing again, an ice layer is formed which has important implications in ecology.
+This might be calculated / predicted with the `late_frost()` function, but then applied as 'early thaw' function.
 
@@
 
-This work was supported by The Nordic Joint Committee for Agricultural and Food Research (grant num. 202100-2817).
+This work was supported by The Nordic Joint Committee for Agricultural and Food Research (grant num.
+202100-2817).
+We thank Julian Ramirez-Villegas and Marcel Schrijvers-Gonlag for the useful insights and discussion that helped in the development of this study.
 

### 2020-08-14 inst/paper/climatology.R

@@ -5,19 +5,19 @@ library("nasapower")
 library("tidyverse")
 library("mapview")
 
-e <- matrix(c(7, 59,
-              17, 59,
-              17, 63,
-              7, 63,
-              7, 59),
-            nrow = 5, ncol = 2, byrow = TRUE)
-
-# e <- matrix(c(-7.5, 37.2,
-#                  -2, 37.2,
-#                  -2, 40.5,
-#               -7.5, 40.5,
-#               -7.5, 37.2),
+# e <- matrix(c(7, 59,
+#               17, 59,
+#               17, 63,
+#               7, 63,
+#               7, 59),
 #             nrow = 5, ncol = 2, byrow = TRUE)
+
+e <- matrix(c(-7.5, 37.2,
+                 -2, 37.2,
+                 -2, 40.5,
+              -7.5, 40.5,
+              -7.5, 37.2),
+            nrow = 5, ncol = 2, byrow = TRUE)
 e
 
 e <- st_polygon(list(e))
@@ -29,16 +29,18 @@ p <- st_as_sf(p, crs = 4326)
 mapview(p)
 
 temp <- temperature(p, 
-                    day.one = "2020-03-01", 
-                    last.day = "2020-04-01", 
+                    day.one = "2000-01-01", 
+                    last.day = "2019-12-31", 
                     timeseries = TRUE, 
                     intervals = 365)
+
+save(temp, file = "inst/paper/climatology.rda")
 
 i <- c("CSDI","WSDI")
 
 
 temp %>% 
-  #filter(index %in% i) %>% 
+  filter(index %in% i) %>% 
   group_by(index) %>% 
   mutate(ab = mean(value)) %>% 
   ggplot() +
@@ -67,7 +69,7 @@ i <- c("T10p","T90p")
 
 
 temp %>% 
-  filter(index %in% i) %>% 
+  #filter(index %in% i) %>% 
   group_by(index) %>% 
   mutate(ab = mean(value)) %>% 
   ggplot() +
@@ -94,5 +96,4 @@ gg2
 
 
 gg1 / gg2
-
 

### 2020-10-02 inst/paper/paper.Rmd
paper-remove-ecological-analysis

@@
     affiliation: "1, 2"
+  - name: Magne Neby
+    affiliation: "3"
   - name: Jacob van Etten
@@
     index: 2
+  - name: Department of Forest and Field Sciences, Inland Norway University of Applied Sciences, 2480 Koppang, Norway
+    index: 3
 citation_author: de Sousa et. al.
@@
 
-# Introduction
+# Summary
 
@@
 In the walk of climate change, understanding how these factors drive the physiological processes is a key approach to provide recommendations for adaptation and biodiversity conservation.
-
-[...continue...]
-
-`climatrends` aims to provide the R [@RCoreTeam] toolkit to compute extreme precipitation and temperature indices that serve as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
-
-[...continue...]
+The package `climatrends` aims to provide the methods in R [@RCoreTeam] to compute precipitation and temperature indices that serve as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
 
@@
 
-Six main functions are provided, `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector' and additional methods implemented via the package `methods` [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
-The last two methods are designed to fetch data from cloud sources.
+Six main functions are provided (Table 1), `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector' and additional methods implemented via the package `methods` [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
+The last two methods are designed to fetch data from cloud sources, until now from the packages `nasapower` [@Sparks2018] and `chirps` [@chirps].
 
-[...continue...]
+```{r table1, echo=FALSE, results="asis"}
+tabl <- as.data.frame(matrix(
+  c("crop_sensitive()", "Compute crop sensitive indices",
+    "ETo()", "Reference evapotranspiration using the Blaney-Criddle method",
+    "GDD()", "Compute growing degree-days",
+    "late_frost()","Compute the occurrence of late-spring frost",
+    "rainfall()","Precipitation indices",
+    "temperature()","Temperature indices"),
+  nrow = 6, ncol = 2, byrow = TRUE
+))
 
-say that the idea for the functions started with citizen science projects that is why `day.one` and `span` may be variable across locations.
-For time series analysis where fixed periods are defined across many locations the indices can be adjusted with last.day.
+names(tabl) <- c("Function", "Definition")
 
-[...continue...]
+kable(tabl, 
+      align = "l", 
+      caption = "Main functions available in climatrends.") %>% 
+  kable_styling(latex_options = "hold_position")
+```
+
+These functions started as a set of scripts to compute indices in citizen science trials.
+In these trials.
+Aiming to capture the environmental variation across different sites, which can differ since each data-point generally have a different starting day and duration, the arguments `day.one` and `span` are vectorised and may be variable across data-points.
+For time series analysis, where fixed periods are defined across many locations, the indices can be adjusted with the argument `last.day` linked to the argument `day.one`.
 
@@
 
+The package `climatrends` computes 12 temperature indices and 10 precipitation indices that were suggested by previous research on climatology and crop science [@Kehel2016; @Aguilar2005].
 The indices computed by the functions `temperature()` and `rainfall()` are described in Table 2.
+
+```{r table2, echo=FALSE, results="asis"}
+index <- as.data.frame(matrix(
+  c("maxDT", "Maximun day temperature", "°C",
+    "minDT", "Minimum day temperature", "°C",
+    "maxNT", "Maximun night temperature", "°C",
+    "minNT", "Minimum night temperature", "°C",
+    "DTR", "Diurnal temperature range (mean difference between DT and NT)", "°C",
+    "SU", "Summer days, number of days with maximum temperature > 30 °C", "days", 
+    "TR", "Tropical nights, number of nights with maximum temperature > 25 °C", "days",
+    "CFD", "Consecutive frosty days, number of days with temperature < 0 °C", "days",
+    "WSDI","Maximum warm spell duration, consecutive days with temperature > 90th percentile", "days",
+    "CSDI","Maximum cold spell duration, consecutive nights with temperature < 10th percentile","days",
+    "T10p","The 10th percentile of night tempeture","°C",
+    "T90p","The 90th percentile of day tempeture","°C",
+    "MLDS", "Maximum length of consecutive dry day, rain < 1 mm", "days",
+    "MLWS", "Maximum length of consecutive wet day, rain >= 1 mm", "days",
+    "R10mm", "Heavy precipitation days 10 >= rain < 20 mm", "days",
+    "R20mm", "Very heavy precipitation days rain >= 20", "days",
+    "Rx1day", "Maximum 1-day precipitation", "mm",
+    "Rx5day", "Maximum 5-day precipitation", "mm",
+    "R95p", "Total precipitation when rain > 95th percentile", "mm",
+    "R99p", "Total precipitation when rain > 99th percentile", "mm",
+    "Rtotal", "Total precipitation in wet days, rain >= 1 mm", "mm",
+    "SDII", "Simple daily intensity index, total precipitation divided by the number of wet days", "mm/days"),
+  nrow = 22, ncol = 3, byrow = TRUE
+))
+
+names(index) <- c("Index", "Definition", "Unit")
+
+kable(index,
+      caption = "Temperature and precipitation indices available in climatrends.") %>% 
+  kable_styling(latex_options = "hold_position") %>% 
+  column_spec(2, width = "20em")
+```
 
@@
 
+\pagebreak
+
 Equation [5]
@@
 For example, to change the defaults for hts_max (high temperature stress), a vector with the temperature thresholds is passed through the argument `hts_max.thresholds`.
+
+```{r cropsentbl, echo=FALSE, results="asis"}
+index <- as.data.frame(matrix(
+  c("hts_mean", "High temperature stress using daily mean temperature, and given as percentage number of days a certain threshold is exceeded", "32, 35, 38 °C",
+    "hts_max", "High temperature stress using daily max temperature, and given as percentage number of days a certain threshold is exceeded", "36, 39, 42 °C",
+    "hse", "Heat stress event, and given as percentage number of days a certain threshold is exceeded for at least two consecutive days", "31 °C", 
+    "hse_ms", "Heat stress event, and given the maximum number of days a certain threshold is exceeded for at least two consecutive days", "31 °C",
+    "cdi_mean", "Crop duration index using daily mean temperature, and given as max(Tmean - threshold, 0)", "22, 23, 24 °C",
+    "cdi_max", "Crop duration index using daily max temperature, and given as max(Tmax - threshold, 0)", "27, 28, 29 °C",
+    "lethal", "Lethal temperatures, defined as percentage of days during the timeseries where daily mean temperature exceeds a given threshold", "43, 46, 49 °C"),
+  nrow = 7, ncol = 3, byrow = TRUE
+))
+
+names(index) <- c("Index", "Definition", "Default thresholds")
+
+kable(index,
+      caption = "Crop sensitive indices computed by climatrends.") %>% 
+  kable_styling(latex_options = "hold_position") %>% 
+  column_spec(2, width = "20em")
+
+```
 
@@
 
+```{r fig_cbean, fig.cap="Fig.
+1.
+Application of climatrends functions to support the analysis of a citizen-science data testing 11 common bean varieties in Nicaragua.
+(A) Days required to reach 900 growing-degree days from planting date calculated using the function GDD().
+(B) Maximum night temperature (°C) distributed across seasons computed using the function temperature().
+(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in three different nodes splitted with the summer days (day temperature > 30 °C) and maximum night temperature (°C).
+Note: the first season (primera, Pr) spans from May to August, the second (postrera, Po) from September to October, and the third (apante, Ap) from November to January.", out.width = '80%', echo=FALSE}
+knitr::include_graphics("cbean.png")
+```
+
 ## Trends in climate variability in Norway and Sweden
@@
 <div id="refs"></div>
-
-\pagebreak
-
-# Figures and tables
-
-```{r fig_cbean, fig.cap="Fig.
-1.
-Application of climatrends functions to support the analysis of a citizen-science data testing 11 common bean varieties in Nicaragua.
-(A) Days required to reach 900 growing-degree days from planting date calculated using the function GDD().
-(B) Maximum night temperature (°C) distributed across seasons computed using the function temperature().
-(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in three different nodes splitted with the summer days (day temperature > 30 °C) and maximum night temperature (°C).
-Note: the first season (primera, Pr) spans from May to August, the second (postrera, Po) from September to October, and the third (apante, Ap) from November to January.", out.width = '80%', echo=FALSE}
-knitr::include_graphics("cbean.png")
-```
-
-\pagebreak
-
-```{r table1, echo=FALSE, results="asis"}
-tabl <- as.data.frame(matrix(
-  c("crop_sensitive()", "Compute crop sensitive indices",
-    "ETo()", "Reference evapotranspiration using the Blaney-Criddle method",
-    "GDD()", "Compute growing degree-days",
-    "late_frost()","Compute the occurrence of late-spring frost",
-    "rainfall()","Precipitation indices",
-    "temperature()","Temperature indices"),
-  nrow = 6, ncol = 2, byrow = TRUE
-))
-
-names(tabl) <- c("Function", "Definition")
-
-kable(tabl, 
-      align = "l", 
-      caption = "Main functions available in climatrends.") %>% 
-  kable_styling(latex_options = "hold_position")
-```
-
-\pagebreak
-
-```{r table2, echo=FALSE, results="asis"}
-index <- as.data.frame(matrix(
-  c("maxDT", "Maximun day temperature", "°C",
-    "minDT", "Minimum day temperature", "°C",
-    "maxNT", "Maximun night temperature", "°C",
-    "minNT", "Minimum night temperature", "°C",
-    "DTR", "Diurnal temperature range (mean difference between DT and NT)", "°C",
-    "SU", "Summer days, number of days with maximum temperature > 30 °C", "days", 
-    "TR", "Tropical nights, number of nights with maximum temperature > 25 °C", "days",
-    "CFD", "Consecutive frosty days, number of days with temperature < 0 °C", "days",
-    "WSDI","Maximum warm spell duration, consecutive days with temperature > 90th percentile", "days",
-    "CSDI","Maximum cold spell duration, consecutive nights with temperature < 10th percentile","days",
-    "T10p","The 10th percentile of night tempeture","°C",
-    "T90p","The 90th percentile of day tempeture","°C",
-    "MLDS", "Maximum length of consecutive dry day, rain < 1 mm", "days",
-    "MLWS", "Maximum length of consecutive wet day, rain >= 1 mm", "days",
-    "R10mm", "Heavy precipitation days 10 >= rain < 20 mm", "days",
-    "R20mm", "Very heavy precipitation days rain >= 20", "days",
-    "Rx1day", "Maximum 1-day precipitation", "mm",
-    "Rx5day", "Maximum 5-day precipitation", "mm",
-    "R95p", "Total precipitation when rain > 95th percentile", "mm",
-    "R99p", "Total precipitation when rain > 99th percentile", "mm",
-    "Rtotal", "Total precipitation in wet days, rain >= 1 mm", "mm",
-    "SDII", "Simple daily intensity index, total precipitation divided by the number of wet days", "mm/days"),
-  nrow = 22, ncol = 3, byrow = TRUE
-))
-
-names(index) <- c("Index", "Definition", "Unit")
-
-kable(index,
-      caption = "Temperature and precipitation indices available in climatrends.") %>% 
-  kable_styling(latex_options = "hold_position") %>% 
-  column_spec(2, width = "20em")
-```
-
-```{r cropsentbl, echo=FALSE, results="asis"}
-index <- as.data.frame(matrix(
-  c("hts_mean", "High temperature stress using daily mean temperature, and given as percentage number of days a certain threshold is exceeded", "32, 35, 38 °C",
-    "hts_max", "High temperature stress using daily max temperature, and given as percentage number of days a certain threshold is exceeded", "36, 39, 42 °C",
-    "hse", "Heat stress event, and given as percentage number of days a certain threshold is exceeded for at least two consecutive days", "31 °C", 
-    "hse_ms", "Heat stress event, and given the maximum number of days a certain threshold is exceeded for at least two consecutive days", "31 °C",
-    "cdi_mean", "Crop duration index using daily mean temperature, and given as max(Tmean - threshold, 0)", "22, 23, 24 °C",
-    "cdi_max", "Crop duration index using daily max temperature, and given as max(Tmax - threshold, 0)", "27, 28, 29 °C",
-    "lethal", "Lethal temperatures, defined as percentage of days during the timeseries where daily mean temperature exceeds a given threshold", "43, 46, 49 °C"),
-  nrow = 7, ncol = 3, byrow = TRUE
-))
-
-names(index) <- c("Index", "Definition", "Default thresholds")
-
-kable(index,
-      caption = "Crop sensitive indices computed by climatrends.") %>% 
-  kable_styling(latex_options = "hold_position") %>% 
-  column_spec(2, width = "20em")
-
-```

### 2020-10-02 inst/paper/vole.R

@@ -1,0 +1,132 @@
+# Activity, home range and thus trapability  vs.  variation in weather per session?
+# Activity, home range and thus trapability  vs.  season?
+# climatrends paper
+
+# # Import rawdata
+# rm(list=ls())
+# source("scripts/1_tidying_rawdata.R")
+# 
+# df <- alldata %>% # OBS this is bank voles from two different habitats / trapping designs
+#   select(session_start,
+#          individual_id,
+#          live_myodes_glareolus,
+#          dead_myodes_glareolus,
+#          trap_id,
+#          grid_id)
+# 
+# df <- df %>%
+#   filter(live_myodes_glareolus > 0 | dead_myodes_glareolus > 0) %>%
+#   select(-live_myodes_glareolus, -dead_myodes_glareolus)
+# 
+# df$session_start <- as.Date(as.character(df$session_start))
+# df <- df %>%
+#   filter(session_start > "2017-05-10") %>%
+#   filter(session_start < "2021-01-01")
+# 
+# write.csv(df, "processed_files/bank_vole_data_to_climatrends.csv",
+#           row.names = FALSE)
+
+# set environment
+library(dplyr)
+library(lubridate)
+library(ggplot2)
+se <- function(x) sqrt(var(x)/length(x))
+df <- read.csv("inst/paper/vole_data.csv")
+
+# select habitat / study design
+## Cross design:
+# df <- df %>%
+#   filter(grid_id != "M1") %>%
+#   filter(grid_id != "M2") %>%
+#   select(-grid_id)
+
+## or line transect:
+# df <- df %>%
+#   filter(grid_id == "M1" | grid_id == "M2") %>%
+#   select(-grid_id)
+
+## Merge month id in case we want to combine the two datasets/habitats and treat as one
+# df$session_start <- as.Date(paste(substr(df$session_start, 0,7), "10", sep = "-"))
+
+
+# make overview / table of what traps that was used by whom in each session (ie. month)
+dfSession <- df %>%
+  group_by(session_start) %>%
+  distinct(individual_id, trap_id) %>%
+  ungroup()
+
+
+# count number of unique traps an individual is trapped in during the same trapping session
+dfCount <- as.data.frame(table(dfSession$individual_id, dfSession$session_start)) 
+dfCount <- dfCount %>% #remove ghosts
+  filter(Freq > 0)
+
+# rename
+dfCount <- dfCount %>% 
[…]

### 2020-10-02 inst/paper/paper.Rmd
paper

@@
 citation_author: de Sousa et. al.
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: "A preprint"
 year: 2020
@@
     "CSDI","Maximum cold spell duration, consecutive nights with temperature < 10th percentile","days",
-    "T10p","The 10th percentile of night tempeture","°C",
-    "T90p","The 90th percentile of day tempeture","°C",
+    "T10p","The 10th percentile of night temperature","°C",
+    "T90p","The 90th percentile of day temperature","°C",
     "MLDS", "Maximum length of consecutive dry day, rain < 1 mm", "days",
@@
 
-The degree-days spanned from 54 to 100 days as shown in Fig. 2a. For simplicity we take the average per season and use this vector to compute the temperature indices.
+The degree-days spanned from 54 to 100 days as shown in Fig. 1a. For simplicity we take the average per season and use this vector to compute the temperature indices.
 
@@
 Across-season distribution of maxNT captured for each sample plot in this experiment is shown in Fig.
-2b.
+1b.
 The data has a bimodal distribution which is reflected in the splitting value (18.7 $^\circ C$) for the Plackett-Luce trees in Fig.
-2c.
+1c.
 The upper node splits with 49 summer days (SU).
@@
 (C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in three different nodes splitted with the summer days (day temperature > 30 °C) and maximum night temperature (°C).
-Note: the first season (primera, Pr) spans from May to August, the second (postrera, Po) from September to October, and the third (apante, Ap) from November to January.", out.width = '80%', echo=FALSE}
+Note: the first season (primera, Pr) spans from May to August, the second (postrera, Po) from September to October, and the third (apante, Ap) from November to January.", out.width = '80%', echo=FALSE, fig.align="center"}
 knitr::include_graphics("cbean.png")
@@
 
+\pagebreak
+
 ## Trends in climate variability in Norway and Sweden
 
-Pick some random points in Norway and Sweden and check how the trends on temperature indices over the last 20 years.
+We randomly selected 100 points in hexagonal within the coordinates 7$^\circ$ and 17$^\circ$ W, and 59 $^\circ$ and 63 $^\circ$ N, that comprises Norway and Sweden before the Arctic Circle.
+We compute the temperature indices from 2000-01-01 to 2019-12-31 using the function `temperature()` with the method for objects of class 'sf'.
+The temperature data is fetched from from the NASA Langley Research Center POWER Project funded through the NASA Earth Science Directorate Applied Science Program (<https://power.larc.nasa.gov/>), using the R package `nasapower` [@Sparks2018].
 
-## Impact of early thaw on vole populations
+```{r nordic, message=FALSE, echo=TRUE, eval=FALSE}
+library("climatrends")
+library("sf")
+library("nasapower")
 
-From [Johnsen et al (2016)](https://doi.org/10.1002/ece3.2635) and [Boonstra et al (2016)](http://bioscience.oxfordjournals.org/content/early/2016/07/19/biosci.biw080.full.pdf).
-The snowpack of European boreal forest contains relatively dense snow because of snow falling at temperatures just below freezing and because of the formation of ice crusts owing to thaw–freeze cycles during the relatively mild winters.
-The formation of ice due to freeze–thaw cycles during winter may be limiting vole availability to bilberries and other edible food items by encrusting them in ice and therefore limiting the peak phase of the cycles.
-If there falls snow in autumn but then it becomes above zero and the snow starts melting but then it starts freezing again, an ice layer is formed which has important implications in ecology.
-This might be calculated / predicted with the `late_frost()` function, but then applied as 'early thaw' function.
+# create a polygon within the coordinates 7, 17, 59, 63
+e <- matrix(c(7, 59, 17, 59, 17, 63,
+              7, 63, 7, 59), 
+            nrow = 5, ncol = 2, byrow = TRUE)
+
+e <- st_polygon(list(e))
+
+# sample 100 points in the hexagonal type
+p <- st_sample(e, 100, type = "hexagonal")
+p <- st_as_sf(p, crs = 4326)
+
+# compute the temperature indices using the random points 
+temp <- temperature(p, 
+                    day.one = "2000-01-01", 
+                    last.day = "2019-12-31", 
+                    timeseries = TRUE, 
+                    intervals = 365)
+
+```
+
+We then select the indices CSDI (cold spell duration of night temperature), WSDI (warm spell duration of day temperature), and their associated indices the T10p (the 10th percentile of night temperature) and T90p (the 90th percentile of day temperature), in Figure 2.
+Plots are generated with `ggplot2` [@ggplot2] and `patchwork` [@patchwork].
+
+The trends show a decrease in the cold spell duration (number of consecutive cold nights bellow the 10th percentile) and warm spell duration (number of consecutive warm days above the 90th percentile).
+However, the values of the percentiles show an increase over the time series.
+The T10p index shows a decrease around the year of 2010, but again rises up to the a value around the -10 $^\circ$C, meaning that the could nights are becoming a bit warmer over the time.
+The T90p index also shows an increase in the temperature across the sampled area, with the average 90th percentile rising from ~ 16 $^\circ$C to ~ 18 $^\circ$C over the time series.
+
+```{r fig_nordic, fig.cap="Fig.
+2.
+Trends in temperature indices across Southern Norway and Sweeden from 2000 to 2019.
+CSDI, maximum cold spell duration, consecutive nights with temperature < 10th percentile.
+WSDI, maximum warm spell duration, consecutive days with temperature > 90th percentile.
+T10p, the 10th percentile of night temperature.
+T90p, the 90th percentile of day temperature.
+Red line indicates the historical mean of each index in the time series.
+Blue line indicates the smoothed trends in each index using the 'loess' method.", out.width = '80%', echo=FALSE, fig.align="center"}
+knitr::include_graphics("nordic.png")
+```
 
@@
 
-Integration with other datasets as they become available in `R` via API client packages. New indices related to the physiology of crops to be implemented while I work on the rice data.
+The package can support the integration with other datasets as they become available in `R` via API client packages. Also new indices related to the physiology of crops could be implemented.
 
@@
 
-# Authors' contributions
-
-KdS and JvE conceived the ideas of implementing an `R` package. JvE and SØS assisted in the implementation of methods. KdS wrote and documented the software. [add other contributions...]
-
 # Data availability statement
 
-The source code has been archived at [**add link**] as `climatrends` version **X.Y.Z**.
 To explore the latest functionalities of `climatrends`, please check the package's updates at CRAN (https://cran.r-project.org/package=climatrends).

### 2020-10-02 inst/paper/climatology.R

@@ -5,20 +5,20 @@ library("nasapower")
 library("tidyverse")
 library("mapview")
 
-# e <- matrix(c(7, 59,
-#               17, 59,
-#               17, 63,
-#               7, 63,
-#               7, 59),
-#             nrow = 5, ncol = 2, byrow = TRUE)
-
-e <- matrix(c(-7.5, 37.2,
-                 -2, 37.2,
-                 -2, 40.5,
-              -7.5, 40.5,
-              -7.5, 37.2),
+e <- matrix(c(7, 59,
+              17, 59,
+              17, 63,
+              7, 63,
+              7, 59),
             nrow = 5, ncol = 2, byrow = TRUE)
-e
+
+# e <- matrix(c(-7.5, 37.2,
+#                  -2, 37.2,
+#                  -2, 40.5,
+#               -7.5, 40.5,
+#               -7.5, 37.2),
+#             nrow = 5, ncol = 2, byrow = TRUE)
+# e
 
 e <- st_polygon(list(e))
 plot(e)
@@ -38,6 +38,7 @@ save(temp, file = "inst/paper/climatology.rda")
 
 i <- c("CSDI","WSDI")
 
+#i <- c("TR", "minDT", "maxDT", "maxNT", "minNT", "DTR")
 
 temp %>% 
   filter(index %in% i) %>% 
@@ -69,7 +70,7 @@ i <- c("T10p","T90p")
 
 
 temp %>% 
-  #filter(index %in% i) %>% 
+  filter(index %in% i) %>% 
   group_by(index) %>% 
   mutate(ab = mean(value)) %>% 
   ggplot() +
@@ -95,5 +96,11 @@ temp %>%
 gg2
 
 
-gg1 / gg2
+pl <- gg1 / gg2
 
+ggsave(paste0("inst/paper/nordic.png"),
+       plot = pl,
+       width = 15,
+       height = 15,
+       dpi = 800,
+       units = "cm")

### 2021-03-14 inst/paper/paper.Rmd
update-urls

@@
 
->Kauê de Sousa^1,2[*]^, Magne Neby^3^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
+\pagenumbering{gobble}
+
+>Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
 ^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
 ^2^ Bioversity International, 00054 Maccarese, Rome, Italy    
-^3^ Department of Forest and Field Sciences, Inland Norway University of Applied Sciences, 2480 Koppang, Norway    
 ^[*]^ Correspondence should be addressed to: <kaue.desousa@inn.no>
@@
 
-The data used here is available as supplementary material (**zenodo ref**) as `cbean`.
+The data used here is available as supplementary material as `cbean`.
 This contains a data.frame with a Plackett-Luce grouped rankings, the geographical coordinates of each sampled plot and the planting dates from where each farmer decided to start the experiment.

### 2022-02-10 inst/paper/paper.Rmd
travis

@@
     index: 1
-  - name: Bioversity International, Rome, Italy
+  - name: Systems Transformation Group, Bioversity International, Montpellier, France
     index: 2
@@
 citation_author: de Sousa et. al.
-date: "A preprint"
-year: 2020
+year: 2022
 bibliography: paper.bib
-csl: citationstyle.csl
-output:
-  pdf_document:
-    keep_tex: false
-header-includes: \usepackage{caption} \captionsetup[figure]{labelformat=empty}
+output: rticles::joss_article
+journal: JOSS
+# output:
+#   pdf_document:
+#     keep_tex: false
+# header-includes: \usepackage{caption} \captionsetup[figure]{labelformat=empty}
 ---
@@
 library("PlackettLuce")
-load("cbean.rda")
 ```
-
-\pagenumbering{gobble}
-
->Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
-^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
-^2^ Bioversity International, 00054 Maccarese, Rome, Italy    
-^[*]^ Correspondence should be addressed to: <kaue.desousa@inn.no>
-
-\
 
@@
 
-Abiotic factors play an important role in most ecological and crop systems that depend on certain levels of temperature, light and precipitation (and their interplay) to initiate important physiological events [@PlantEcology].
-In the walk of climate change, understanding how these factors drive the physiological processes is a key approach to provide recommendations for adaptation and biodiversity conservation.
+Abiotic factors play an important role in most ecological and crop systems that depend on certain levels of temperature, light and precipitation to initiate important physiological events [@PlantEcology].
+Understanding how these factors drive the physiological processes is a key approach to provide recommendations for adaptation and biodiversity conservation in applied ecology studies.
 The package `climatrends` aims to provide the methods in R [@RCoreTeam] to compute precipitation and temperature indices that serve as input for climate and crop models [@vanEtten2019; @Kehel2016], trends in climate change [@Aguilar2005; @deSousa2018] and applied ecology [@Prentice1992; @YLiu2018].
 
-# Methods and features
+# Implementation
 
-## Implementation
-
-Six main functions are provided (Table 1), `crop_sensitive()`, `ETo()`, `GDD()`, `late_frost()`, `rainfall()` and `temperature()` with a default method for numeric 'vector' and additional methods implemented via the package `methods` [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
-The last two methods are designed to fetch data from cloud sources, until now from the packages `nasapower` [@Sparks2018] and `chirps` [@chirps].
+Six main functions are provided (Table 1), with a default method for numeric 'vector' and additional methods implemented via the package `methods` [@RCoreTeam] for classes 'matrix' (or array), 'data.frame', and 'sf' (of geometry POINT or POLYGON) [@sf].
+The last two methods are designed to fetch data from cloud sources, currently from the packages `nasapower` [@Sparks2018] and `chirps` [@chirps].
 
@@
 
-These functions started as a set of scripts to compute indices in citizen science trials.
-In these trials.
-Aiming to capture the environmental variation across different sites, which can differ since each data-point generally have a different starting day and duration, the arguments `day.one` and `span` are vectorised and may be variable across data-points.
-For time series analysis, where fixed periods are defined across many locations, the indices can be adjusted with the argument `last.day` linked to the argument `day.one`.
+These functions started as a set of scripts to compute indices from on-farm testing sites following a citizen science approach [@vanEtten2019].
+Aiming to capture the environmental variation across different sites, which can differ as each on-farm trial generally have a different starting day and duration, the arguments `day.one` and `span` are vectorised and may be used to indicate the starting date for each data-point and the duration of the timespan to be considered for the computation of the indices.
+For time series analysis, fixed periods can be adjusted with the argument `last.day` linked to the argument `day.one`.
 
@@
 
-## Crop-related indices
+## Crop ecology indices
 
@@
 
-The crop sensitive indices available in **climatrends** are described in Table 3.
+The crop ecology indices available in **climatrends** are described in Table 3.
 These indices were previously used in crop models to project the impacts of climate change on crop yield [@Challinor2016; @Trnka2014].
@@
 
-# compute the number of days required to accumulate
-# gdd from planting date to maturity
+# number of days required to accumulate gdd from planting date to maturity
 gdd <- GDD(modis, 
@@
 
-# add gdd to the cbean data and take the average 
-# of gdd per season
+# gdd to the cbean data and take the average gdd per season
 cbean %<>%  
@@
 
-To compute the temperature indices we use the array with tempeture data, the vector with planting dates, and the seasonal averaged degree-days passed as a vector using the argument `span`.
-`temperature()` concatenates the data from the given `day.one` to the given `span` and compute the indices for each row.
+To compute the temperature indices we use the array with temperature data, the vector with planting dates, and the seasonal averaged degree-days passed as a vector using the argument `span`.
+The function `temperature()` concatenates the data from the given `day.one` to the given `span` and computes the indices for each row.
 
@@
 ```{r pltree_bean, message=FALSE, echo=TRUE, eval=FALSE}
-# compute the temperature indices from planting date to the 
+# temperature indices from planting date to the 
 # number of days required to accumulate the gdd in each season
@@
 
-# combine the indices with the main data
 cbean <- cbind(cbean, temp)
@@
 The upper node splits with 49 summer days (SU).
-We can interpret these results as that differences in growing performance of common beans is led by a considerable amount of diurnal temperature above a warmer threshold of 30 $^\circ C$ (in this case >70% of the growing days) and warmer nights (> 18.7 $^\circ C$).
+We can interpret these results as that the on-farm performance of common beans varieties is led by heat accumulation of diurnal temperature above 30 $^\circ C$ (in this case >70% of the growing days) and warmer nights (> 18.7 $^\circ C$).
 
@@
 1.
-Application of climatrends functions to support the analysis of a citizen-science data testing 11 common bean varieties in Nicaragua.
+Application of climatrends functions to support the analysis of on-farm trial data.
 (A) Days required to reach 900 growing-degree days from planting date calculated using the function GDD().
 (B) Maximum night temperature (°C) distributed across seasons computed using the function temperature().
-(C) Plackett-Luce Tree showing the probability of one common bean variety has to win against the others (axys X) in three different nodes splitted with the summer days (day temperature > 30 °C) and maximum night temperature (°C).
+(C) Plackett-Luce Tree showing the probability that a given variety to outperform the other varieties (axys X) in three different nodes splitted with the summer days (day temperature > 30 °C) and maximum night temperature (°C).
 Note: the first season (primera, Pr) spans from May to August, the second (postrera, Po) from September to October, and the third (apante, Ap) from November to January.", out.width = '80%', echo=FALSE, fig.align="center"}
@@
 We compute the temperature indices from 2000-01-01 to 2019-12-31 using the function `temperature()` with the method for objects of class 'sf'.
-The temperature data is fetched from from the NASA Langley Research Center POWER Project funded through the NASA Earth Science Directorate Applied Science Program (<https://power.larc.nasa.gov/>), using the R package `nasapower` [@Sparks2018].
+The temperature data is fetched from the NASA Langley Research Center POWER Project funded through the NASA Earth Science Directorate Applied Science Program (<https://power.larc.nasa.gov/>), using the R package `nasapower` [@Sparks2018].
 
@@
 
-The package can support the integration with other datasets as they become available in `R` via API client packages. Also new indices related to the physiology of crops could be implemented.
+The package can support the integration with other datasets as they become available in `R` via API client packages.
+Also new indices related to the physiology of crops could be implemented.
+To explore the latest functionalities of `climatrends`, please check the package's updates at CRAN (https://cran.r-project.org/package=climatrends).
 
@@
 
-# Data availability statement
-
-To explore the latest functionalities of `climatrends`, please check the package's updates at CRAN (https://cran.r-project.org/package=climatrends).
-
 # References

### 2022-02-10 R/get_timeseries.R

@@ -463,8 +463,8 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
     # define geographic boundaries for lonlat
     lims <- with(lonlat, c(floor(min(lonlat_i[, 1])), 
                            floor(min(lonlat_i[, 2])),
-                           ceiling(max(lonlat_i[, 1])), 
-                           ceiling(max(lonlat_i[, 2]))))
+                           ceiling(max(lonlat_i[, 1]) + .5), 
+                           ceiling(max(lonlat_i[, 2])) + .5))
     
     args <- list(community = community,
                  lonlat = lims,

### 2022-02-10 inst/paper/cbean.R

@@ -41,8 +41,13 @@ temp[1:ncol(temp)] <- lapply(temp[1:ncol(temp)], function(x) round(x, 2))
 # combine the indices with the main data
 cbean <- cbind(cbean, temp, rain)
 
+names(cbean)
+
 # fit a PL tree
-plt <- pltree(G ~ maxNT + SU, data = cbean, minsize = 50)
+plt <- pltree(G ~ maxNT + SU + Rx5day, 
+              data = cbean, 
+              minsize = 50,
+              gamma = TRUE)
 
 plot(plt)
 

### 2022-02-14 R/get_timeseries.R
v0.2

@@ -213,6 +213,7 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
   
   return(r)
 }
+
 #' Set up span length and organise dates
 #' 
 #' @param day.one the first day
@@ -399,10 +400,10 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
 #' @param pars character vector of solar, meteorological or climatology parameters 
 #' to download. See help("parameters", "nasapower") for details.
 #' @examples 
-#'   
+#'  
 #' library("nasapower")
-#' lonlat <- data.frame(lon = c(-66.48, -83.08, -66.45, -66.4),
-#'                      lat = c(-4.60, 9.85, -5.19, -0.15))
+#' lonlat <- data.frame(lon = c(-66.48),# -83.08, -66.45, -66.4),
+#'                      lat = c(-4.60))#, 9.85, -5.19, -0.15))
 #' 
 #' .nasapower(dates = c("2010-01-01", "2010-01-30"),
 #'            lonlat = lonlat,
@@ -411,7 +412,11 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
 #' 
 #'                      
 #' @noRd
-.nasapower <- function(dates, lonlat, pars, community = NULL, temporal_average = NULL){
+.nasapower <- function(dates, 
+                       lonlat, 
+                       pars, 
+                       community = NULL, 
+                       temporal_average = NULL){
   
   message("Getting climate data from NASA POWER \n")
   
@@ -431,7 +436,7 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
     
     h <- stats::hclust(h)
     
-    regions <- stats::cutree(h, h = 5)
+    regions <- stats::cutree(h, h = 3)
     
     nregions <- max(regions)
   }
@@ -460,17 +465,17 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
     r_i <- which(regions == i)
     lonlat_i <- lonlat[r_i, ]
     
-    # define geographic boundaries for lonlat
-    lims <- with(lonlat, c(floor(min(lonlat_i[, 1])), 
+    #define geographic boundaries for lonlat
+    lims <- with(lonlat, c(floor(min(lonlat_i[, 1])),
                            floor(min(lonlat_i[, 2])),
-                           ceiling(max(lonlat_i[, 1]) + .5), 
-                           ceiling(max(lonlat_i[, 2])) + .5))
-    
[…]

### 2022-02-14 inst/paper/climatology.R

@@ -28,9 +28,13 @@ p <- st_as_sf(p, crs = 4326)
 
 mapview(p)
 
+p <- matrix(as.vector(unlist(p)), nrow = nrow(p), ncol = 2, by = TRUE)
+
+p <- as.data.frame(p)
+
 temp <- temperature(p, 
                     day.one = "2000-01-01", 
-                    last.day = "2019-12-31", 
+                    last.day = "2000-12-31", 
                     timeseries = TRUE, 
                     intervals = 365)
 

### 2022-02-23 R/GDD.R
v0.2

@@ -63,8 +63,7 @@
 #'   adjust tmax = tbase_max if tmax < tbase_max
 #' 
 #' @references 
-#' Prentice I. C., et al. (1992) Journal of Biogeography, 19(2), 117. 
-#' \doi{https://doi.org/10.2307/2845499}
+#' Prentice I. C., et al. (1992) Journal of Biogeography, 19(2), 117.
 #' 
 #' Baskerville, G., & Emin, P. (1969). Ecology, 50(3), 514-517. 
 #' \doi{https://doi.org/10.2307/1933912}

### 2022-02-23 R/cumMLWS.R

@@ -1,101 +1,0 @@
-#' Cumulative sum of rainfall
-#' 
-#' Returns a vector with the cumulative sum of 
-#'  the maximum length of wet spell (MLWS)
-#'  
-#' @param object a numeric vector
-#' @return a vector with the cumulative sum of MLWS
-#' @examples 
-#' 
-#' # Example 1
-#' 
-#' rain <- c(0,0.2,1.4,6.1,1.4,5.1,1.5,1.6,0.1,0,7,6,4,6,1.1,1.2,1.5,0)
-#' 
-#' cumMLWS(rain)
-#' 
-#' # should return this vector
-#' # raincum <- c(0,0,1,2,3,4,5,6,6,6,6,6,6,6,6,6,7,7)
-#' 
-#' # Example 2
-#' 
-#' rain2 <- c(1,0,1,1,0,0,1,1,1,0,0,1,1)
-#' 
-#' cumMLWS(rain2)
-#' 
-#' # should return this 
-#' # raincum2 <- c(1,1,1,2,2,2,2,2,3,3,3,3,3)
-#' 
-#' @export
-cumMLWS <- function(object){
-  
-  # values higher equal or higher than 1 are coded as 1
-  # lower than 1 are coded as 0
-  r <- ifelse(object >= 1, 1, 0)
-  
-  # get the list with length encoding showing 
-  # the values with 0s and 1s to build the vector
-  len <- rle(r)
-  
-  # now start the vector by making the first 
-  # cumulative values
-  result <- integer()
-  
-  # check if the first sequence is with 1s or 0s
-  zero <- which(len$values == 0)[1]
-  one  <- which(len$values == 1)[1]
-  
-  # if it starts with 0 then 
-  # start by a sequence with 0s until 
-  # the 1s starts and then do a 
-  # cumulative sequence until the third sequence starts
-  if(zero == 1) {
-    result <- rep(0, len$lengths[1])
-    result <- c(result, seq_len(len$lengths[2]))
-  }
-  
-  # if starts with 1 then add the cumulative 
-  # and then repeats the last value until 
-  # the third sequence starts
-  if(one == 1) {
-    result <- seq_len(len$lengths[1])
-    result <- c(result, 
-                rep(result[length(result)], seq_len(len$lengths[2])))
-  }
-  
-  result
-  
-  # if only two sequences then return the result 
-  if (length(len$lengths) == 2) {
-    return(result)
-  }
-  
-  # or continue with the sequence
-  # remove the first two sequences
[…]

### 2022-02-23 R/cumrain.R

@@ -1,0 +1,58 @@
+#' Cumulative sum of rainfall days
+#' 
+#' Returns a vector with the cumulative sum of 
+#'  the maximum length of wet spell (MLWS)
+#'  
+#' @param x a numeric vector
+#' @return a vector with the cumulative sum of MLWS
+#' @examples 
+#' 
+#' # Example 1
+#' 
+#' rain <- c(0,0.2,1.4,6.1,1.4,5.1,1.5,1.6,0.1,0,7,6,4,6,1.1,1.2,1.5,0)
+#' 
+#' cumrain(rain)
+#' 
+#' # should return this vector
+#' # raincum <- c(0,0,1,2,3,4,5,6,6,6,6,6,6,6,6,6,7,7)
+#' 
+#' # Example 2
+#' 
+#' rain2 <- c(1,0,1,1,0,0,1,1,1,0,0,1,1)
+#' 
+#' cumrain(rain2)
+#' 
+#' # should return this 
+#' # raincum2 <- c(1,1,1,2,2,2,2,2,3,3,3,3,3)
+#' 
+#' @export
+cumrain <- function(x){
+  
+  a <- ifelse(x >= 1, 1, 0)
+  for(i in 2:length(a))  a[i] <- ifelse(a[i] == 1, sum(a[i-1], a[i]), 0)
+  cummax(a)
+  
+}
+
+#' Cumulative sum of dry days
+#' 
+#' Returns a vector with the cumulative sum of 
+#'  the maximum length of dry spell (MLDS)
+#'  
+#' @param x a numeric vector
+#' @return a vector with the cumulative sum of MLDS
+#' @examples 
+#' 
+#' rain <- c(0,0.2,0.4,0.1,0.4,5.1,1.5,1.6,0.1,0,7,6,4,6,0.1,1.2,0.5,0)
+#' 
+#' cumdrought(rain)
+#' 
+#' @export
+cumdrought <- function(x){
+  
+  a <- ifelse(x < 1, 1, 0)
+  for(i in 2:length(a))  a[i] <- ifelse(a[i] == 1, sum(a[i-1], a[i]), 0)
+  cummax(a)
+  
+}
+

### 2023-01-04 R/clima_data.R
0.4-0003

@@ -4,10 +4,13 @@
 #' 
 #' @details 
 #' 
-#' \code{temp_dat}: array with two layers 1) day temperature and 2) night temperature. 
-#'  An excerpt of MODIS (MYD11A1) data to represent an example 
-#'  of the input data in \code{temperature()} and \code{get_timeseries()} 
-#'  when an array is provided.
+#' \code{commonbean}: a list with three data sets 1) on-farm variety performance data,
+#'  2) chirps data, and 3) modis data. 
+#' 
+#' \code{rain_dat}: matrix with precipitation from CHIRPS. An excerpt to represent 
+#'  an example of the input data in \code{rainfall()} or 
+#' 
+#' \code{get_timeseries()} when a matrix is provided.
 #' 
 #' \code{rain_dat}: matrix with precipitation from CHIRPS. An excerpt to represent 
 #'  an example of the input data in \code{rainfall()} or 
@@ -19,7 +22,7 @@
 #' \code{lonlatsf}: a 'sf' object with 'POINT' geometry with five random points around 
 #'  the municipality of Sinop, Brazil 
 #'  
-#' In \code{temp_dat} and \code{rain_dat}, rows represents the coordinates for the 
+#' \code{temp_dat} and \code{rain_dat}, rows represents the coordinates for the 
 #'  given lonlat provided in \code{raster::extract()} and colunms represents the dates 
 #'  from the observed precipitation/temperature.
 #' 
@@ -30,6 +33,7 @@
 #' @aliases temp_dat
 #' @aliases rain_dat
 #' @aliases innlandet
+#' @aliases commonbean
 #' @keywords datasets
 #' @source 
 #' 
@@ -37,12 +41,16 @@
 #' with stations—a new environmental record for monitoring extremes. 
 #' Scientific Data, 2, 150066. \doi{https://doi.org/10.1038/sdata.2015.66}
 #' 
+#' van Etten, J. et. al. (2019). PNAS 116(10) 4194-4199 
+#'    \doi{http://doi.org/10.1073/pnas.1813720116}
+#' 
 #' Wan Z, Hook S, Hulley G (2015) MYD11A1 MODIS/Aqua Land Surface 
 #' Temperature/Emissivity 8-Day L3 Global 1km
 #' SIN Grid V006 \doi{http://doi.org/10.5067/MODIS/MYD11A2.006}.
 #' 
 #' @format 
[…]

### 2023-01-04 inst/paper/cbean.R

@@ -1,142 +1,0 @@
-library("climatrends")
-library("gosset")
-library("tidyverse")
-library("PlackettLuce")
-library("patchwork")
-library("qvcalc")
-library("ggparty")
-
-
-data("beans", package = "PlackettLuce")
-
-
-# compute the number of days required to accumulate gdd from 
-# planting date to maturity
-gdd <- GDD(modis, 
-           day.one = beans$planting_date, 
-           degree.days = 900, 
-           return.as = "ndays")
-
-# add gdd to the cbean data
-# and take the average of gdd per season
-beans %<>%  
-  mutate(gdd = gdd$gdd) %>% 
-  group_by(season) %>% 
-  mutate(gdds = as.integer(mean(gdd)))
-  
-# compute the temperature indices from planting date to the number of days 
-# required to accumulate the gdd in each season
-temp <- temperature(modis, 
-                    day.one = beans$planting_date, 
-                    span = beans$gdds)
-
-# round values to 3 decimals, this will not change the main result, but will 
-# help in visualizing the tree 
-temp[1:ncol(temp)] <- lapply(temp[1:ncol(temp)], function(x) round(x, 2))
-
-# combine the indices with the main data
-G <- rank_tricot(beans, 
-                 items = 
-                 input = )
-
-names(cbean)
-
-# fit a PL tree
-plt <- pltree(G ~ maxNT + SU + Rx5day, 
-              data = cbean, 
-              minsize = 50,
-              gamma = TRUE)
-
-plot(plt)
-
-# ..................................
-# ..................................
-# Prepare charts ####
-A <- 
-ggplot(cbean, aes(gdd)) +
-  geom_density() +
-  labs(x = "GDD", y = "Density", title = "(a)") +
-  theme_bw() + 
-  theme(panel.background = element_blank(), 
-        panel.grid = element_blank(),
-        plot.title = element_text(size=16, 
-                                  colour = "grey20", 
-                                  face = "bold"),
-        axis.text.x = element_text(size = 10, angle = 0,
-                                   face="plain", colour = "grey20"),
-        axis.title.x = element_text(size=10, colour = "grey20"),
-        axis.text.y = element_text(size=10, angle = 0,
-                                   hjust=1, vjust=0.5,
-                                   face="plain", colour = "grey20"),
-        axis.title.y = element_text(size=10, colour = "grey20"))
-
-B <- 
[…]

### 2023-01-05 R/get_timeseries.R
paper

@@ -392,7 +392,6 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
   
 }
 
-
 #' Get data from NASAPOWER using nasapower::get_power()
 #' 
 #' @param dates character with first and final date in the format YYYY-MM-DD
@@ -555,3 +554,4 @@ get_timeseries.array <- function(object, day.one, span = NULL, last.day = NULL,
   return(result)
   
 }
+

### 2023-01-05 R/internal_functions.R

@@ -176,3 +176,4 @@
   return(x)
   
 }
+

### 2023-01-05 inst/paper/climatology.R
add-new-vignette

@@ -42,8 +42,6 @@ save(temp, file = "inst/paper/climatology.rda")
 
 i <- c("CSDI","WSDI")
 
-#i <- c("TR", "minDT", "maxDT", "maxNT", "minNT", "DTR")
-
 temp %>% 
   filter(index %in% i) %>% 
   group_by(index) %>% 

### 2023-01-05 data-raw/DATASET.R
package-after-testing-as-cran

@@ -1,0 +1,5 @@
+## code to prepare `DATASET` dataset goes here
+
+usethis::use_data(commonbean, overwrite = TRUE, compress = "xz")
+
+dimnames(commonbean[[2]])

### 2023-01-05 R/get_timeseries.R
enable-return-matrix-in-get_timeseries

@@ -16,6 +16,7 @@
 #'  any other object that can be coerced to \code{Date} (e.g. integer, character 
 #'  YYYY-MM-DD)  for the last day of the time series
 #' @param data.from character, for the source of climate data. See details.
+#' @param as.matrix logical, optional, to return a matrix or array instead of a data.frame 
 #' @param ... additional arguments passed to methods. See details.
 #' @details 
 #' The \code{default} method and the \code{sf} method assumes that the climate 
@@ -38,7 +39,7 @@
 #' @return A list with class \code{clima_ls} with data.frame(s) with 
 #'  the class \code{clima_df} 
 #' @family GET functions
-#' @examples 
+#' @examplesIf interactive()
 #' # Using local sources
 #' # an array with temperature data
 #' data("temp_dat", package = "climatrends")
@@ -53,19 +54,18 @@
 #' 
 #' get_timeseries(rain_dat, "2013-10-28", span = span)
 #' 
-#' # library("nasapower")
-#' # library("sf")
-#' # # Fetch data from NASA POWER using 'sf' method
-#' # data("lonlatsf", package = "climatrends")
-#' # 
-#' # g <- get_timeseries(lonlatsf, 
-#' #                     day.one = "2018-05-16", 
-#' #                     last.day = "2018-05-30",
-#' #                     pars = c("PRECTOT", "T2M", "T10M"))
-#' # 
-#' # 
-#' # g
+#' \donttest{
+#' # data can be returned as matrix
+#' library("sf")
+#' # Fetch data from NASA POWER using 'sf' method
+#' data("lonlatsf", package = "climatrends")
 #' 
+#' g <- get_timeseries(object = lonlatsf,
+#'                     day.one = "2018-05-16",
+#'                     last.day = "2018-05-30",
+#'                     pars = c("PRECTOTCORR", "T2M"),
+#'                     as.matrix = TRUE)
+#' }
 #' @importFrom stats dist hclust cutree
 #' @export
 get_timeseries <- function(object, day.one, ...) {
@@ -76,7 +76,10 @@ get_timeseries <- function(object, day.one, ...) {
 
 #' @rdname get_timeseries
 #' @export
-get_timeseries.default <- function(object, day.one, span = NULL, last.day = NULL,
+get_timeseries.default <- function(object, day.one, 
+                                   span = NULL, 
+                                   last.day = NULL, 
[…]

### 2023-01-05 inst/paper/climatology.R
citation

@@ -1,108 +1,0 @@
-library("climatrends")
-library("patchwork")
-library("sf")
-library("nasapower")
-library("tidyverse")
-library("mapview")
-
-e <- matrix(c(7, 59,
-              17, 59,
-              17, 63,
-              7, 63,
-              7, 59),
-            nrow = 5, ncol = 2, byrow = TRUE)
-
-# e <- matrix(c(-7.5, 37.2,
-#                  -2, 37.2,
-#                  -2, 40.5,
-#               -7.5, 40.5,
-#               -7.5, 37.2),
-#             nrow = 5, ncol = 2, byrow = TRUE)
-# e
-
-e <- st_polygon(list(e))
-plot(e)
-set.seed(435)
-p <- st_sample(e, 100, type = "hexagonal")
-p <- st_as_sf(p, crs = 4326)
-
-mapview(p)
-
-p <- matrix(as.vector(unlist(p)), nrow = nrow(p), ncol = 2, by = TRUE)
-
-p <- as.data.frame(p)
-
-temp <- temperature(p, 
-                    day.one = "2000-01-01", 
-                    last.day = "2000-12-31", 
-                    timeseries = TRUE, 
-                    intervals = 365)
-
-save(temp, file = "inst/paper/climatology.rda")
-
-i <- c("CSDI","WSDI")
-
-temp %>% 
-  filter(index %in% i) %>% 
-  group_by(index) %>% 
-  mutate(ab = mean(value)) %>% 
-  ggplot() +
-  geom_line(aes(x = date, y = value, group = id)) + 
-  geom_hline(aes(yintercept = ab), colour = "red", lwd = 0.7) +
-  geom_smooth(aes(x = date, y = value), method = "loess", se = FALSE) +
-  facet_wrap(. ~ index, scales = "free") +
-  labs(x = "", y = "Index (days)") +
-  theme(axis.text.y = element_text(size = 12, angle = 0, hjust = 1, 
-                                   vjust = 0.5, face = "plain", colour = "black"),
-        axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5, 
-                                   vjust = 1, face = "plain", colour = "black"),
-        axis.title = element_text(size = 12, face = "plain"),
-        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
-        plot.background = element_blank(),
-        panel.background = element_blank(),
-        strip.text.x = element_text(size = 12, colour = "black"),
[…]

### 2025-11-21 R/ETo.R
website

@@ -54,7 +54,6 @@
 #' Brouwer C. & Heibloem M. (1986). Irrigation water management: 
 #' Irrigation water needs. Food and Agriculture Organization of The 
 #' United Nations, Rome, Italy. 
-#' \url{https://www.fao.org/3/S2022E/s2022e00.htm}
 #' 
 #' @examples
 #' # the default method

### 2025-11-21 vignettes/Overview.R

@@ -3,27 +3,26 @@ library("climatrends")
 
 data("innlandet", package = "climatrends")
 
-temp1 <- temperature(innlandet$tmax, innlandet$tmin)
+temp1 = temperature(innlandet$tmax, innlandet$tmin)
 
 temp1
 
-
 ## ----temperature2, message=TRUE, eval=TRUE, echo=TRUE-------------------------
-temp2 <- temperature(innlandet$tmax, innlandet$tmin,
-                     dates = innlandet$dates,
-                     timeseries = TRUE, 
-                     intervals = 30)
+temp2 = temperature(innlandet$tmax, innlandet$tmin,
+                    dates = innlandet$dates,
+                    timeseries = TRUE, 
+                    intervals = 30)
 
 temp2
 
 ## ----gdd, message=TRUE, eval=TRUE, echo=TRUE----------------------------------
-gdd <- GDD(innlandet$tmax, innlandet$tmin, tbase = 2, equation = "b")
+gdd = GDD(innlandet$tmax, innlandet$tmin, tbase = 2, equation = "b")
 
 gdd
 
 ## ----gdd2, message=FALSE, eval=FALSE, echo=TRUE-------------------------------
-#  lonlat <- data.frame(lon = 129.19,
-#                       lat = 36.39)
+#  lonlat = data.frame(lon = 129.19,
+#                      lat = 36.39)
 #  
 #  GDD(lonlat,
 #      day.one = "2019-04-01",
@@ -32,7 +31,7 @@ gdd
 #      return.as = "ndays")
 
 ## ----latefrost, message=TRUE, eval=TRUE, echo=TRUE----------------------------
-lf <- late_frost(innlandet$tmax, 
+lf = late_frost(innlandet$tmax, 
                  innlandet$tmin, 
                  dates = innlandet$date, 
                  base = 2)
@@ -42,10 +41,10 @@ lf
 ## ----rain, message=FALSE, eval=FALSE, echo=TRUE-------------------------------
 #  library("nasapower")
 #  
-#  lonlat <- data.frame(lon = c(-73.3, -74.5),
+#  lonlat = data.frame(lon = c(-73.3, -74.5),
 #                       lat = c(-6.1, - 6.2))
 #  
-#  rain <- rainfall(lonlat,
+#  rain = rainfall(lonlat,
 #                   day.one = "2018-11-01",
 #                   last.day = "2018-12-31")
 #  
[…]
